---
title: "Avatar"
description: "Avatar component for displaying user images, initials, or icons"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Avatar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(muiMaterial)
library(shiny)
```

Avatars are fundamental UI elements used to represent users throughout Material Design interfaces. They appear in tables, menus, dialogs, and more. The Avatar component can display images, text initials, or icons.

## Image avatars

Image avatars are created by passing `src` and `alt` properties to the `Avatar()` component:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1),
      Avatar(
        alt = "Remy Sharp",
        src = "https://mui.com/static/images/avatar/1.jpg"
      ),
      Avatar(
        alt = "Travis Howard",
        src = "https://mui.com/static/images/avatar/2.jpg"
      ),
      Avatar(
        alt = "Cindy Baker",
        src = "https://mui.com/static/images/avatar/3.jpg"
      )
    )
  )
)
```

## Letter avatars

Avatars can display text initials or single characters as children:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1),
      Avatar("H"),
      Avatar(
        "N",
        sx = list(bgcolor = "#FF7043")
      ),
      Avatar(
        "OP",
        sx = list(bgcolor = "#7E57C2")
      )
    )
  )
)
```

## Avatar sizes

Control avatar size using the `sx` prop with `width` and `height` CSS properties:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1, alignItems = "center"),
      Avatar(
        alt = "Remy Sharp",
        src = "https://mui.com/static/images/avatar/1.jpg",
        sx = list(width = 24, height = 24)
      ),
      Avatar(
        alt = "Remy Sharp",
        src = "https://mui.com/static/images/avatar/1.jpg"
      ),
      Avatar(
        alt = "Remy Sharp",
        src = "https://mui.com/static/images/avatar/1.jpg",
        sx = list(width = 56, height = 56)
      )
    )
  )
)
```

## Icon avatars

Use Material Design icons as avatar content:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1),
      Avatar(
        shiny::icon("folder"),
        sx = list(bgcolor = "#E91E63")
      ),
      Avatar(
        shiny::icon("file"),
        sx = list(bgcolor = "#FF1744")
      ),
      Avatar(
        shiny::icon("image"),
        sx = list(bgcolor = "#4CAF50")
      )
    )
  )
)
```

## Avatar variants

Change avatar shape using the `variant` prop. Options include:

- `"circular"` (default)
- `"rounded"`
- `"square"`

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1),
      Avatar(
        "N",
        variant = "circular",
        sx = list(bgcolor = "#FF7043")
      ),
      Avatar(
        "N",
        variant = "rounded",
        sx = list(bgcolor = "#FF7043")
      ),
      Avatar(
        "N",
        variant = "square",
        sx = list(bgcolor = "#FF7043")
      )
    )
  )
)
```

## Avatar fallbacks

When an image fails to load, avatars fall back in the following order:

1. Provided children (text/icon)
2. First letter of `alt` text
3. Generic avatar icon

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 1),
      # Fallback to children
      Avatar(
        alt = "Remy Sharp",
        src = "https://invalid-url.jpg",
        "B",
        sx = list(bgcolor = "#FF7043")
      ),
      # Fallback to first letter of alt
      Avatar(
        alt = "Remy Sharp",
        src = "https://invalid-url.jpg",
        sx = list(bgcolor = "#FF7043")
      ),
      # Fallback to generic icon
      Avatar(
        src = "https://invalid-url.jpg"
      )
    )
  )
)
```

## Avatar groups

Use `AvatarGroup()` to stack multiple avatars. Control the maximum number displayed with the `max` prop:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 2),
      AvatarGroup(
        max = 4,
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        ),
        Avatar(
          alt = "Cindy Baker",
          src = "https://mui.com/static/images/avatar/3.jpg"
        ),
        Avatar(
          alt = "Agnes Walker",
          src = "https://mui.com/static/images/avatar/4.jpg"
        ),
        Avatar(
          alt = "Trevor Henderson",
          src = "https://mui.com/static/images/avatar/5.jpg"
        )
      )
    )
  )
)
```

## Avatar group with total

Control the total number of avatars not shown using the `total` prop:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2),
      AvatarGroup(
        total = 24,
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        ),
        Avatar(
          alt = "Agnes Walker",
          src = "https://mui.com/static/images/avatar/4.jpg"
        ),
        Avatar(
          alt = "Trevor Henderson",
          src = "https://mui.com/static/images/avatar/5.jpg"
        )
      )
    )
  )
)
```

## Avatar group spacing

Control the spacing between avatars in a group using the `spacing` prop:

- `"small"` - compact spacing
- `"medium"` (default) - normal spacing
- Numeric value - custom pixel spacing

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", flexDirection = "column", gap = 2),
      Typography("Small spacing:", variant = "h6"),
      AvatarGroup(
        spacing = "small",
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        ),
        Avatar(
          alt = "Cindy Baker",
          src = "https://mui.com/static/images/avatar/3.jpg"
        )
      ),
      Typography("Medium spacing (default):", variant = "h6"),
      AvatarGroup(
        spacing = "medium",
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        ),
        Avatar(
          alt = "Cindy Baker",
          src = "https://mui.com/static/images/avatar/3.jpg"
        )
      ),
      Typography("Custom spacing (24px):", variant = "h6"),
      AvatarGroup(
        spacing = 24,
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        ),
        Avatar(
          alt = "Cindy Baker",
          src = "https://mui.com/static/images/avatar/3.jpg"
        )
      )
    )
  )
)
```

## Avatar with Badge

Combine avatars with the `Badge()` component to add status indicators:

```{r}
muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 2),
      Badge(
        overlap = "circular",
        anchorOrigin = list(vertical = "bottom", horizontal = "right"),
        variant = "dot",
        sx = list(
          "& .MuiBadge-badge" = list(
            bgcolor = "#44b700",
            boxShadow = "0 0 0 2px white"
          )
        ),
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        )
      ),
      Badge(
        overlap = "circular",
        anchorOrigin = list(vertical = "bottom", horizontal = "right"),
        badgeContent = Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg",
          sx = list(width = 22, height = 22)
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        )
      )
    )
  )
)
```

See the [Material UI Avatar documentation](https://mui.com/material-ui/api/avatar/) and [AvatarGroup documentation](https://mui.com/material-ui/api/avatar-group/) for complete API details.
