---
title: "Backdrop"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Backdrop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-backdrop/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Backdrop

The Backdrop component narrows the user's focus to a particular element on the screen.

The Backdrop signals a state change within the application and can be used for creating loaders, dialogs, and more.
In its simplest form, the Backdrop component will add a dimmed layer over your application.

## Example

The demo below shows a basic Backdrop with a Circular Progress component in the foreground to indicate a loading state.
After clicking **Show Backdrop**, you can click anywhere on the page to close it.

The open state is kept in the URL with
[reactRouter](https://felixluginbuhl.com/reactRouter/): the button links to
`#/backdrop`, and a click on the backdrop navigates back to `#/`.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/backdrop", "Show backdrop"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "backdrop",
        element = Backdrop(
          sx = list(color = "#fff", zIndex = 1201), # theme.zIndex.drawer + 1
          open = TRUE,
          onClick = JS("() => { window.location.hash = '/'; }"),
          CircularProgress(color = "inherit")
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Backdrop from '@mui/material/Backdrop';
import CircularProgress from '@mui/material/CircularProgress';
import Button from '@mui/material/Button';

export default function SimpleBackdrop() {
  const [open, setOpen] = React.useState(false);
  const handleClose = () => {
    setOpen(false);
  };
  const handleOpen = () => {
    setOpen(true);
  };

  return (
    <div>
      <Button onClick={handleOpen}>Show backdrop</Button>
      <Backdrop
        sx={(theme) => ({ color: '#fff', zIndex: theme.zIndex.drawer + 1 })}
        open={open}
        onClick={handleClose}
      >
        <CircularProgress color="inherit" />
      </Backdrop>
    </div>
  );
}
```
</details>

## Transitions

Backdrop uses [Fade](transitions.html#fade) by default.
Use `slots$transition` and `slotProps$transition` to replace it with another transition or to pass transition props.

## Backdrops in Shiny apps

A backdrop is a simple way to block the page while the server works. Render it
from the server with `renderReact()`:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Button.shinyInput("run", "Run long computation", variant = "contained"),
  reactOutput("backdrop"),
  verbatimTextOutput("result")
)

server <- function(input, output, session) {
  busy <- reactiveVal(FALSE)
  result <- reactiveVal(NULL)

  observeEvent(input$run, {
    busy(TRUE)
    # Run the work after the backdrop has been sent to the browser
    later::later(function() {
      Sys.sleep(2) # the long computation
      result(Sys.time())
      busy(FALSE)
    }, 0.1)
  })

  output$backdrop <- renderReact({
    Backdrop(sx = list(color = "#fff", zIndex = 1201), open = busy(), CircularProgress(color = "inherit"))
  })
  output$result <- renderPrint(result())
}

shinyApp(ui, server)
```
