---
title: "Badge"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Badge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Badge component is used to display a small notification or count next to another element. It's commonly used for notification counts, status indicators, or other small pieces of information.

## Basic badge

Examples of badges containing text, using primary and secondary colors. The badge is applied to its children.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        badgeContent = 4,
        color = "primary",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "secondary",
        shiny::icon(name = "envelope")
      )
    )
  )
)
```

## Color

Use the `color` prop to apply theme palette to the badge.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        badgeContent = 4,
        color = "primary",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "secondary",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "success",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "error",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "warning",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 4,
        color = "info",
        shiny::icon(name = "envelope")
      )
    )
  )
)
```

## Badge visibility

The visibility of badges can be controlled using the `invisible` prop.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        badgeContent = 1,
        color = "secondary",
        shiny::icon(name = "envelope")
      ),
      Badge(
        badgeContent = 1,
        invisible = TRUE,
        color = "secondary",
        shiny::icon(name = "envelope")
      )
    )
  )
)
```

The badge hides automatically when `badgeContent` is zero. You can override this with the `showZero` prop.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        color = "secondary",
        badgeContent = 0,
        shiny::icon(name = "envelope")
      ),
      Badge(
        color = "secondary",
        badgeContent = 0,
        showZero = TRUE,
        shiny::icon(name = "envelope")
      )
    )
  )
)
```

## Maximum value

You can use the `max` prop to cap the value of the badge content.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        color = "secondary",
        badgeContent = 99,
        shiny::icon(name = "envelope")
      ),
      Badge(
        color = "secondary",
        badgeContent = 100,
        shiny::icon(name = "envelope")
      ),
      Badge(
        color = "secondary",
        badgeContent = 1000,
        max = 999,
        shiny::icon(name = "envelope")
      )
    )
  )
)
```

## Dot badge

The `variant = "dot"` prop changes a badge into a small dot. This can be used as a notification that something has changed without giving a count.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 2, p = 2),
      Badge(
        color = "secondary",
        variant = "dot",
        shiny::icon(name = "envelope")
      ),
      Badge(
        color = "error",
        variant = "dot",
        shiny::icon(name = "phone")
      ),
      Badge(
        color = "success",
        variant = "dot",
        shiny::icon(name = "check")
      )
    )
  )
)
```

## Badge overlap

You can use the `overlap` prop to place the badge relative to the corner of the wrapped element. Use `"rectangular"` (default) for rectangular children or `"circular"` for circular children like avatars.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(display = "flex", gap = 3, p = 2),
      # Rectangular overlap
      Badge(
        color = "secondary",
        badgeContent = " ",
        overlap = "rectangular",
        Paper(
          sx = list(width = 40, height = 40, bgcolor = "lightgray")
        )
      ),
      # Dot variant with rectangular
      Badge(
        color = "secondary",
        variant = "dot",
        overlap = "rectangular",
        Paper(
          sx = list(width = 40, height = 40, bgcolor = "lightgray")
        )
      ),
      # Circular overlap with Avatar
      Badge(
        color = "secondary",
        badgeContent = " ",
        overlap = "circular",
        Avatar(
          alt = "John Doe",
          src = "https://mui.com/static/images/avatar/1.jpg"
        )
      ),
      # Dot variant with circular
      Badge(
        color = "success",
        variant = "dot",
        overlap = "circular",
        Avatar(
          alt = "Jane Smith",
          src = "https://mui.com/static/images/avatar/2.jpg"
        )
      )
    )
  )
)
```

## Badge alignment

You can use the `anchorOrigin` prop to move the badge to any corner of the wrapped element.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2),
      Typography("Badge Positions", variant = "h5", sx = list(mb = 3)),
      Box(
        sx = list(display = "grid", gridTemplateColumns = "repeat(3, 100px)", gap = 2),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "top", horizontal = "left"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "top", horizontal = "right"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "bottom", horizontal = "left"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "bottom", horizontal = "right"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "top", horizontal = "center"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "bottom", horizontal = "center"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "center", horizontal = "left"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "center", horizontal = "right"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        ),
        Badge(
          badgeContent = 1,
          anchorOrigin = list(vertical = "center", horizontal = "center"),
          Paper(sx = list(width = 40, height = 40, bgcolor = "lightgray"))
        )
      )
    )
  )
)
```

## Badge with avatar

Combine badges with avatars to show online status or other information.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2, display = "flex", gap = 3, flexWrap = "wrap"),
      # Online status indicator
      Badge(
        overlap = "circular",
        anchorOrigin = list(vertical = "bottom", horizontal = "right"),
        variant = "dot",
        sx = list(
          "& .MuiBadge-badge" = list(
            bgcolor = "#44b700",
            boxShadow = "0 0 0 2px white"
          )
        ),
        Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg"
        )
      ),
      # Badge with nested avatar
      Badge(
        overlap = "circular",
        anchorOrigin = list(vertical = "bottom", horizontal = "right"),
        badgeContent = Avatar(
          alt = "Remy Sharp",
          src = "https://mui.com/static/images/avatar/1.jpg",
          sx = list(width = 22, height = 22, border = "2px solid white")
        ),
        Avatar(
          alt = "Travis Howard",
          src = "https://mui.com/static/images/avatar/2.jpg"
        )
      ),
      # Error status
      Badge(
        overlap = "circular",
        anchorOrigin = list(vertical = "bottom", horizontal = "right"),
        variant = "dot",
        sx = list(
          "& .MuiBadge-badge" = list(
            bgcolor = "#f44336"
          )
        ),
        Avatar(
          alt = "Cindy Baker",
          src = "https://mui.com/static/images/avatar/3.jpg"
        )
      )
    )
  )
)
```

## Accessibility

You should provide a full description for badges when displaying counts, especially for important notifications. Use `aria-label` on the parent element to describe the badge content.

```{r}
library(muiMaterial)

muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2),
      IconButton(
        `aria-label` = "100 unread notifications",
        Badge(
          badgeContent = 100,
          color = "error",
          shiny::icon(name = "bell")
        )
      ),
      IconButton(
        `aria-label` = "shopping cart with 5 items",
        Badge(
          badgeContent = 5,
          color = "secondary",
          shiny::icon(name = "shop")
        )
      )
    )
  )
)
```
