---
title: "CSS conflicts with Bootstrap"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{CSS conflicts with Bootstrap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(muiMaterial)
library(shiny)
```

# CSS conflicts with Bootstrap

## Why `muiMaterialPage()`?

`muiMaterialPage()` is the recommended way to build MUI apps with `muiMaterial`. It suppresses Bootstrap by default (`suppressBootstrap = TRUE`), giving MUI components a clean CSS environment where they render exactly as intended.

This matters because MUI has its own design system — its own typography scale, spacing, color palette, and CSS reset — that is independent of Bootstrap. When both frameworks are loaded, their global CSS rules conflict, leading to visual inconsistencies (e.g. text appearing too small, margins being off).

By removing Bootstrap, `muiMaterialPage()` avoids these conflicts entirely. MUI provides its own complete set of input components (`TextField()`, `Button()`, `Select()`, `Slider()`, etc.) through their `.shinyInput()` wrappers, so you don't need Bootstrap-styled inputs.

## Mixing MUI with Bootstrap-dependent packages

You can still use other R packages that depend on Bootstrap (e.g. `shiny`, `bslib`, `DT`, `plotly`, `leaflet`) alongside `muiMaterial`. You just need to account for the CSS conflict.

### Using `muiMaterialPage()` with shiny components

You can add Bootstrap 3 dependency use by default by shiny when using `muiMaterialPage()` with `suppressBootstrap = FALSE`. 

Bootstrap 3 and 4 sets `html { font-size: 10px }`, which breaks MUI's `rem`-based sizing, making components like `Typography()` appear too small. To fix this, wrap MUI components in a `ThemeProvider()` with `htmlFontSize = 10`, which tells MUI to recalculate its `rem` values for the actual root font-size.

```{r, eval=FALSE}
library(shiny)
library(muiMaterial)

muiMaterialPage(
  suppressBootstrap = FALSE, # TRUE by default, FALSE add Bootstrap 3 as dependency
  shiny::tags$h3("muiMaterial with Bootstrap 3"),
  shiny::textInput("name", "Shiny textInput"),
  shiny::actionButton("btn", "Shiny actionButton"),
  ThemeProvider(
    theme = list(typography = list(htmlFontSize = 10)), # fix Bootstrap 3 and html font size 
    Card(
      sx = list(m = 3),
      CardContent(
        Typography("Card title", variant = "h5"),
        Typography("Some content inside a Card component.", variant = "body2")
      )
    )
  )
)
```

### Using MUI components with bslib

The `bslib` R package gives access to Bootstrap 5, which font sizing system is align with MUI components. If `muiMaterialPage()` cannot be use, it is recommended to use bslib functions with `theme = bs_theme(version = 5)`.

As Bootstrap 5 and MUI have two design system, always wrap MUI components in `ScopedCssBaseline()` to ensure their correct rendering.

```{r, eval=FALSE}
library(bslib)
library(muiMaterial)

bslib::page_fluid(
  theme = bs_theme(version = 5),
  shiny::tags$h3("Bootstrap 5 with muiMaterial"),
  shiny::textInput("name", "Shiny textInput"),
  shiny::actionButton("btn", "Shiny actionButton"),
  ScopedCssBaseline(
    Card(
      sx = list(m = 3),
      CardContent(
        Typography("Card title", variant = "h5"),
        Typography("Some content inside a Card component.", variant = "body2")
      )
    )
  )
)
```
