---
title: "Bottom Navigation"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Bottom Navigation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-bottom-navigation/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Bottom Navigation

The Bottom Navigation bar allows movement between primary destinations in an app.

Bottom navigation bars display three to five destinations at the bottom of a screen. Each destination is represented by an icon and an optional text label. When a bottom navigation icon is tapped, the user is taken to the top-level navigation destination associated with that icon.

<div class="alert alert-secondary">
**How the live demos on this page work.** The selected action is kept in the
URL with [reactRouter](https://felixluginbuhl.com/reactRouter/), as in the
[Third-party routing library](#third-party-routing-library) section of the MUI
documentation: each action is a link (`component = "a"` and `href`), and each
route renders the bar with its `value`. In a Shiny app, use
`BottomNavigation.shinyInput()`: see
[Bottom navigation in Shiny apps](#bottom-navigation-in-shiny-apps).
</div>

## Bottom navigation

When there are only **three** actions, display both icons and text labels at all times.

```{r}
actions <- list(
  list(value = "recents", label = "Recents", icon = "restore"),
  list(value = "favorites", label = "Favorites", icon = "favorite"),
  list(value = "nearby", label = "Nearby", icon = "location_on")
)

navigationAction <- function(action, route) {
  BottomNavigationAction(
    component = "a",
    href = sprintf("#/%s/%s", route, action$value),
    value = action$value,
    label = action$label,
    icon = Icon(action$icon)
  )
}

simpleBottomNavigation <- function(value) {
  Box(
    sx = list(width = 500, maxWidth = "100%"),
    BottomNavigation(
      showLabels = TRUE,
      value = value,
      lapply(actions, navigationAction, route = "simple-bottom-navigation")
    )
  )
}

navigationRoutes <- function(route, actions, render) {
  c(
    lapply(actions, function(action) Route(path = paste0(route, "/", action$value), element = render(action$value))),
    list(Route(path = "*", element = render(actions[[1]]$value)))
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(navigationRoutes("simple-bottom-navigation", actions, simpleBottomNavigation)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import BottomNavigation from '@mui/material/BottomNavigation';
import BottomNavigationAction from '@mui/material/BottomNavigationAction';
import RestoreIcon from '@mui/icons-material/Restore';
import FavoriteIcon from '@mui/icons-material/Favorite';
import LocationOnIcon from '@mui/icons-material/LocationOn';

export default function SimpleBottomNavigation() {
  const [value, setValue] = React.useState(0);

  return (
    <Box sx={{ width: 500 }}>
      <BottomNavigation
        showLabels
        value={value}
        onChange={(event, newValue) => {
          setValue(newValue);
        }}
      >
        <BottomNavigationAction label="Recents" icon={<RestoreIcon />} />
        <BottomNavigationAction label="Favorites" icon={<FavoriteIcon />} />
        <BottomNavigationAction label="Nearby" icon={<LocationOnIcon />} />
      </BottomNavigation>
    </Box>
  );
}
```
</details>

## Bottom navigation with no label

If there are **four** or **five** actions, display inactive views as icons only.

```{r}
labelActions <- c(actions, list(list(value = "folder", label = "Folder", icon = "folder")))

labelBottomNavigation <- function(value) {
  BottomNavigation(
    sx = list(width = 500, maxWidth = "100%"),
    value = value,
    lapply(labelActions, navigationAction, route = "label-bottom-navigation")
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(navigationRoutes("label-bottom-navigation", labelActions, labelBottomNavigation)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import BottomNavigation from '@mui/material/BottomNavigation';
import BottomNavigationAction from '@mui/material/BottomNavigationAction';
import FolderIcon from '@mui/icons-material/Folder';
import RestoreIcon from '@mui/icons-material/Restore';
import FavoriteIcon from '@mui/icons-material/Favorite';
import LocationOnIcon from '@mui/icons-material/LocationOn';

export default function LabelBottomNavigation() {
  const [value, setValue] = React.useState('recents');

  const handleChange = (event, newValue) => {
    setValue(newValue);
  };

  return (
    <BottomNavigation sx={{ width: 500 }} value={value} onChange={handleChange}>
      <BottomNavigationAction
        label="Recents"
        value="recents"
        icon={<RestoreIcon />}
      />
      <BottomNavigationAction
        label="Favorites"
        value="favorites"
        icon={<FavoriteIcon />}
      />
      <BottomNavigationAction
        label="Nearby"
        value="nearby"
        icon={<LocationOnIcon />}
      />
      <BottomNavigationAction label="Folder" value="folder" icon={<FolderIcon />} />
    </BottomNavigation>
  );
}
```
</details>

## Fixed positioning

This demo keeps bottom navigation fixed to the bottom, no matter the amount of content on-screen.
It is rendered in a framed box: its `transform` makes the fixed-position bar
stay at the bottom of the frame, while the inner box scrolls.

```{r}
messages <- list(
  list(primary = "Brunch this week?", secondary = "I'll be in the neighbourhood this week. Let's grab a bite to eat", person = 5),
  list(primary = "Birthday Gift", secondary = "Do you have a suggestion for a good present for John on his work anniversary. I am really confused & would love your thoughts on it.", person = 1),
  list(primary = "Recipe to try", secondary = "I am try out this new BBQ recipe, I think this might be amazing", person = 2),
  list(primary = "Yes!", secondary = "I have the tickets to the ReactConf for this year.", person = 3),
  list(primary = "Doctor's Appointment", secondary = "My appointment for the doctor was rescheduled for next Saturday.", person = 4),
  list(primary = "Discussion", secondary = "Menus that are generated by the bottom app bar (such as a bottom navigation drawer or overflow menu) open as bottom sheets at a higher elevation than the bar.", person = 5),
  list(primary = "Summer BBQ", secondary = "Who wants to have a cookout this weekend? I just got some furniture for my backyard and would love to fire up the grill.", person = 1)
)
fixedActions <- list(
  list(value = "recents", label = "Recents", icon = "restore"),
  list(value = "favorites", label = "Favorites", icon = "favorite"),
  list(value = "archive", label = "Archive", icon = "archive")
)

fixedBottomNavigation <- function(value) {
  set.seed(match(value, sapply(fixedActions, `[[`, "value"))) # new messages for each action
  Box(
    sx = list(position = "relative", height = 400, overflow = "hidden", transform = "translateZ(0)", border = 1, borderColor = "divider"),
    Box(
      sx = list(height = "100%", overflow = "auto", pb = 7),
      List(lapply(sample(messages, 20, replace = TRUE), function(m) {
        ListItemButton(
          ListItemAvatar(Avatar(alt = "Profile Picture", src = sprintf("https://mui.com/static/images/avatar/%s.jpg", m$person))),
          ListItemText(primary = m$primary, secondary = m$secondary)
        )
      }))
    ),
    Paper(
      sx = list(position = "fixed", bottom = 0, left = 0, right = 0),
      elevation = 3,
      BottomNavigation(showLabels = TRUE, value = value, lapply(fixedActions, navigationAction, route = "fixed-bottom-navigation"))
    )
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(navigationRoutes("fixed-bottom-navigation", fixedActions, fixedBottomNavigation)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import CssBaseline from '@mui/material/CssBaseline';
import BottomNavigation from '@mui/material/BottomNavigation';
import BottomNavigationAction from '@mui/material/BottomNavigationAction';
import RestoreIcon from '@mui/icons-material/Restore';
import FavoriteIcon from '@mui/icons-material/Favorite';
import ArchiveIcon from '@mui/icons-material/Archive';
import Paper from '@mui/material/Paper';
import List from '@mui/material/List';
import ListItemButton from '@mui/material/ListItemButton';
import ListItemAvatar from '@mui/material/ListItemAvatar';
import ListItemText from '@mui/material/ListItemText';
import Avatar from '@mui/material/Avatar';

function refreshMessages() {
  const getRandomInt = (max) => Math.floor(Math.random() * Math.floor(max));

  return Array.from(new Array(50)).map(
    () => messageExamples[getRandomInt(messageExamples.length)],
  );
}

export default function FixedBottomNavigation() {
  const [value, setValue] = React.useState(0);
  const ref = React.useRef(null);
  const [messages, setMessages] = React.useState(() => refreshMessages());

  React.useEffect(() => {
    ref.current.ownerDocument.body.scrollTop = 0;
    setMessages(refreshMessages());
  }, [value, setMessages]);

  return (
    <Box sx={{ pb: 7 }} ref={ref}>
      <CssBaseline />
      <List>
        {messages.map(({ primary, secondary, person }, index) => (
          <ListItemButton key={index + person}>
            <ListItemAvatar>
              <Avatar alt="Profile Picture" src={person} />
            </ListItemAvatar>
            <ListItemText primary={primary} secondary={secondary} />
          </ListItemButton>
        ))}
      </List>
      <Paper sx={{ position: 'fixed', bottom: 0, left: 0, right: 0 }} elevation={3}>
        <BottomNavigation
          showLabels
          value={value}
          onChange={(event, newValue) => {
            setValue(newValue);
          }}
        >
          <BottomNavigationAction label="Recents" icon={<RestoreIcon />} />
          <BottomNavigationAction label="Favorites" icon={<FavoriteIcon />} />
          <BottomNavigationAction label="Archive" icon={<ArchiveIcon />} />
        </BottomNavigation>
      </Paper>
    </Box>
  );
}

const messageExamples = [
  {
    primary: 'Brunch this week?',
    secondary: "I'll be in the neighbourhood this week. Let's grab a bite to eat",
    person: '/static/images/avatar/5.jpg',
  },
  {
    primary: 'Birthday Gift',
    secondary: `Do you have a suggestion for a good present for John on his work
      anniversary. I am really confused & would love your thoughts on it.`,
    person: '/static/images/avatar/1.jpg',
  },
  {
    primary: 'Recipe to try',
    secondary: 'I am try out this new BBQ recipe, I think this might be amazing',
    person: '/static/images/avatar/2.jpg',
  },
  {
    primary: 'Yes!',
    secondary: 'I have the tickets to the ReactConf for this year.',
    person: '/static/images/avatar/3.jpg',
  },
  {
    primary: "Doctor's Appointment",
    secondary: 'My appointment for the doctor was rescheduled for next Saturday.',
    person: '/static/images/avatar/4.jpg',
  },
  {
    primary: 'Discussion',
    secondary: `Menus that are generated by the bottom app bar (such as a bottom
      navigation drawer or overflow menu) open as bottom sheets at a higher elevation
      than the bar.`,
    person: '/static/images/avatar/5.jpg',
  },
  {
    primary: 'Summer BBQ',
    secondary: `Who wants to have a cookout this weekend? I just got some furniture
      for my backyard and would love to fire up the grill.`,
    person: '/static/images/avatar/1.jpg',
  },
];
```
</details>

## Third-party routing library

One frequent use case is to perform navigation on the client only, without an HTTP round-trip to the server.
The `BottomNavigationAction` component provides the `component` prop to handle this use case.
All the demos on this page use it with [reactRouter](https://felixluginbuhl.com/reactRouter/),
and [Using a router](routing.html) shows complete multi-page apps.

## Bottom navigation in Shiny apps

`BottomNavigation.shinyInput()` sends the `value` of the selected action to
`input[[inputId]]`. Give each `BottomNavigationAction()` a `value`, and pass the
initial one as `value`:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Box(
    sx = list(width = 500, p = 2),
    reactOutput("page"),
    BottomNavigation.shinyInput(
      inputId = "nav",
      value = "recents",
      showLabels = TRUE,
      BottomNavigationAction(label = "Recents", value = "recents", icon = Icon("restore")),
      BottomNavigationAction(label = "Favorites", value = "favorites", icon = Icon("favorite")),
      BottomNavigationAction(label = "Nearby", value = "nearby", icon = Icon("location_on"))
    )
  )
)

server <- function(input, output, session) {
  output$page <- renderReact({
    Typography(variant = "h5", sx = list(p = 2), paste("Page:", input$nav))
  })
}

shinyApp(ui, server)
```

Run the bundled version with `muiMaterialExample("BottomNavigation")`.
