---
title: "Box"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Box}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-box/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
```

# Box

The Box component is a generic, theme-aware container with access to CSS utilities from MUI System.

## Introduction

The Box component is a generic container for grouping other components.
It's a fundamental building block when working with Material UI. You can think of it as a `<div>` with extra built-in features, like access to your app's theme and the [`sx` prop](#the-sx-prop-in-r).

### Usage

The Box component differs from other containers available in Material UI in that its usage is intended to be multipurpose and open-ended, just like a `<div>`.
Components like [Container](container.html), [Stack](stack.html) and [Paper](paper.html), by contrast, feature usage-specific props that make them ideal for certain use cases.

## Basics

The Box component renders as a `<div>` by default, but you can swap in any other valid HTML tag or React component using the `component` prop.
The demo below replaces the `<div>` with a `<section>` element:

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(component = "section", sx = list(p = 2, border = "1px dashed grey"), "This Box renders as an HTML section element.")
)
```

<details><summary>JS code</summary>
```jsx
import Box from '@mui/material/Box';

export default function BoxBasic() {
  return (
    <Box component="section" sx={{ p: 2, border: '1px dashed grey' }}>
      This Box renders as an HTML section element.
    </Box>
  );
}
```
</details>

## Customization

Use the [`sx` prop](https://mui.com/system/getting-started/the-sx-prop/) to quickly customize any Box instance using a superset of CSS that has access to all the style functions and theme-aware properties exposed in the MUI System package.
The demo below shows how to apply colors from the theme using this prop:

```{r}
muiMaterialPage(
  CssBaseline(),
  ThemeProvider(
    theme = list(palette = list(primary = list(main = "#007FFF", dark = "#0066CC"))),
    Box(
      sx = list(
        width = 100,
        height = 100,
        borderRadius = 1,
        bgcolor = "primary.main",
        "&:hover" = list(bgcolor = "primary.dark")
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Box from '@mui/material/Box';
import { ThemeProvider } from '@mui/material/styles';

export default function BoxSx() {
  return (
    <ThemeProvider
      theme={{
        palette: {
          primary: {
            main: '#007FFF',
            dark: '#0066CC',
          },
        },
      }}
    >
      <Box
        sx={{
          width: 100,
          height: 100,
          borderRadius: 1,
          bgcolor: 'primary.main',
          '&:hover': {
            bgcolor: 'primary.dark',
          },
        }}
      />
    </ThemeProvider>
  );
}
```
</details>

## The `sx` prop in R

Every muiMaterial component accepts `sx`. The JavaScript object becomes a named
R `list()`, following a few rules:

| JavaScript | R |
|---|---|
| `sx={{ p: 2, bgcolor: 'primary.main' }}` | `sx = list(p = 2, bgcolor = "primary.main")` |
| `'&:hover': { color: 'red' }` | `"&:hover" = list(color = "red")` |
| `'& .MuiChip-label': { ... }` | `"& .MuiChip-label" = list(...)` |
| `width: { xs: '100%', md: 400 }` | `width = list(xs = "100%", md = 400)` |
| `sx={(theme) => ({ ... })}` | `sx = JS("(theme) => ({ ... })")` |

- **Spacing** values (`m`, `p`, `mx`, `pt`, `gap`, ...) are multiples of the
  theme spacing (8px by default): `p = 2` is `16px`. Strings are used as is:
  `p = "3px"`.
- **Colors** can be theme paths such as `"primary.main"`, `"text.secondary"` or
  `"grey.300"`, or any CSS color.
- **Breakpoints** (`xs`, `sm`, `md`, `lg`, `xl`) make any value responsive:
  `display = list(xs = "none", md = "flex")` hides an element on small screens.
- Keys that are not valid R names (selectors, `z-index` style names) must be
  quoted: `"&:hover"`, `"@media all"`.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(
      display = "flex",
      flexDirection = list(xs = "column", sm = "row"),
      gap = 2,
      p = 2,
      bgcolor = "grey.100",
      borderRadius = 2
    ),
    lapply(c("primary", "secondary", "success"), function(color) {
      Box(
        sx = list(
          flex = 1,
          p = 2,
          color = "common.white",
          bgcolor = paste0(color, ".main"),
          borderRadius = 1,
          textAlign = "center",
          transition = "transform 150ms",
          "&:hover" = list(transform = "scale(1.05)")
        ),
        color
      )
    })
  )
)
```

See the [MUI System documentation](https://mui.com/system/getting-started/the-sx-prop/)
for all available properties.

## Anatomy

The Box component is composed of a single root `<div>` element:

```html
<div className="MuiBox-root">
  <!-- contents of the Box -->
</div>
```
