---
title: "Breadcrumbs"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Breadcrumbs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-breadcrumbs/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# Breadcrumbs

A breadcrumbs is a list of links that help visualize a page's location within a site's hierarchical structure, it allows navigation up to any of the ancestors.

As in the MUI documentation, the links of the demos do not navigate: a
`preventDefault()` handler on the wrapper logs the click to the browser console
instead.

```{r}
handleClick <- JS("(event) => { event.preventDefault(); console.info('You clicked a breadcrumb.'); }")
```

## Basic breadcrumbs

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    role = "presentation",
    onClick = handleClick,
    Breadcrumbs(
      `aria-label` = "breadcrumb",
      Link(underline = "hover", color = "inherit", href = "/", "MUI"),
      Link(underline = "hover", color = "inherit", href = "/material-ui/getting-started/installation/", "Core"),
      Typography(sx = list(color = "text.primary"), "Breadcrumbs")
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Typography from '@mui/material/Typography';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Link from '@mui/material/Link';

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function BasicBreadcrumbs() {
  return (
    <div role="presentation" onClick={handleClick}>
      <Breadcrumbs aria-label="breadcrumb">
        <Link underline="hover" color="inherit" href="/">
          MUI
        </Link>
        <Link
          underline="hover"
          color="inherit"
          href="/material-ui/getting-started/installation/"
        >
          Core
        </Link>
        <Typography sx={{ color: 'text.primary' }}>Breadcrumbs</Typography>
      </Breadcrumbs>
    </div>
  );
}
```
</details>

## Active last breadcrumb

Keep the last breadcrumb interactive.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    role = "presentation",
    onClick = handleClick,
    Breadcrumbs(
      `aria-label` = "breadcrumb",
      Link(underline = "hover", color = "inherit", href = "/", "MUI"),
      Link(underline = "hover", color = "inherit", href = "/material-ui/getting-started/installation/", "Core"),
      Link(
        underline = "hover",
        href = "/material-ui/react-breadcrumbs/",
        `aria-current` = "page",
        sx = list(color = "text.primary"),
        "Breadcrumbs"
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Link from '@mui/material/Link';

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function ActiveLastBreadcrumb() {
  return (
    <div role="presentation" onClick={handleClick}>
      <Breadcrumbs aria-label="breadcrumb">
        <Link underline="hover" color="inherit" href="/">
          MUI
        </Link>
        <Link
          underline="hover"
          color="inherit"
          href="/material-ui/getting-started/installation/"
        >
          Core
        </Link>
        <Link
          underline="hover"
          href="/material-ui/react-breadcrumbs/"
          aria-current="page"
          sx={{
            color: 'text.primary',
          }}
        >
          Breadcrumbs
        </Link>
      </Breadcrumbs>
    </div>
  );
}
```
</details>

## Custom separator

In the following examples, we are using two string separators and an icon.

```{r}
breadcrumbs <- tagList(
  Link(underline = "hover", color = "inherit", href = "/", onClick = handleClick, "MUI"),
  Link(underline = "hover", color = "inherit", href = "/material-ui/getting-started/installation/", onClick = handleClick, "Core"),
  Typography(sx = list(color = "text.primary"), "Breadcrumb")
)

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Stack(
    spacing = 2,
    Breadcrumbs(separator = "›", `aria-label` = "breadcrumb", breadcrumbs),
    Breadcrumbs(separator = "-", `aria-label` = "breadcrumb", breadcrumbs),
    Breadcrumbs(separator = Icon("navigate_next", fontSize = "small"), `aria-label` = "breadcrumb", breadcrumbs)
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Typography from '@mui/material/Typography';
import Link from '@mui/material/Link';
import Stack from '@mui/material/Stack';
import NavigateNextIcon from '@mui/icons-material/NavigateNext';

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function CustomSeparator() {
  const breadcrumbs = [
    <Link underline="hover" key="1" color="inherit" href="/" onClick={handleClick}>
      MUI
    </Link>,
    <Link
      underline="hover"
      key="2"
      color="inherit"
      href="/material-ui/getting-started/installation/"
      onClick={handleClick}
    >
      Core
    </Link>,
    <Typography key="3" sx={{ color: 'text.primary' }}>
      Breadcrumb
    </Typography>,
  ];

  return (
    <Stack spacing={2}>
      <Breadcrumbs separator="›" aria-label="breadcrumb">
        {breadcrumbs}
      </Breadcrumbs>
      <Breadcrumbs separator="-" aria-label="breadcrumb">
        {breadcrumbs}
      </Breadcrumbs>
      <Breadcrumbs
        separator={<NavigateNextIcon fontSize="small" />}
        aria-label="breadcrumb"
      >
        {breadcrumbs}
      </Breadcrumbs>
    </Stack>
  );
}
```
</details>

## Breadcrumbs with icons

```{r}
iconLink <- function(icon, label, href) {
  Link(
    underline = "hover",
    sx = list(display = "flex", alignItems = "center"),
    color = "inherit",
    href = href,
    Icon(icon, sx = list(mr = 0.5), fontSize = "inherit"),
    label
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Box(
    role = "presentation",
    onClick = handleClick,
    Breadcrumbs(
      `aria-label` = "breadcrumb",
      iconLink("home", "MUI", "/"),
      iconLink("whatshot", "Core", "/material-ui/getting-started/installation/"),
      Typography(
        sx = list(color = "text.primary", display = "flex", alignItems = "center"),
        Icon("grain", sx = list(mr = 0.5), fontSize = "inherit"),
        "Breadcrumb"
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Typography from '@mui/material/Typography';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Link from '@mui/material/Link';
import HomeIcon from '@mui/icons-material/Home';
import WhatshotIcon from '@mui/icons-material/Whatshot';
import GrainIcon from '@mui/icons-material/Grain';

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function IconBreadcrumbs() {
  return (
    <div role="presentation" onClick={handleClick}>
      <Breadcrumbs aria-label="breadcrumb">
        <Link
          underline="hover"
          sx={{ display: 'flex', alignItems: 'center' }}
          color="inherit"
          href="/"
        >
          <HomeIcon sx={{ mr: 0.5 }} fontSize="inherit" />
          MUI
        </Link>
        <Link
          underline="hover"
          sx={{ display: 'flex', alignItems: 'center' }}
          color="inherit"
          href="/material-ui/getting-started/installation/"
        >
          <WhatshotIcon sx={{ mr: 0.5 }} fontSize="inherit" />
          Core
        </Link>
        <Typography
          sx={{ color: 'text.primary', display: 'flex', alignItems: 'center' }}
        >
          <GrainIcon sx={{ mr: 0.5 }} fontSize="inherit" />
          Breadcrumb
        </Typography>
      </Breadcrumbs>
    </div>
  );
}
```
</details>

## Collapsed breadcrumbs

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    role = "presentation",
    onClick = handleClick,
    Breadcrumbs(
      maxItems = 2,
      `aria-label` = "breadcrumb",
      lapply(c("Home", "Catalog", "Accessories", "New Collection"), function(label) {
        Link(underline = "hover", color = "inherit", href = "#", label)
      }),
      Typography(sx = list(color = "text.primary"), "Belts")
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Typography from '@mui/material/Typography';
import Link from '@mui/material/Link';

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function CollapsedBreadcrumbs() {
  return (
    <div role="presentation" onClick={handleClick}>
      <Breadcrumbs maxItems={2} aria-label="breadcrumb">
        <Link underline="hover" color="inherit" href="#">
          Home
        </Link>
        <Link underline="hover" color="inherit" href="#">
          Catalog
        </Link>
        <Link underline="hover" color="inherit" href="#">
          Accessories
        </Link>
        <Link underline="hover" color="inherit" href="#">
          New Collection
        </Link>
        <Typography sx={{ color: 'text.primary' }}>Belts</Typography>
      </Breadcrumbs>
    </div>
  );
}
```
</details>

## Condensed with menu

As an alternative, consider adding a [Menu](menu.html) component to display the condensed links in a dropdown list.
`Menu.triggerId()` opens the menu next to the "more" button:

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Menu.triggerId(
    triggerId = "with-menu-demo-breadcrumbs",
    `aria-labelledby` = "with-menu-demo-breadcrumbs",
    MenuItem("Breadcrumb 2"),
    MenuItem("Breadcrumb 3"),
    MenuItem("Breadcrumb 4")
  ),
  Breadcrumbs(
    `aria-label` = "breadcrumbs",
    Link(color = "primary", href = "#condensed-with-menu", "Breadcrumb 1"),
    IconButton(id = "with-menu-demo-breadcrumbs", color = "primary", size = "small", Icon("more_horiz")),
    Link(color = "primary", href = "#condensed-with-menu", "Breadcrumb 5"),
    Link(color = "primary", href = "#condensed-with-menu", "Breadcrumb 6")
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Link from '@mui/material/Link';
import Menu from '@mui/material/Menu';
import MenuItem from '@mui/material/MenuItem';
import IconButton from '@mui/material/IconButton';
import MoreHorizIcon from '@mui/icons-material/MoreHoriz';

export default function CondensedWithMenu() {
  const [anchorEl, setAnchorEl] = React.useState(null);
  const open = Boolean(anchorEl);

  const handleClick = (event) => {
    if (event) {
      setAnchorEl(event.currentTarget);
    }
  };

  const handleClose = () => {
    setAnchorEl(null);
  };

  return (
    <React.Fragment>
      <Menu
        anchorEl={anchorEl}
        open={open}
        onClose={handleClose}
        aria-labelledby="with-menu-demo-breadcrumbs"
      >
        <MenuItem onClick={handleClose}>Breadcrumb 2</MenuItem>
        <MenuItem onClick={handleClose}>Breadcrumb 3</MenuItem>
        <MenuItem onClick={handleClose}>Breadcrumb 4</MenuItem>
      </Menu>
      <Breadcrumbs aria-label="breadcrumbs">
        <Link color="primary" href="#condensed-with-menu">
          Breadcrumb 1
        </Link>
        <IconButton color="primary" size="small" onClick={handleClick}>
          <MoreHorizIcon />
        </IconButton>
        <Link color="primary" href="#condensed-with-menu">
          Breadcrumb 5
        </Link>
        <Link color="primary" href="#condensed-with-menu">
          Breadcrumb 6
        </Link>
      </Breadcrumbs>
    </React.Fragment>
  );
}
```
</details>

## Customization

Here is an example of customizing the component.
You can learn more about this in the [overrides documentation page](https://mui.com/material-ui/customization/how-to-customize/).
The `styled(Chip)` of the original demo becomes an R function returning a
`Chip()` styled with `sx`:

```{r}
StyledBreadcrumb <- function(...) {
  Chip(
    sx = list(
      backgroundColor = "grey.100",
      height = 24,
      color = "text.primary",
      fontWeight = "fontWeightRegular",
      "&:hover, &:focus" = list(backgroundColor = "#e5e5e5"), # emphasize(grey[100], 0.06)
      "&:active" = list(boxShadow = 1, backgroundColor = "#d6d6d6") # emphasize(grey[100], 0.12)
    ),
    ...
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Box(
    role = "presentation",
    onClick = handleClick,
    Breadcrumbs(
      `aria-label` = "breadcrumb",
      StyledBreadcrumb(component = "a", href = "#", label = "Home", icon = Icon("home", fontSize = "small")),
      StyledBreadcrumb(component = "a", href = "#", label = "Catalog"),
      StyledBreadcrumb(label = "Accessories", deleteIcon = Icon("expand_more"), onDelete = handleClick)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { emphasize, styled } from '@mui/material/styles';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import Chip from '@mui/material/Chip';
import HomeIcon from '@mui/icons-material/Home';
import ExpandMoreIcon from '@mui/icons-material/ExpandMore';

const StyledBreadcrumb = styled(Chip)(({ theme }) => {
  return {
    backgroundColor: theme.palette.grey[100],
    height: theme.spacing(3),
    color: (theme.vars || theme).palette.text.primary,
    fontWeight: theme.typography.fontWeightRegular,
    '&:hover, &:focus': {
      backgroundColor: emphasize(theme.palette.grey[100], 0.06),
      ...theme.applyStyles('dark', {
        backgroundColor: emphasize(theme.palette.grey[800], 0.06),
      }),
    },
    '&:active': {
      boxShadow: theme.shadows[1],
      backgroundColor: emphasize(theme.palette.grey[100], 0.12),
      ...theme.applyStyles('dark', {
        backgroundColor: emphasize(theme.palette.grey[800], 0.12),
      }),
    },
    ...theme.applyStyles('dark', {
      backgroundColor: theme.palette.grey[800],
    }),
  };
}); // TypeScript only: need a type cast here because https://github.com/Microsoft/TypeScript/issues/26591

function handleClick(event) {
  event.preventDefault();
  console.info('You clicked a breadcrumb.');
}

export default function CustomizedBreadcrumbs() {
  return (
    <div role="presentation" onClick={handleClick}>
      <Breadcrumbs aria-label="breadcrumb">
        <StyledBreadcrumb
          component="a"
          href="#"
          label="Home"
          icon={<HomeIcon fontSize="small" />}
        />
        <StyledBreadcrumb component="a" href="#" label="Catalog" />
        <StyledBreadcrumb
          label="Accessories"
          deleteIcon={<ExpandMoreIcon />}
          onDelete={handleClick}
        />
      </Breadcrumbs>
    </div>
  );
}
```
</details>

## Integration with react-router

With [reactRouter](https://felixluginbuhl.com/reactRouter/), the breadcrumbs of
each route are built in R from its path. The functions of reactRouter are
called with the `reactRouter::` prefix here, because reactRouter also exports a
`Link()` function, which masks `muiMaterial::Link()` once reactRouter is
attached.

```{r}
breadcrumbNameMap <- c(
  "/inbox" = "Inbox",
  "/inbox/important" = "Important",
  "/trash" = "Trash",
  "/spam" = "Spam",
  "/drafts" = "Drafts"
)

routerBreadcrumbs <- function(path) {
  segments <- strsplit(sub("^/", "", path), "/")[[1]]
  parents <- vapply(seq_along(segments), function(i) paste0("/", paste(segments[seq_len(i)], collapse = "/")), "")
  Box(
    sx = list(display = "flex", flexDirection = "column", width = 360),
    Breadcrumbs(
      `aria-label` = "breadcrumb",
      Link(underline = "hover", color = "inherit", href = "#/", "Home"),
      lapply(seq_along(parents), function(i) {
        if (i == length(parents)) {
          Typography(sx = list(color = "text.primary"), breadcrumbNameMap[[parents[i]]])
        } else {
          Link(underline = "hover", color = "inherit", href = paste0("#", parents[i]), breadcrumbNameMap[[parents[i]]])
        }
      })
    ),
    Box(
      sx = list(bgcolor = "background.paper", mt = 1),
      component = "nav",
      `aria-label` = "mailbox folders",
      List(lapply(names(breadcrumbNameMap), function(to) {
        ListItemButton(
          component = "a",
          href = paste0("#", to),
          selected = identical(to, path),
          sx = list(pl = if (grepl("/.+/", to)) 4 else 2),
          ListItemText(primary = breadcrumbNameMap[[to]])
        )
      }))
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  reactRouter::RouterProvider(
    router = reactRouter::createHashRouter(
      lapply(names(breadcrumbNameMap), function(path) {
        reactRouter::Route(path = path, element = routerBreadcrumbs(path))
      }),
      reactRouter::Route(path = "*", element = routerBreadcrumbs("/inbox"))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import PropTypes from 'prop-types';
import Box from '@mui/material/Box';
import List from '@mui/material/List';
import Link from '@mui/material/Link';

import ListItemButton from '@mui/material/ListItemButton';
import Collapse from '@mui/material/Collapse';
import ListItemText from '@mui/material/ListItemText';
import Typography from '@mui/material/Typography';
import ExpandLess from '@mui/icons-material/ExpandLess';
import ExpandMore from '@mui/icons-material/ExpandMore';
import Breadcrumbs from '@mui/material/Breadcrumbs';
import {
  Link as RouterLink,
  Route,
  Routes,
  MemoryRouter,
  useLocation,
} from 'react-router';

const breadcrumbNameMap = {
  '/inbox': 'Inbox',
  '/inbox/important': 'Important',
  '/trash': 'Trash',
  '/spam': 'Spam',
  '/drafts': 'Drafts',
};

function ListItemLink(props) {
  const { to, open, ...other } = props;
  const primary = breadcrumbNameMap[to];

  let icon = null;
  if (open != null) {
    icon = open ? <ExpandLess /> : <ExpandMore />;
  }

  return (
    <li>
      <ListItemButton component={RouterLink} to={to} {...other}>
        <ListItemText primary={primary} />
        {icon}
      </ListItemButton>
    </li>
  );
}

ListItemLink.propTypes = {
  open: PropTypes.bool,
  to: PropTypes.string.isRequired,
};

function LinkRouter(props) {
  return <Link {...props} component={RouterLink} />;
}

function Page() {
  const location = useLocation();
  const pathnames = location.pathname.split('/').filter((x) => x);

  return (
    <Breadcrumbs aria-label="breadcrumb">
      <LinkRouter underline="hover" color="inherit" to="/">
        Home
      </LinkRouter>
      {pathnames.map((value, index) => {
        const last = index === pathnames.length - 1;
        const to = `/${pathnames.slice(0, index + 1).join('/')}`;

        return last ? (
          <Typography key={to} sx={{ color: 'text.primary' }}>
            {breadcrumbNameMap[to]}
          </Typography>
        ) : (
          <LinkRouter underline="hover" color="inherit" to={to} key={to}>
            {breadcrumbNameMap[to]}
          </LinkRouter>
        );
      })}
    </Breadcrumbs>
  );
}

export default function RouterBreadcrumbs() {
  const [open, setOpen] = React.useState(true);

  const handleClick = () => {
    setOpen((prevOpen) => !prevOpen);
  };

  return (
    <MemoryRouter initialEntries={['/inbox']} initialIndex={0}>
      <Box sx={{ display: 'flex', flexDirection: 'column', width: 360 }}>
        <Routes>
          <Route path="*" element={<Page />} />
        </Routes>
        <Box
          sx={{ bgcolor: 'background.paper', mt: 1 }}
          component="nav"
          aria-label="mailbox folders"
        >
          <List>
            <ListItemLink to="/inbox" open={open} onClick={handleClick} />
            <Collapse component="li" in={open} timeout="auto" unmountOnExit>
              <List disablePadding>
                <ListItemLink sx={{ pl: 4 }} to="/inbox/important" />
              </List>
            </Collapse>
            <ListItemLink to="/trash" />
            <ListItemLink to="/spam" />
          </List>
        </Box>
      </Box>
    </MemoryRouter>
  );
}
```
</details>

## Accessibility

(WAI-ARIA: <https://www.w3.org/WAI/ARIA/apg/patterns/breadcrumb/>)

Be sure to add a `aria-label` description on the `Breadcrumbs` component.

The accessibility of this component relies on:

- The set of links is structured using an ordered list (`<ol>` element).
- To prevent screen reader announcement of the visual separators between links, they are hidden with `aria-hidden`.
- A nav element labeled with `aria-label` identifies the structure as a breadcrumb trail and makes it a navigation landmark so that it is easy to locate.
