---
title: "Button"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Button}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-button/).</div>

```{r setup}
library(shinyMaterialUI)
library(shiny)
```

# Button

Buttons allow users to take actions, and make choices, with a single tap.

Buttons communicate actions that users can take. They are typically placed throughout your UI, in places like:

- Modal windows
- Forms
- Cards
- Toolbars

## Basic button

The Button comes with three variants: text (default), contained, and outlined.

```{r}
CssBaseline(
  Stack(
    spacing = 2, 
    direction = "row",
    Button(variant = "text", "Text"),
    Button(variant = "contained", "Contained"),
    Button(variant = "outlined", "Outlined")
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Stack from '@mui/material/Stack';
import Button from '@mui/material/Button';

export default function BasicButtons() {
  return (
    <Stack spacing={2} direction="row">
      <Button variant="text">Text</Button>
      <Button variant="contained">Contained</Button>
      <Button variant="outlined">Outlined</Button>
    </Stack>
  );
}
```
</details>

## Text button

Text buttons are typically used for less-pronounced actions, including those located: in dialogs, in cards. In cards, text buttons help maintain an emphasis on card content.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 2,
    Button("Primary"),
    Button(disabled = TRUE, "Disabled"),
    Button(href = "#text-buttons", "Link")
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Stack from '@mui/material/Stack';

export default function TextButtons() {
  return (
    <Stack direction="row" spacing={2}>
      <Button>Primary</Button>
      <Button disabled>Disabled</Button>
      <Button href="#text-buttons">Link</Button>
    </Stack>
  );
}
```
</details>

## Contained button

Contained buttons are high-emphasis, distinguished by their use of elevation and fill. They contain actions that are primary to your app.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 2,
    Button(variant = "contained", "Contained"),
    Button(variant = "contained", disabled = TRUE, "Disabled"),
    Button(variant = "contained", href = "#contained-buttons", "Link")
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Stack from '@mui/material/Stack';

export default function ContainedButtons() {
  return (
    <Stack direction="row" spacing={2}>
      <Button variant="contained">Contained</Button>
      <Button variant="contained" disabled>
        Disabled
      </Button>
      <Button variant="contained" href="#contained-buttons">
        Link
      </Button>
    </Stack>
  );
}
```
</details>

You can remove the elevation with the disableElevation prop.

```{r}
CssBaseline(
  Button(variant = "contained", disableElevation = TRUE, "Disable elevation")
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';

export default function DisableElevation() {
  return (
    <Button variant="contained" disableElevation>
      Disable elevation
    </Button>
  );
}
```
</details>

## Outlined button

Outlined buttons are medium-emphasis buttons. They contain actions that are important but aren't the primary action in an app.

Outlined buttons are also a lower emphasis alternative to contained buttons, or a higher emphasis alternative to text buttons.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 2,
    Button(variant = "outlined", "Primary"),
    Button(variant = "outlined", disabled = TRUE, "Disabled"),
    Button(variant = "outlined", href = "#outlined-buttons", "Link")
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Stack from '@mui/material/Stack';

export default function OutlinedButtons() {
  return (
    <Stack direction="row" spacing={2}>
      <Button variant="outlined">Primary</Button>
      <Button variant="outlined" disabled>
        Disabled
      </Button>
      <Button variant="outlined" href="#outlined-buttons">
        Link
      </Button>
    </Stack>
  );
}
```
</details>

## Handling clicks

All components accept an onClick handler that is applied to the root DOM element.

```{r eval=FALSE}
ui <- shinyMaterialUIPage(
  Button.shinyInput(
    inputId = "clickButton",
    "Click me"
  )
)

server <- function(input, output, session) {
  observeEvent(input$clickButton, {
    showModal(modalDialog(
      title = "Clicked",
      "Button was clicked"
    ))
  })
}

shinyApp(ui, server)
```


<details><summary>JS code</summary>
```jsx
<Button
  onClick={() => {
    alert('clicked');
  }}
>
  Click me
</Button>
```
</details>

Note that the documentation avoids mentioning native props (there are a lot) in the API section of the components.

## Color

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 2,
    Button(color = "secondary", "Secondary"),
    Button(variant = "contained", color = "success", "Success"),
    Button(variant = "outlined", color = "error", "Error")
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Stack from '@mui/material/Stack';
import Button from '@mui/material/Button';

export default function ColorButtons() {
  return (
    <Stack direction="row" spacing={2}>
      <Button color="secondary">Secondary</Button>
      <Button variant="contained" color="success">
        Success
      </Button>
      <Button variant="outlined" color="error">
        Error
      </Button>
    </Stack>
  );
}
```
</details>

In addition to using the default button colors, you can add custom ones, or disable any you don't need. See the Adding new colors examples for more info.

## Sizes

For larger or smaller buttons, use the size prop.

```{r}
CssBaseline(
  Box(
    sx = list('& button' = list(m = 1)),
    div(
      Button(size = "small", "Small"),
      Button(size = "medium", "Medium"),
      Button(size = "large", "Large")
    ),
    div(
      Button(variant = "outlined", size = "small", "Small"),
      Button(variant = "outlined", size = "medium", "Medium"),
      Button(variant = "outlined", size = "large", "Large")
    ),
    div(
      Button(variant = "contained", size = "small", "Small"),
      Button(variant = "contained", size = "medium", "Medium"),
      Button(variant = "contained", size = "large", "Large")
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Button from '@mui/material/Button';

export default function ButtonSizes() {
  return (
    <Box sx={{ '& button': { m: 1 } }}>
      <div>
        <Button size="small">Small</Button>
        <Button size="medium">Medium</Button>
        <Button size="large">Large</Button>
      </div>
      <div>
        <Button variant="outlined" size="small">
          Small
        </Button>
        <Button variant="outlined" size="medium">
          Medium
        </Button>
        <Button variant="outlined" size="large">
          Large
        </Button>
      </div>
      <div>
        <Button variant="contained" size="small">
          Small
        </Button>
        <Button variant="contained" size="medium">
          Medium
        </Button>
        <Button variant="contained" size="large">
          Large
        </Button>
      </div>
    </Box>
  );
}
```
</details>

## Buttons with icons and label

Sometimes you might want to have icons for certain buttons to enhance the UX of the application as we recognize logos more easily than plain text. For example, if you have a delete button you can label it with a dustbin icon.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 2,
    Button(
      variant = "outlined", 
      startIcon = shiny::icon("trash"),
      "Delete"
    ),
    Button(
      variant = "contained", 
      endIcon = shiny::icon("paper-plane"),
      "Send"
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import DeleteIcon from '@mui/icons-material/Delete';
import SendIcon from '@mui/icons-material/Send';
import Stack from '@mui/material/Stack';

export default function IconLabelButtons() {
  return (
    <Stack direction="row" spacing={2}>
      <Button variant="outlined" startIcon={<DeleteIcon />}>
        Delete
      </Button>
      <Button variant="contained" endIcon={<SendIcon />}>
        Send
      </Button>
    </Stack>
  );
}
```
</details>

## Icon button

Icon buttons are commonly found in app bars and toolbars.

Icons are also appropriate for toggle buttons that allow a single choice to be selected or deselected, such as adding or removing a star to an item.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 1,
    IconButton(
      `aria-label` = "delete",
      shiny::icon("trash")
    ),
    IconButton(
      `aria-label` = "delete", 
      disabled = TRUE, 
      color = "primary",
      shiny::icon("trash")
    ),
    IconButton(
      color = "secondary", 
      `aria-label` = "add an alarm",
      shiny::icon("clock")
    ),
    IconButton(
      color = "primary", 
      `aria-label` = "add to shopping cart",
      shiny::icon("shopping-cart")
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import IconButton from '@mui/material/IconButton';
import Stack from '@mui/material/Stack';
import DeleteIcon from '@mui/icons-material/Delete';
import AlarmIcon from '@mui/icons-material/Alarm';
import AddShoppingCartIcon from '@mui/icons-material/AddShoppingCart';

export default function IconButtons() {
  return (
    <Stack direction="row" spacing={1}>
      <IconButton aria-label="delete">
        <DeleteIcon />
      </IconButton>
      <IconButton aria-label="delete" disabled color="primary">
        <DeleteIcon />
      </IconButton>
      <IconButton color="secondary" aria-label="add an alarm">
        <AlarmIcon />
      </IconButton>
      <IconButton color="primary" aria-label="add to shopping cart">
        <AddShoppingCartIcon />
      </IconButton>
    </Stack>
  );
}
```
</details>

## Sizes

For larger or smaller icon buttons, use the size prop.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 1, 
    sx = list(alignItems = "center"),
    IconButton(
      `aria-label` = "delete", 
      size = "small",
      shiny::icon("trash")
    ),
    IconButton(
      `aria-label` = "delete", 
      size = "small",
      shiny::icon("trash", class = "fa-sm")
    ),
    IconButton(
      `aria-label` = "delete", 
      size = "large",
      shiny::icon("trash")
    ),
    IconButton(
      `aria-label` = "delete", 
      size = "large",
      shiny::icon("trash", class = "fa-lg")
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Stack from '@mui/material/Stack';
import IconButton from '@mui/material/IconButton';
import DeleteIcon from '@mui/icons-material/Delete';

export default function IconButtonSizes() {
  return (
    <Stack direction="row" spacing={1} sx={{ alignItems: 'center' }}>
      <IconButton aria-label="delete" size="small">
        <DeleteIcon fontSize="inherit" />
      </IconButton>
      <IconButton aria-label="delete" size="small">
        <DeleteIcon fontSize="small" />
      </IconButton>
      <IconButton aria-label="delete" size="large">
        <DeleteIcon />
      </IconButton>
      <IconButton aria-label="delete" size="large">
        <DeleteIcon fontSize="inherit" />
      </IconButton>
    </Stack>
  );
}
```
</details>

## Colors

Use color prop to apply theme color palette to component.

```{r}
CssBaseline(
  Stack(
    direction = "row", 
    spacing = 1,
    IconButton(
      `aria-label` = "fingerprint", 
      color = "secondary",
      shiny::icon("fingerprint")
    ),
    IconButton(
      `aria-label` = "fingerprint", 
      color = "success",
      shiny::icon("fingerprint")
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Stack from '@mui/material/Stack';
import IconButton from '@mui/material/IconButton';
import Fingerprint from '@mui/icons-material/Fingerprint';

export default function IconButtonColors() {
  return (
    <Stack direction="row" spacing={1}>
      <IconButton aria-label="fingerprint" color="secondary">
        <Fingerprint />
      </IconButton>
      <IconButton aria-label="fingerprint" color="success">
        <Fingerprint />
      </IconButton>
    </Stack>
  );
}
```
</details>

## Loading

Starting from v6.4.0, use loading prop to set icon buttons in a loading state and disable interactions.

```{r eval=FALSE}
ui <- shinyMaterialUIPage(
  Tooltip(
    title = "Click to see loading",
    IconButton.shinyInput(
      inputId = "loadingButton",
      shiny::icon("shopping-cart")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$loadingButton, {
    updateIconButton.shinyInput(
      inputId = "loadingButton",
      loading = TRUE
    )
    
    # Reset after 2 seconds
    Sys.sleep(2)
    updateIconButton.shinyInput(
      inputId = "loadingButton",
      loading = FALSE
    )
  })
}

shinyApp(ui, server)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Tooltip from '@mui/material/Tooltip';
import IconButton from '@mui/material/IconButton';
import ShoppingCartIcon from '@mui/icons-material/ShoppingCart';

export default function LoadingIconButton() {
  const [loading, setLoading] = React.useState(false);
  React.useEffect(() => {
    const timeout = setTimeout(() => {
      setLoading(false);
    }, 2000);
    return () => clearTimeout(timeout);
  });
  return (
    <Tooltip title="Click to see loading">
      <IconButton onClick={() => setLoading(true)} loading={loading}>
        <ShoppingCartIcon />
      </IconButton>
    </Tooltip>
  );
}
```
</details>

## Badge

You can use the Badge component to add a badge to an IconButton.

```{r}
CssBaseline(
  IconButton(
    shiny::icon("shopping-cart", class = "fa-sm"),
    Badge(
      badgeContent = 2, 
      color = "primary", 
      overlap = "circular",
      sx = list(
        "& .MuiBadge-badge" = list(
          top = "-12px",
          right = "-6px"
        )
      )
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import IconButton from '@mui/material/IconButton';
import Badge, { badgeClasses } from '@mui/material/Badge';
import ShoppingCartIcon from '@mui/icons-material/ShoppingCartOutlined';

const CartBadge = styled(Badge)`
  & .${badgeClasses.badge} {
    top: -12px;
    right: -6px;
  }
`;

export default function IconButtonWithBadge() {
  return (
    <IconButton>
      <ShoppingCartIcon fontSize="small" />
      <CartBadge badgeContent={2} color="primary" overlap="circular" />
    </IconButton>
  );
}
```
</details>

## File upload

To create a file upload button, turn the button into a label using component="label" and then create a visually-hidden input with type file.

```{r eval=FALSE}
ui <- shinyMaterialUIPage(
  Button(
    component = "label",
    role = NULL,
    variant = "contained",
    tabIndex = -1,
    startIcon = shiny::icon("cloud-upload-alt"),
    "Upload files",
    tags$input(
      type = "file",
      id = "fileUpload",
      style = "clip: rect(0 0 0 0); clipPath: inset(50%); height: 1px; overflow: hidden; position: absolute; bottom: 0; left: 0; whiteSpace: nowrap; width: 1px;",
      multiple = TRUE
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$fileUpload, {
    # Handle file upload
    req(input$fileUpload)
    print(input$fileUpload$name)
  })
}

shinyApp(ui, server)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import Button from '@mui/material/Button';
import CloudUploadIcon from '@mui/icons-material/CloudUpload';

const VisuallyHiddenInput = styled('input')({
  clip: 'rect(0 0 0 0)',
  clipPath: 'inset(50%)',
  height: 1,
  overflow: 'hidden',
  position: 'absolute',
  bottom: 0,
  left: 0,
  whiteSpace: 'nowrap',
  width: 1,
});

export default function InputFileUpload() {
  return (
    <Button
      component="label"
      role={undefined}
      variant="contained"
      tabIndex={-1}
      startIcon={<CloudUploadIcon />}
    >
      Upload files
      <VisuallyHiddenInput
        type="file"
        onChange={(event) => console.log(event.target.files)}
        multiple
      />
    </Button>
  );
}
```
</details>

## Loading

Starting from v6.4.0, use the loading prop to set buttons in a loading state and disable interactions.

```{r}
CssBaseline(
  Stack(
    spacing = 2,
    Stack(
      direction = "row", 
      spacing = 2,
      Button(loading = TRUE, variant = "outlined", "Submit"),
      Button(loading = TRUE, loadingIndicator = "Loading…", variant = "outlined", "Fetch data"),
      Button(
        loading = TRUE,
        loadingPosition = "start",
        startIcon = shiny::icon("save"),
        variant = "outlined",
        "Save"
      )
    ),
    Button(
      fullWidth = TRUE,
      loading = TRUE,
      loadingPosition = "start",
      startIcon = shiny::icon("save"),
      variant = "outlined",
      "Full width"
    ),
    Button(
      fullWidth = TRUE,
      loading = TRUE,
      loadingPosition = "end",
      endIcon = shiny::icon("save"),
      variant = "outlined",
      "Full width"
    ),
    Stack(
      direction = "row", 
      spacing = 2,
      Button(loading = TRUE, variant = "outlined", loadingPosition = "start", "Submit"),
      Button(loading = TRUE, variant = "outlined", loadingPosition = "end", "Submit"),
      Button(
        loading = TRUE,
        variant = "outlined",
        loadingPosition = "end",
        startIcon = shiny::icon("save"),
        "Save"
      )
    )
  )
)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import SaveIcon from '@mui/icons-material/Save';
import Stack from '@mui/material/Stack';

export default function LoadingButtons() {
  return (
    <Stack spacing={2}>
      <Stack direction="row" spacing={2}>
        <Button loading variant="outlined">
          Submit
        </Button>
        <Button loading loadingIndicator="Loading…" variant="outlined">
          Fetch data
        </Button>
        <Button
          loading
          loadingPosition="start"
          startIcon={<SaveIcon />}
          variant="outlined"
        >
          Save
        </Button>
      </Stack>
      <Button
        fullWidth
        loading
        loadingPosition="start"
        startIcon={<SaveIcon />}
        variant="outlined"
      >
        Full width
      </Button>
      <Button
        fullWidth
        loading
        loadingPosition="end"
        endIcon={<SaveIcon />}
        variant="outlined"
      >
        Full width
      </Button>
      <Stack direction="row" spacing={2}>
        <Button loading variant="outlined" loadingPosition="start">
          Submit
        </Button>
        <Button loading variant="outlined" loadingPosition="end">
          Submit
        </Button>
        <Button
          loading
          variant="outlined"
          loadingPosition="end"
          startIcon={<SaveIcon />}
        >
          Save
        </Button>
      </Stack>
    </Stack>
  );
}
```
</details>

Toggle the loading switch to see the transition between the different states.

## Loading

```{r eval=FALSE}
ui <- shinyMaterialUIPage(
  div(
    FormControlLabel(
      sx = list(display = "block"),
      control = Switch.shinyInput(
        inputId = "loadingSwitch",
        value = TRUE,
        name = "loading",
        color = "primary"
      ),
      label = "Loading"
    ),
    Box(
      sx = list("& > button" = list(m = 1)),
      Button.shinyInput(
        inputId = "basicButton",
        size = "small",
        loading = TRUE,
        variant = "outlined",
        disabled = TRUE,
        "Disabled"
      ),
      Button.shinyInput(
        inputId = "fetchButton",
        size = "small",
        loading = TRUE,
        loadingIndicator = "Loading…",
        variant = "outlined",
        "Fetch data"
      ),
      Button.shinyInput(
        inputId = "sendButton",
        size = "small",
        endIcon = shiny::icon("paper-plane"),
        loading = TRUE,
        loadingPosition = "end",
        variant = "contained",
        "Send"
      ),
      Button.shinyInput(
        inputId = "saveButton",
        size = "small",
        color = "secondary",
        loading = TRUE,
        loadingPosition = "start",
        startIcon = shiny::icon("save"),
        variant = "contained",
        "Save"
      )
    ),
    Box(
      sx = list("& > button" = list(m = 1)),
      Button.shinyInput(
        inputId = "basicButton2",
        loading = TRUE, 
        variant = "outlined", 
        disabled = TRUE,
        "Disabled"
      ),
      Button.shinyInput(
        inputId = "fetchButtonLarge",
        loading = TRUE,
        loadingIndicator = "Loading…",
        variant = "outlined",
        "Fetch data"
      ),
      Button.shinyInput(
        inputId = "sendButtonLarge",
        endIcon = shiny::icon("paper-plane"),
        loading = TRUE,
        loadingPosition = "end",
        variant = "contained",
        "Send"
      ),
      Button.shinyInput(
        inputId = "saveButtonLarge",
        color = "secondary",
        loading = TRUE,
        loadingPosition = "start",
        startIcon = shiny::icon("save"),
        variant = "contained",
        "Save"
      )
    )
  )
)

server <- function(input, output, session) {
  observe({
    loading <- input$loadingSwitch
    
    updateButton.shinyInput(inputId = "basicButton", loading = loading)
    updateButton.shinyInput(inputId = "fetchButton", loading = loading)
    updateButton.shinyInput(inputId = "sendButton", loading = loading)
    updateButton.shinyInput(inputId = "saveButton", loading = loading)
    updateButton.shinyInput(inputId = "basicButton2", loading = loading)
    updateButton.shinyInput(inputId = "fetchButtonLarge", loading = loading)
    updateButton.shinyInput(inputId = "sendButtonLarge", loading = loading)
    updateButton.shinyInput(inputId = "saveButtonLarge", loading = loading)
  })
}

shinyApp(ui, server)
```


<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Box from '@mui/material/Box';
import FormControlLabel from '@mui/material/FormControlLabel';
import Switch from '@mui/material/Switch';
import SaveIcon from '@mui/icons-material/Save';
import SendIcon from '@mui/icons-material/Send';

export default function LoadingButtonsTransition() {
  const [loading, setLoading] = React.useState(true);
  function handleClick() {
    setLoading(true);
  }

  return (
    <div>
      <FormControlLabel
        sx={{ display: 'block' }}
        control={
          <Switch
            checked={loading}
            onChange={() => setLoading(!loading)}
            name="loading"
            color="primary"
          />
        }
        label="Loading"
      />
      <Box sx={{ '& > button': { m: 1 } }}>
        <Button
          size="small"
          onClick={handleClick}
          loading={loading}
          variant="outlined"
          disabled
        >
          Disabled
        </Button>
        <Button
          size="small"
          onClick={handleClick}
          loading={loading}
          loadingIndicator="Loading…"
          variant="outlined"
        >
          Fetch data
        </Button>
        <Button
          size="small"
          onClick={handleClick}
          endIcon={<SendIcon />}
          loading={loading}
          loadingPosition="end"
          variant="contained"
        >
          Send
        </Button>
        <Button
          size="small"
          color="secondary"
          onClick={handleClick}
          loading={loading}
          loadingPosition="start"
          startIcon={<SaveIcon />}
          variant="contained"
        >
          Save
        </Button>
      </Box>
      <Box sx={{ '& > button': { m: 1 } }}>
        <Button onClick={handleClick} loading={loading} variant="outlined" disabled>
          Disabled
        </Button>
        <Button
          onClick={handleClick}
          loading={loading}
          loadingIndicator="Loading…"
          variant="outlined"
        >
          Fetch data
        </Button>
        <Button
          onClick={handleClick}
          endIcon={<SendIcon />}
          loading={loading}
          loadingPosition="end"
          variant="contained"
        >
          Send
        </Button>
        <Button
          color="secondary"
          onClick={handleClick}
          loading={loading}
          loadingPosition="start"
          startIcon={<SaveIcon />}
          variant="contained"
        >
          Save
        </Button>
      </Box>
    </div>
  );
}
```
</details>

When the loading prop is set to boolean, the loading wrapper is always present in the DOM to prevent a Google Translation Crash.

The loading value should always be null or boolean. The pattern below is not recommended as it can cause the Google Translation crash:

```jsx
<Button {...(isFetching && { loading: true })}> // ❌ Don't do this
```

## Customization

Here are some examples of customizing the component.
You can learn more about this in the overrides documentation page.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import Button from '@mui/material/Button';
import Stack from '@mui/material/Stack';
import { purple } from '@mui/material/colors';

const BootstrapButton = styled(Button)({
  boxShadow: 'none',
  textTransform: 'none',
  fontSize: 16,
  padding: '6px 12px',
  border: '1px solid',
  lineHeight: 1.5,
  backgroundColor: '#0063cc',
  borderColor: '#0063cc',
  fontFamily: [
    '-apple-system',
    'BlinkMacSystemFont',
    '"Segoe UI"',
    'Roboto',
    '"Helvetica Neue"',
    'Arial',
    'sans-serif',
    '"Apple Color Emoji"',
    '"Segoe UI Emoji"',
    '"Segoe UI Symbol"',
  ].join(','),
  '&:hover': {
    backgroundColor: '#0069d9',
    borderColor: '#0062cc',
    boxShadow: 'none',
  },
  '&:active': {
    boxShadow: 'none',
    backgroundColor: '#0062cc',
    borderColor: '#005cbf',
  },
  '&:focus': {
    boxShadow: '0 0 0 0.2rem rgba(0,123,255,.5)',
  },
});

const ColorButton = styled(Button)(({ theme }) => ({
  color: theme.palette.getContrastText(purple[500]),
  backgroundColor: purple[500],
  '&:hover': {
    backgroundColor: purple[700],
  },
}));

export default function CustomizedButtons() {
  return (
    <Stack spacing={2} direction="row">
      <ColorButton variant="contained">Custom CSS</ColorButton>
      <BootstrapButton variant="contained" disableRipple>
        Bootstrap
      </BootstrapButton>
    </Stack>
  );
}
```
</details>

```{r}
library(shiny)

CustomizedButtons <- function() {
  CssBaseline(
    ThemeProvider(
      theme = list(
        components = list(
          MuiButton = list(
            styleOverrides = list(
              root = list(
                "&.bootstrap-button" = list(
                  boxShadow = "none",
                  textTransform = "none",
                  fontSize = 16,
                  padding = "6px 12px",
                  border = "1px solid",
                  lineHeight = 1.5,
                  backgroundColor = "#0063cc",
                  borderColor = "#0063cc",
                  fontFamily = paste(
                    "-apple-system",
                    "BlinkMacSystemFont",
                    '"Segoe UI"',
                    "Roboto",
                    '"Helvetica Neue"',
                    "Arial",
                    "sans-serif",
                    '"Apple Color Emoji"',
                    '"Segoe UI Emoji"',
                    '"Segoe UI Symbol"',
                    sep = ","
                  ),
                  "&:hover" = list(
                    backgroundColor = "#0069d9",
                    borderColor = "#0062cc",
                    boxShadow = "none"
                  ),
                  "&:active" = list(
                    boxShadow = "none",
                    backgroundColor = "#0062cc",
                    borderColor = "#005cbf"
                  ),
                  "&:focus" = list(
                    boxShadow = "0 0 0 0.2rem rgba(0,123,255,.5)"
                  )
                ),
                "&.color-button" = list(
                  color = "#fff",
                  backgroundColor = "#9c27b0",
                  "&:hover" = list(
                    backgroundColor = "#7b1fa2"
                  )
                )
              )
            )
          )
        )
      ),
      Stack(
        spacing = 2,
        direction = "row",
        Button(
          variant = "contained",
          className = "color-button",
          "Custom CSS"
        ),
        Button(
          variant = "contained",
          className = "bootstrap-button",
          disableRipple = TRUE,
          "Bootstrap"
        )
      )
    )
  )
}
CustomizedButtons()
```

## Complex button

The Text Buttons, Contained Buttons, Floating Action Buttons and Icon Buttons are built on top of the same component: the ButtonBase.
You can take advantage of this lower-level component to build custom interactions.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import ButtonBase from '@mui/material/ButtonBase';
import Typography from '@mui/material/Typography';

const images = [
  {
    url: '/static/images/buttons/breakfast.jpg',
    title: 'Breakfast',
    width: '40%',
  },
  {
    url: '/static/images/buttons/burgers.jpg',
    title: 'Burgers',
    width: '30%',
  },
  {
    url: '/static/images/buttons/camera.jpg',
    title: 'Camera',
    width: '30%',
  },
];

export default function ButtonBaseDemo() {
  return (
    <Box sx={{ display: 'flex', flexWrap: 'wrap', minWidth: 300, width: '100%' }}>
      {images.map((image) => (
        <ButtonBase
          focusRipple
          key={image.title}
          sx={{
            position: 'relative',
            width: image.width,
            height: 200,
            '&:hover, &.Mui-focusVisible': {
              zIndex: 1,
              '& .MuiImageBackdrop-root': {
                opacity: 0.15,
              },
              '& .MuiImageMarked-root': {
                opacity: 0,
              },
              '& .MuiTypography-root': {
                border: '4px solid currentColor',
              },
            },
          }}
        >
          <Box
            sx={{
              position: 'absolute',
              left: 0,
              right: 0,
              top: 0,
              bottom: 0,
              backgroundSize: 'cover',
              backgroundPosition: 'center 40%',
              backgroundImage: `url(${image.url})`,
            }}
          />
          <Box
            className="MuiImageBackdrop-root"
            sx={{
              position: 'absolute',
              left: 0,
              right: 0,
              top: 0,
              bottom: 0,
              backgroundColor: 'common.black',
              opacity: 0.4,
              transition: 'opacity 0.3s',
            }}
          />
          <Box
            sx={{
              position: 'absolute',
              left: 0,
              right: 0,
              top: 0,
              bottom: 0,
              display: 'flex',
              alignItems: 'center',
              justifyContent: 'center',
              color: 'common.white',
            }}
          >
            <Typography
              component="span"
              variant="subtitle1"
              color="inherit"
              sx={{
                position: 'relative',
                p: 4,
                pt: 2,
                pb: (theme) => `calc(${theme.spacing(1)} + 6px)`,
              }}
            >
              {image.title}
              <Box
                className="MuiImageMarked-root"
                sx={{
                  height: 3,
                  width: 18,
                  backgroundColor: 'common.white',
                  position: 'absolute',
                  bottom: -2,
                  left: 'calc(50% - 9px)',
                  transition: 'opacity 0.3s',
                }}
              />
            </Typography>
          </Box>
        </ButtonBase>
      ))}
    </Box>
  );
}
```
</details>

```{r}
library(shiny)

ButtonBaseDemo <- function() {
  images <- list(
    list(
      url = "https://mui.com/static/images/buttons/breakfast.jpg",
      title = "Breakfast",
      width = "40%"
    ),
    list(
      url = "https://mui.com/static/images/buttons/burgers.jpg",
      title = "Burgers",
      width = "30%"
    ),
    list(
      url = "https://mui.com/static/images/buttons/camera.jpg",
      title = "Camera",
      width = "30%"
    )
  )
  
  CssBaseline(
    Box(
      sx = list(
        display = "flex",
        flexWrap = "wrap",
        minWidth = 300,
        width = "100%"
      ),
      lapply(images, function(image) {
        ButtonBase(
          focusRipple = TRUE,
          key = image$title,
          sx = list(
            position = "relative",
            width = image$width,
            height = 200,
            "&:hover, &.Mui-focusVisible" = list(
              zIndex = 1,
              "& .MuiImageBackdrop-root" = list(
                opacity = 0.15
              ),
              "& .MuiImageMarked-root" = list(
                opacity = 0
              ),
              "& .MuiTypography-root" = list(
                border = "4px solid currentColor"
              )
            )
          ),
          Box(
            sx = list(
              position = "absolute",
              left = 0,
              right = 0,
              top = 0,
              bottom = 0,
              backgroundSize = "cover",
              backgroundPosition = "center 40%",
              backgroundImage = paste0("url(", image$url, ")")
            )
          ),
          Box(
            className = "MuiImageBackdrop-root",
            sx = list(
              position = "absolute",
              left = 0,
              right = 0,
              top = 0,
              bottom = 0,
              backgroundColor = "common.black",
              opacity = 0.4,
              transition = "opacity 0.3s"
            )
          ),
          Box(
            sx = list(
              position = "absolute",
              left = 0,
              right = 0,
              top = 0,
              bottom = 0,
              display = "flex",
              alignItems = "center",
              justifyContent = "center",
              color = "common.white"
            ),
            Typography(
              component = "span",
              variant = "subtitle1",
              color = "inherit",
              sx = list(
                position = "relative",
                p = 4,
                pt = 2,
                pb = "calc(8px + 6px)"
              ),
              image$title,
              Box(
                className = "MuiImageMarked-root",
                sx = list(
                  height = 3,
                  width = 18,
                  backgroundColor = "common.white",
                  position = "absolute",
                  bottom = -2,
                  left = "calc(50% - 9px)",
                  transition = "opacity 0.3s"
                )
              )
            )
          )
        )
      })
    )
  )
}
ButtonBaseDemo()
```
