---
title: "Button Group"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Button Group}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-button-group/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# Button Group

The ButtonGroup component can be used to group related buttons.

## Basic button group

The buttons can be grouped by wrapping them with the `ButtonGroup` component.
They need to be immediate children.

```{r}
muiMaterialPage(
  CssBaseline(),
  ButtonGroup(variant = "contained", `aria-label` = "Basic button group", Button("One"), Button("Two"), Button("Three"))
)
```

<details><summary>JS code</summary>
```jsx
import Button from '@mui/material/Button';
import ButtonGroup from '@mui/material/ButtonGroup';

export default function BasicButtonGroup() {
  return (
    <ButtonGroup variant="contained" aria-label="Basic button group">
      <Button>One</Button>
      <Button>Two</Button>
      <Button>Three</Button>
    </ButtonGroup>
  );
}
```
</details>

## Button variants

All the standard button variants are supported.

```{r}
buttons <- tagList(Button("One"), Button("Two"), Button("Three"))

muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(display = "flex", flexDirection = "column", alignItems = "center", "& > *" = list(m = 1)),
    ButtonGroup(variant = "outlined", `aria-label` = "Basic button group", buttons),
    ButtonGroup(variant = "text", `aria-label` = "Basic button group", buttons)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Button from '@mui/material/Button';
import ButtonGroup from '@mui/material/ButtonGroup';
import Box from '@mui/material/Box';

export default function VariantButtonGroup() {
  return (
    <Box
      sx={{
        display: 'flex',
        flexDirection: 'column',
        alignItems: 'center',
        '& > *': {
          m: 1,
        },
      }}
    >
      <ButtonGroup variant="outlined" aria-label="Basic button group">
        <Button>One</Button>
        <Button>Two</Button>
        <Button>Three</Button>
      </ButtonGroup>
      <ButtonGroup variant="text" aria-label="Basic button group">
        <Button>One</Button>
        <Button>Two</Button>
        <Button>Three</Button>
      </ButtonGroup>
    </Box>
  );
}
```
</details>

## Sizes and colors

The `size` and `color` props can be used to control the appearance of the button group.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(display = "flex", flexDirection = "column", alignItems = "center", "& > *" = list(m = 1)),
    ButtonGroup(size = "small", `aria-label` = "Small button group", buttons),
    ButtonGroup(color = "secondary", `aria-label` = "Medium-sized button group", buttons),
    ButtonGroup(size = "large", `aria-label` = "Large button group", buttons)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Button from '@mui/material/Button';
import Box from '@mui/material/Box';
import ButtonGroup from '@mui/material/ButtonGroup';

const buttons = [
  <Button key="one">One</Button>,
  <Button key="two">Two</Button>,
  <Button key="three">Three</Button>,
];

export default function GroupSizesColors() {
  return (
    <Box
      sx={{
        display: 'flex',
        flexDirection: 'column',
        alignItems: 'center',
        '& > *': {
          m: 1,
        },
      }}
    >
      <ButtonGroup size="small" aria-label="Small button group">
        {buttons}
      </ButtonGroup>
      <ButtonGroup color="secondary" aria-label="Medium-sized button group">
        {buttons}
      </ButtonGroup>
      <ButtonGroup size="large" aria-label="Large button group">
        {buttons}
      </ButtonGroup>
    </Box>
  );
}
```
</details>

## Vertical group

The button group can be displayed vertically using the `orientation` prop.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(display = "flex", "& > *" = list(m = 1)),
    ButtonGroup(orientation = "vertical", `aria-label` = "Vertical button group", buttons),
    ButtonGroup(orientation = "vertical", `aria-label` = "Vertical button group", variant = "contained", buttons),
    ButtonGroup(orientation = "vertical", `aria-label` = "Vertical button group", variant = "text", buttons)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Button from '@mui/material/Button';
import ButtonGroup from '@mui/material/ButtonGroup';
import Box from '@mui/material/Box';

const buttons = [
  <Button key="one">One</Button>,
  <Button key="two">Two</Button>,
  <Button key="three">Three</Button>,
];

export default function GroupOrientation() {
  return (
    <Box
      sx={{
        display: 'flex',
        '& > *': {
          m: 1,
        },
      }}
    >
      <ButtonGroup orientation="vertical" aria-label="Vertical button group">
        {buttons}
      </ButtonGroup>
      <ButtonGroup
        orientation="vertical"
        aria-label="Vertical button group"
        variant="contained"
      >
        {buttons}
      </ButtonGroup>
      <ButtonGroup
        orientation="vertical"
        aria-label="Vertical button group"
        variant="text"
      >
        {buttons}
      </ButtonGroup>
    </Box>
  );
}
```
</details>

## Split button

`ButtonGroup` can also be used to create a split button. The dropdown can change the button action (as in this example) or be used to immediately trigger a related action.

The MUI demo positions a `Popper()` with a React ref. In R, `Menu.triggerId()`
opens the options next to the arrow button (see [Menu](menu.html)), and the
selected option is kept in the URL with
[reactRouter](https://felixluginbuhl.com/reactRouter/) (`#/merge/<index>`):

```{r, message = FALSE}
library(reactRouter)

options <- c("Create a merge commit", "Squash and merge", "Rebase and merge")

splitButton <- function(selectedIndex) {
  tagList(
    ButtonGroup(
      variant = "contained",
      `aria-label` = "Button group with a nested menu",
      Button(onClick = JS(sprintf("() => { console.info('You clicked %s'); }", options[selectedIndex])), options[selectedIndex]),
      Button(
        id = "split-button-toggle",
        size = "small",
        `aria-label` = "select merge strategy",
        `aria-haspopup` = "menu",
        Icon("arrow_drop_down")
      )
    ),
    Menu.triggerId(
      triggerId = "split-button-toggle",
      id = "split-button-menu",
      anchorOrigin = list(vertical = "bottom", horizontal = "right"),
      transformOrigin = list(vertical = "top", horizontal = "right"),
      lapply(seq_along(options), function(index) {
        MenuItem(
          component = "a",
          href = paste0("#/merge/", index),
          disabled = index == 3,
          selected = index == selectedIndex,
          options[index]
        )
      })
    )
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(1:2, function(index) Route(path = paste0("merge/", index), element = splitButton(index))),
      Route(path = "*", element = splitButton(2))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import ButtonGroup from '@mui/material/ButtonGroup';
import ArrowDropDownIcon from '@mui/icons-material/ArrowDropDown';
import ClickAwayListener from '@mui/material/ClickAwayListener';
import Grow from '@mui/material/Grow';
import Paper from '@mui/material/Paper';
import Popper from '@mui/material/Popper';
import MenuItem from '@mui/material/MenuItem';
import MenuList from '@mui/material/MenuList';

const options = ['Create a merge commit', 'Squash and merge', 'Rebase and merge'];

export default function SplitButton() {
  const [open, setOpen] = React.useState(false);
  const anchorRef = React.useRef(null);
  const [selectedIndex, setSelectedIndex] = React.useState(1);

  const handleClick = () => {
    console.info(`You clicked ${options[selectedIndex]}`);
  };

  const handleMenuItemClick = (event, index) => {
    setSelectedIndex(index);
    setOpen(false);
  };

  const handleToggle = () => {
    setOpen((prevOpen) => !prevOpen);
  };

  const handleClose = (event) => {
    if (anchorRef.current && anchorRef.current.contains(event.target)) {
      return;
    }

    setOpen(false);
  };

  return (
    <React.Fragment>
      <ButtonGroup
        variant="contained"
        ref={anchorRef}
        aria-label="Button group with a nested menu"
      >
        <Button onClick={handleClick}>{options[selectedIndex]}</Button>
        <Button
          size="small"
          aria-controls={open ? 'split-button-menu' : undefined}
          aria-expanded={open ? 'true' : undefined}
          aria-label="select merge strategy"
          aria-haspopup="menu"
          onClick={handleToggle}
        >
          <ArrowDropDownIcon />
        </Button>
      </ButtonGroup>
      <Popper
        sx={{ zIndex: 1 }}
        open={open}
        anchorEl={anchorRef.current}
        role={undefined}
        transition
        disablePortal
      >
        {({ TransitionProps, placement }) => (
          <Grow
            {...TransitionProps}
            style={{
              transformOrigin:
                placement === 'bottom' ? 'center top' : 'center bottom',
            }}
          >
            <Paper>
              <ClickAwayListener onClickAway={handleClose}>
                <MenuList id="split-button-menu" autoFocusItem>
                  {options.map((option, index) => (
                    <MenuItem
                      key={option}
                      disabled={index === 2}
                      selected={index === selectedIndex}
                      onClick={(event) => handleMenuItemClick(event, index)}
                    >
                      {option}
                    </MenuItem>
                  ))}
                </MenuList>
              </ClickAwayListener>
            </Paper>
          </Grow>
        )}
      </Popper>
    </React.Fragment>
  );
}
```
</details>

## Disabled elevation

You can remove the elevation with the `disableElevation` prop.

```{r}
muiMaterialPage(
  CssBaseline(),
  ButtonGroup(disableElevation = TRUE, variant = "contained", `aria-label` = "Disabled button group", Button("One"), Button("Two"))
)
```

<details><summary>JS code</summary>
```jsx
import ButtonGroup from '@mui/material/ButtonGroup';
import Button from '@mui/material/Button';

export default function DisableElevation() {
  return (
    <ButtonGroup
      disableElevation
      variant="contained"
      aria-label="Disabled button group"
    >
      <Button>One</Button>
      <Button>Two</Button>
    </ButtonGroup>
  );
}
```
</details>

## Loading

Use the `loading` prop from `Button` to set buttons in a loading state and disable interactions.

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  ButtonGroup(
    variant = "outlined",
    `aria-label` = "Loading button group",
    Button("Submit"),
    Button("Fetch data"),
    Button(loading = TRUE, loadingPosition = "start", startIcon = Icon("save"), "Save")
  )
)
```

<details><summary>JS code</summary>
```jsx
import ButtonGroup from '@mui/material/ButtonGroup';
import Button from '@mui/material/Button';
import SaveIcon from '@mui/icons-material/Save';

export default function LoadingButtonGroup() {
  return (
    <ButtonGroup variant="outlined" aria-label="Loading button group">
      <Button>Submit</Button>
      <Button>Fetch data</Button>
      <Button loading loadingPosition="start" startIcon={<SaveIcon />}>
        Save
      </Button>
    </ButtonGroup>
  );
}
```
</details>

## Button groups in Shiny apps

Each button of a group can be a `Button.shinyInput()`, which reports its own
click counter:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(p = 2),
    ButtonGroup(
      variant = "contained",
      Button.shinyInput("one", "One"),
      Button.shinyInput("two", "Two"),
      Button.shinyInput("three", "Three")
    ),
    verbatimTextOutput("clicked")
  )
)

server <- function(input, output, session) {
  last <- reactiveVal("none")
  observeEvent(input$one, last("One"))
  observeEvent(input$two, last("Two"))
  observeEvent(input$three, last("Three"))
  output$clicked <- renderText(paste("Last clicked:", last()))
}

shinyApp(ui, server)
```

For a group of mutually exclusive options, use [Toggle Button Group](toggle-button-group.html).
