---
title: "Click-Away Listener"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Click-Away Listener}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-click-away-listener/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Click-Away Listener

The Click-Away Listener component detects when a click event happens outside of its child element.

## Introduction

Click-Away Listener is a utility component that listens for click events outside of its child.
(Note that it only accepts _one_ child element.)
This is useful for components like the [Popper](popper.html) which should close when the user clicks anywhere else in the document.
Click-Away Listener also supports the [Portal](https://mui.com/material-ui/react-portal/) component.

The demo below shows how to hide a menu dropdown when users click anywhere else on the page.
The open state is kept in the URL with
[reactRouter](https://felixluginbuhl.com/reactRouter/): the button toggles
between `#/click-away` and `#/`, and `onClickAway` navigates back to `#/`.

```{r}
dropdownStyles <- list(position = "absolute", top = 28, right = 0, left = 0, zIndex = 1, border = "1px solid", p = 1, bgcolor = "background.paper")

clickAway <- function(route, open, mouseEvent = NULL, touchEvent = NULL, portal = FALSE, styles = dropdownStyles) {
  dropdown <- Box(sx = styles, "Click me, I will stay visible until you click outside.")
  do.call(ClickAwayListener, c(
    list(onClickAway = JS(sprintf(
      "() => { if (window.location.hash === '#/%s') window.location.hash = '/'; }", route
    ))),
    if (!is.null(mouseEvent)) list(mouseEvent = mouseEvent, touchEvent = touchEvent),
    list(
      Box(
        sx = list(position = "relative"),
        Button(variant = "outlined", size = "small", href = if (open) "#/" else paste0("#/", route), "Open menu dropdown"),
        if (open) (if (portal) Portal(dropdown) else dropdown)
      )
    )
  ))
}

clickAwayDemo <- function(route, ...) {
  RouterProvider(
    router = createHashRouter(
      Route(path = route, element = clickAway(route, TRUE, ...)),
      Route(path = "*", element = clickAway(route, FALSE, ...))
    )
  )
}

muiMaterialPage(CssBaseline(), clickAwayDemo("click-away"))
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import ClickAwayListener from '@mui/material/ClickAwayListener';

export default function ClickAway() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen((prev) => !prev);
  };

  const handleClickAway = () => {
    setOpen(false);
  };

  const styles = {
    position: 'absolute',
    top: 28,
    right: 0,
    left: 0,
    zIndex: 1,
    border: '1px solid',
    p: 1,
    bgcolor: 'background.paper',
  };

  return (
    <ClickAwayListener onClickAway={handleClickAway}>
      <Box sx={{ position: 'relative' }}>
        <button type="button" onClick={handleClick}>
          Open menu dropdown
        </button>
        {open ? (
          <Box sx={styles}>
            Click me, I will stay visible until you click outside.
          </Box>
        ) : null}
      </Box>
    </ClickAwayListener>
  );
}
```
</details>

## Customization

### Use with Portal

The following demo uses the `Portal()` component to render the dropdown into a new subtree outside of the current DOM hierarchy:

```{r}
muiMaterialPage(
  CssBaseline(),
  clickAwayDemo(
    "portal-click-away",
    portal = TRUE,
    styles = list(
      position = "fixed", width = 200, top = "50%", left = "50%", transform = "translate(-50%, -50%)",
      border = "1px solid", p = 1, bgcolor = "background.paper"
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import ClickAwayListener from '@mui/material/ClickAwayListener';
import Portal from '@mui/material/Portal';

export default function PortalClickAway() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen((prev) => !prev);
  };

  const handleClickAway = () => {
    setOpen(false);
  };

  const styles = {
    position: 'fixed',
    width: 200,
    top: '50%',
    left: '50%',
    transform: 'translate(-50%, -50%)',
    border: '1px solid',
    p: 1,
    bgcolor: 'background.paper',
  };

  return (
    <ClickAwayListener onClickAway={handleClickAway}>
      <div>
        <button type="button" onClick={handleClick}>
          Open menu dropdown
        </button>
        {open ? (
          <Portal>
            <Box sx={styles}>
              Click me, I will stay visible until you click outside.
            </Box>
          </Portal>
        ) : null}
      </div>
    </ClickAwayListener>
  );
}
```
</details>

### Listening for leading events

By default, the Click-Away Listener component responds to **trailing events**: the _end_ of a click or touch.

You can set the component to listen for **leading events** (the start of a click or touch) using the `mouseEvent` and `touchEvent` props, as shown in the following demo:

<div class="alert alert-warning">
When the component is set to listen for leading events, interactions with the scrollbar are ignored.
</div>

```{r}
muiMaterialPage(CssBaseline(), clickAwayDemo("leading-click-away", mouseEvent = "onMouseDown", touchEvent = "onTouchStart"))
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import ClickAwayListener from '@mui/material/ClickAwayListener';

export default function LeadingClickAway() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen((prev) => !prev);
  };

  const handleClickAway = () => {
    setOpen(false);
  };

  const styles = {
    position: 'absolute',
    top: 28,
    right: 0,
    left: 0,
    zIndex: 1,
    border: '1px solid',
    p: 1,
    bgcolor: 'background.paper',
  };

  return (
    <ClickAwayListener
      mouseEvent="onMouseDown"
      touchEvent="onTouchStart"
      onClickAway={handleClickAway}
    >
      <Box sx={{ position: 'relative' }}>
        <button type="button" onClick={handleClick}>
          Open menu dropdown
        </button>
        {open ? (
          <Box sx={styles}>
            Click me, I will stay visible until you click outside.
          </Box>
        ) : null}
      </Box>
    </ClickAwayListener>
  );
}
```
</details>

## Accessibility

By default, Click-Away Listener adds an `onClick` handler to its child.
This can result in screen readers announcing that the child is clickable, even though this `onClick` handler has no effect on the child itself.

To prevent this behavior, add `role = "presentation"` to the child element:

```r
ClickAwayListener(
  onClickAway = JS("() => {}"),
  Box(role = "presentation", tags$h1("non-interactive heading"))
)
```

This is also required to fix a known issue in NVDA when using Firefox that prevents the announcement of alert messages. See [this GitHub issue](https://github.com/mui/material-ui/issues/29080).

## Click-away in Shiny apps

Send the click-away to the server with `triggerEvent()`, for example to close a
panel rendered with `renderUI()`:

```r
ClickAwayListener(onClickAway = triggerEvent("clicked_away"), Box(uiOutput("panel")))
```
