---
title: "Container"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Container}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-container/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
```

# Container

The container centers your content horizontally. It's the most basic layout element.

While containers can be nested, most layouts do not require a nested container.

## Fluid

A fluid container width is bounded by the `maxWidth` prop value.
(The demos use a height of 200px instead of the full viewport height `100vh`.)

```{r}
muiMaterialPage(
  CssBaseline(),
  Container(maxWidth = "sm", Box(sx = list(bgcolor = "#cfe8fc", height = 200)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import CssBaseline from '@mui/material/CssBaseline';
import Box from '@mui/material/Box';
import Container from '@mui/material/Container';

export default function SimpleContainer() {
  return (
    <React.Fragment>
      <CssBaseline />
      <Container maxWidth="sm">
        <Box sx={{ bgcolor: '#cfe8fc', height: '100vh' }} />
      </Container>
    </React.Fragment>
  );
}
```
</details>

```r
Container(maxWidth = "sm")
```

## Fixed

If you prefer to design for a fixed set of sizes instead of trying to accommodate a fully fluid viewport, you can set the `fixed` prop.
The max-width matches the min-width of the current breakpoint.

```{r}
muiMaterialPage(
  CssBaseline(),
  Container(fixed = TRUE, Box(sx = list(bgcolor = "#cfe8fc", height = 200)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import CssBaseline from '@mui/material/CssBaseline';
import Box from '@mui/material/Box';
import Container from '@mui/material/Container';

export default function FixedContainer() {
  return (
    <React.Fragment>
      <CssBaseline />
      <Container fixed>
        <Box sx={{ bgcolor: '#cfe8fc', height: '100vh' }} />
      </Container>
    </React.Fragment>
  );
}
```
</details>

```r
Container(fixed = TRUE)
```

## Containers in Shiny apps

`Container()` is the usual wrapper of a page's main content, below an
[App Bar](app-bar.html):

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  AppBar(position = "static", Toolbar(Typography(variant = "h6", "My app"))),
  Container(
    maxWidth = "md",
    sx = list(py = 3),
    Typography(variant = "h4", gutterBottom = TRUE, "Welcome"),
    Typography("The content is centered and at most 900px wide.")
  )
)

shinyApp(ui, function(input, output, session) {})
```
