---
title: "CSS Baseline"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{CSS Baseline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-css-baseline/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# CSS Baseline

The CssBaseline component helps to kickstart an elegant, consistent, and simple baseline to build upon.

## Global reset

You might be familiar with [normalize.css](https://github.com/necolas/normalize.css), a collection of HTML element and attribute style-normalizations.
Add `CssBaseline()` once, at the top of your page:

```r
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  # The rest of your application
  Typography("Hello")
)
```

## Scoping on children

However, you might be progressively migrating a website to Material UI, using a global reset might not be an option.
It's possible to apply the baseline only to the children by using the `ScopedCssBaseline` component.
This is useful, for example, when mixing MUI components with a Bootstrap page
(see [CSS conflicts with Bootstrap](bootstrap.html)):

```r
ScopedCssBaseline(
  # The rest of your application
  Typography("Hello")
)
```

## Approach

### Page

The `<html>` and `<body>` elements are updated to provide better page-wide defaults. More specifically:

- The margin in all browsers is removed.
- The default Material Design background color is applied.
  It's using `theme.palette.background.default` for standard devices and a white background for print devices.
- If `enableColorScheme` is provided to `CssBaseline`, native components color will be set by applying [`color-scheme`](https://web.dev/articles/color-scheme) on `<html>`.
  The value used is provided by the theme property `theme.palette.mode`.

### Layout

- `box-sizing` is set globally on the `<html>` element to `border-box`.
  Every element (including `*::before` and `*::after`) is declared to inherit this property,
  which ensures that the declared width of the element is never exceeded due to padding or border.

### Color scheme

This API switches between `"light"` and `"dark"` modes of native components such as scrollbar, using the `color-scheme` CSS property.
To enable it, you can set `enableColorScheme = TRUE` as follows:

```r
CssBaseline(enableColorScheme = TRUE)
# or
ScopedCssBaseline(enableColorScheme = TRUE, "The rest of your application using color-scheme")
```

### Typography

- No base font-size is declared on the `<html>`, but 16px is assumed (the browser default).
  You can learn more about the implications of changing the `<html>` default font size in [the theme documentation](https://mui.com/material-ui/customization/typography/#html-font-size) page.
- Set the `theme.typography.body1` style on the `<body>` element.
- Set the font-weight to `theme.typography.fontWeightBold` for the `<b>` and `<strong>` elements.
- Custom font-smoothing is enabled for better display of the Roboto font.

Here is a scoped baseline next to unstyled content. The scoped part gets the
MUI body typography and background:

```{r}
muiMaterialPage(
  Box(
    sx = list(display = "flex", gap = 2),
    Box(sx = list(p = 2, border = 1, borderColor = "divider", flex = 1), tags$b("Without baseline"), tags$p("Plain text.")),
    ScopedCssBaseline(
      Box(sx = list(p = 2, border = 1, borderColor = "divider"), tags$b("With ScopedCssBaseline"), tags$p("Plain text."))
    )
  )
)
```

## Customization

Head to the [global customization](https://mui.com/material-ui/customization/how-to-customize/#4-global-css-override) section of the documentation to change the output of these components.
In R, `GlobalStyles()` adds global CSS rules with the same syntax as `sx`:

```{r}
muiMaterialPage(
  CssBaseline(),
  GlobalStyles(styles = list(".highlight-me" = list(backgroundColor = "#fff59d", padding = "0 4px"))),
  Typography("Some ", tags$span(class = "highlight-me", "highlighted"), " text.")
)
```
