---
title: "Custom components"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Custom components}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

This guide shows three ways to go beyond the components shipped with
muiMaterial, from the simplest to the most flexible:

1. [Reusable components in R](#reusable-components-in-r): R functions that return muiMaterial components.
2. [Custom React components](#custom-react-components): JavaScript components built with the MUI modules bundled in muiMaterial.
3. [Custom Shiny inputs](#custom-shiny-inputs): JavaScript components that send a value to the Shiny server.

## Reusable components in R

In React, a reusable component is a function or a `styled()` component. In R,
it is a function that returns muiMaterial components. Arguments set the props,
and `...` passes the rest:

```{r}
# A reusable "stat card"
StatCard <- function(title, value, trend, ...) {
  up <- trend >= 0
  Card(
    variant = "outlined",
    ...,
    CardContent(
      Typography(sx = list(color = "text.secondary", fontSize = 14), gutterBottom = TRUE, title),
      Typography(variant = "h4", component = "div", value),
      Chip(
        size = "small",
        color = if (up) "success" else "error",
        label = sprintf("%+.0f%%", trend),
        sx = list(mt = 1)
      )
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  Grid(
    container = TRUE,
    spacing = 2,
    Grid(size = list(xs = 12, sm = 4), StatCard("Users", "14k", 25)),
    Grid(size = list(xs = 12, sm = 4), StatCard("Conversions", "325", -12)),
    Grid(size = list(xs = 12, sm = 4), StatCard("Event count", "200k", 5, sx = list(bgcolor = "grey.50")))
  )
)
```

Most `styled()` components of the MUI documentation translate this way: the
styles become an `sx` list (see [The `sx` prop in R](box.html#the-sx-prop-in-r)),
and variants become R arguments. To restyle **all** instances of a component,
use the theme instead (see [Theming](theming.html#themed-components)).

## Custom React components

For behaviour that needs JavaScript (state, effects, `styled()`), define a
React component in a script and render it with `shiny.react::reactElement()`.
The script finds React and the bundled MUI modules on the global `jsmodule`
object, and registers the component under its own module name:

```{r}
CustomComponents <- tags$script(HTML(
  "(function() {
    const React = jsmodule['react'];
    const { Button, styled } = jsmodule['@mui/material'];
    const CustomComponents = jsmodule['CustomComponents'] ??= {};

    // A styled() component, as in the MUI documentation
    const GradientButton = styled(Button)({
      background: 'linear-gradient(45deg, #FE6B8B 30%, #FF8E53 90%)',
      border: 0,
      borderRadius: 3,
      boxShadow: '0 3px 5px 2px rgba(255, 105, 135, .3)',
      color: 'white',
      height: 48,
      padding: '0 30px',
    });

    // A component with its own state: counts its clicks
    CustomComponents.ClickCounter = function ClickCounter({ label }) {
      const [count, setCount] = React.useState(0);
      return React.createElement(GradientButton, { onClick: () => setCount(count + 1) }, `${label}: ${count}`);
    };
  })();"
))

ClickCounter <- function(...) {
  shiny.react::reactElement(
    module = "CustomComponents",
    name = "ClickCounter",
    props = shiny.react::asProps(...),
    deps = muiMaterialDependency()
  )
}

muiMaterialPage(
  CssBaseline(),
  CustomComponents,
  ClickCounter(label = "Clicks")
)
```

- `muiMaterialDependency()` makes sure the muiMaterial bundle (which fills
  `jsmodule`) is loaded before the component renders.
- The script must be on the page before the component, as above.
- Write the component with `React.createElement()`: the script is not
  compiled, so JSX is not available.

## Custom Shiny inputs

shiny.react provides two adapters that turn a React component into a Shiny
input. All the `.shinyInput()` wrappers of muiMaterial are built with them:

- `InputAdapter(Component, (value, setValue, props) => props)` maps the value of
  the input to the props of the component. Call `setValue()` to send a new value
  to `input[[inputId]]`. An optional third argument limits how often the value is
  sent, for example `{ policy: debounce, delay: 250 }`.
- `ButtonAdapter(Component)` sends a click counter, like `Button.shinyInput()`.

The example below builds a text field that upper-cases its input. Its value can
be changed from the server with `shiny.react::updateReactInput()`:

```{r, eval = FALSE}
library(shiny)
library(muiMaterial)

CustomInputs <- tags$script(HTML(
  "(function() {
    const { InputAdapter, debounce } = jsmodule['@/shiny.react'];
    const { TextField } = jsmodule['@mui/material'];
    const CustomInputs = jsmodule['CustomInputs'] ??= {};

    CustomInputs.UpperCaseTextField = InputAdapter(TextField, (value, setValue) => ({
      value: (value ?? '').toUpperCase(),
      onChange: (e) => setValue(e.target.value.toUpperCase()),
    }), { policy: debounce, delay: 250 });
  })();"
))

UpperCaseTextField.shinyInput <- function(inputId, ..., value = "") {
  shiny.react::reactElement(
    module = "CustomInputs",
    name = "UpperCaseTextField",
    props = shiny.react::asProps(inputId = inputId, ..., value = value),
    deps = muiMaterialDependency()
  )
}
updateUpperCaseTextField.shinyInput <- shiny.react::updateReactInput

ui <- muiMaterialPage(
  CssBaseline(),
  CustomInputs,
  Box(
    sx = list(p = 2),
    UpperCaseTextField.shinyInput("code", label = "Code", value = "abc"),
    Button.shinyInput("reset", "Reset"),
    verbatimTextOutput("value")
  )
)

server <- function(input, output, session) {
  output$value <- renderText(input$code)
  observeEvent(input$reset, updateUpperCaseTextField.shinyInput(inputId = "code", value = ""))
}

shinyApp(ui, server)
```

Run the bundled examples with `muiMaterialExample("CustomComponentShinyInput")`
and `muiMaterialExample("CustomComponentShinyInputStyled")` (a `styled()`
slider).

## Available JavaScript modules

The muiMaterial bundle registers these modules on `window.jsmodule`:

| Module | Content |
|---|---|
| `jsmodule['react']`, `jsmodule['react-dom']` | React, provided by shiny.react |
| `jsmodule['@/shiny.react']` | `InputAdapter`, `ButtonAdapter`, `debounce`, ... |
| `jsmodule['@mui/material']` | all Material UI components and `styled`, `createTheme`, `alpha`, ... |
| `jsmodule['@mui/lab']` | the lab components (Timeline, Masonry, TabContext, ...) |
| `jsmodule['@mui/system']`, `jsmodule['@mui/utils']` | MUI System and utilities |
| `jsmodule['@emotion/react']`, `jsmodule['@emotion/styled']`, `jsmodule['@emotion/cache']` | the Emotion styling engine |
| `jsmodule['@/muiMaterial']` | the muiMaterial wrappers (`.shinyInput`, `.triggerId`, `ThemeProvider`, ...) |

`@mui/icons-material` is not bundled; see [Icons](icons.html).

## Compatibility

- **React 18.** muiMaterial is built and tested against React 18, which is
  provided at runtime by shiny.react. A warning is printed in the browser
  console if another React version is loaded.
- **Other shiny.react packages.** Packages built on shiny.react (such as
  [muiDataGrid](https://felixluginbuhl.com/muiDataGrid/), shiny.fluent or
  shiny.blueprint) share the same `jsmodule` object. The MUI modules above are
  shared on purpose, so that a `ThemeProvider()` of muiMaterial also styles the
  components of companion packages. If another package registers a different copy
  of these modules, muiMaterial prints a warning in the browser console.
