---
title: "Getting Started"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(muiMaterial)
library(shiny)
```

# Getting Started

## Installation

```r
install.packages("muiMaterial")
```

Or install the development version:

```r
remotes::install_github("lgnbhl/muiMaterial")
```

## Your first Material UI app

```{r, eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2),
      Typography("Hello Material UI!", variant = "h4")
    )
  )
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

**Important**: Use `muiMaterialPage()` instead of `fluidPage()` and wrap your UI in `CssBaseline()` to ensure proper styling. Material UI uses its own design system and can conflict with Bootstrap (see [Bootstrap conflict] below).

## Shiny input wrappers

Material UI components become Shiny inputs with `*.shinyInput()` wrappers. For example, use `Button.shinyInput()` instead of `Button()` to capture user interactions in Shiny.

Explore available Shiny inputs with the showcase app (live [here](https://lgnbhl-muimaterial-showcase.share.connect.posit.cloud/)):

```r
muiMaterial::muiMaterialExample("showcase")
```

## Server-side rendering

When rendering components from the server, use `shiny::renderUI()` or `shiny.react::renderReact()` in your server function, and `shiny::uiOutput()` or `shiny.react::reactOutput()` in your UI.

## Creating tabs

Use `TabContext.shinyInput()`, `TabList.shinyInput()`, and `TabPanel.shinyInput()` to create tabs with server-side reactivity, or `TabContext.static()`, `TabList.static()`, and `TabPanel.static()` for purely client-side tab switching. See [full example](https://github.com/lgnbhl/muiMaterial/blob/main/inst/examples/Tabs.R).

For more advanced navigation, use client-side routing with [reactRouter](https://felixluginbuhl.com/reactRouter/).

## Styling with sx

Customize any component using the `sx` argument for inline CSS-in-JS styling:

```{r}
Box(
  sx = list(
    bgcolor = "primary.main",
    color = "white",
    p = 3,
    borderRadius = 2
  ),
  Typography("Styled with sx", variant = "h5")
)
```

It is more powerful and maintainable than traditional CSS. See the [MUI sx documentation](https://mui.com/system/getting-started/the-sx-prop/) for all available properties.

## Naming conventions

Function names in `muiMaterial` mirror the original [MUI component names](https://mui.com/material-ui/all-components/) as closely as possible (e.g. `Button()`, `Typography()`, `Drawer()`). Suffixes are used to distinguish variants that add R/Shiny-specific behavior:

- **`.shinyInput`** — Wraps a component as a Shiny input, so its value is available server-side via `input$inputId`. Use these when you need to read or react to user interactions in R. This convention is borrowed from the [shiny.fluent](https://appsilon.github.io/shiny.fluent/) package. Examples: `Button.shinyInput()`, `Slider.shinyInput()`, `Autocomplete.shinyInput()`.

- **`.triggerId`** — Wraps a component that binds to an external DOM element by its HTML `id`. The referenced element acts as a trigger (e.g. a button click opens a drawer or menu), with open/close state managed entirely on the client side. Examples: `Drawer.triggerId()`, `Menu.triggerId()`.

- **`.static`** — Wraps a component whose state is managed entirely on the client side (in React), with no round-trip to the Shiny server. Useful when server-side reactivity is not needed, for example tab switching that only shows/hides content. Examples: `TabContext.static()`, `TabList.static()`, `TabPanel.static()`.

Both `.triggerId` and `.static` variants keep state in the browser and never communicate with the Shiny server. The difference is that `.triggerId` components require a reference to an external trigger element, while `.static` components are self-contained.

## CSS conflicts with Bootstrap

`muiMaterialPage()` suppresses Bootstrap by default, giving MUI components a clean CSS environment. If you need to mix MUI with Bootstrap-dependent packages (e.g. `shiny`, `bslib`, `DT`, `plotly`), see the [CSS conflicts with Bootstrap](bootstrap.html) vignette for details and workarounds.
