---
title: "Getting Started"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(muiMaterial)
library(shiny)
```

# Getting Started

## Installation

```r
install.packages("muiMaterial")
```

Or install the development version:

```r
remotes::install_github("lgnbhl/muiMaterial")
```

## Your first Material UI app

```{r, eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(
    Box(
      sx = list(p = 2),
      Typography("Hello Material UI!", variant = "h4")
    )
  )
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

**Important**: Use `muiMaterialPage()` instead of `fluidPage()` and wrap your UI in `CssBaseline()` to ensure proper styling. Material UI uses its own design system and conflicts with Bootstrap.

## Shiny input wrappers

Material UI components become Shiny inputs with `*.shinyInput()` wrappers. For example, use `Button.shinyInput()` instead of `Button()` to capture user interactions in Shiny.

Explore available Shiny inputs with the showcase app (live [here](https://lgnbhl-muimaterial-showcase.share.connect.posit.cloud/)):

```r
muiMaterial::muiMaterialExample("showcase")
```

## Server-side rendering

When rendering components from the server, use `shiny::renderUI()` or `shiny.react::renderReact()` in your server function, and `shiny::uiOutput()` or `shiny.react::reactOutput()` in your UI.

## Creating tabs

Use `TabContext.shinyInput()`, `TabList.shinyInput()`, and `TabPanel.shinyInput()` instead of the basic `Tabs()` component (which currently doesn't work). See [full example](https://github.com/lgnbhl/muiMaterial/blob/main/inst/examples/Tabs.R).

For more advanced navigation, use client-side routing with [reactRouter](https://felixluginbuhl.com/reactRouter/).

## Styling with sx

Customize any component using the `sx` argument for inline CSS-in-JS styling:

```{r}
Box(
  sx = list(
    bgcolor = "primary.main",
    color = "white",
    p = 3,
    borderRadius = 2
  ),
  Typography("Styled with sx", variant = "h5")
)
```

It is more powerful and maintainable than traditional CSS. See the [MUI sx documentation](https://mui.com/system/getting-started/the-sx-prop/) for all available properties.
