---
title: "Image List"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Image List}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-image-list/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# Image List

The Image List displays a collection of images in an organized grid.

Image lists represent a collection of items in a repeated pattern. They help improve the visual comprehension of the content they hold.

The demos use the photos of the MUI documentation, stored in a data frame:

```{r}
unsplash <- "https://images.unsplash.com/photo-"
photos <- data.frame(
  img = paste0(unsplash, c(
    "1551963831-b3b1ca40c98e", "1551782450-a2132b4ba21d", "1522770179533-24471fcdba45",
    "1444418776041-9c7e33cc5a9c", "1533827432537-70133748f5c8", "1558642452-9d2a7deb7f62",
    "1516802273409-68526ee1bdd6", "1518756131217-31eb79b20e8f", "1597645587822-e99fa5d45d25",
    "1567306301408-9b74779a11af", "1471357674240-e1a485acb3e1", "1589118949245-7d38baf380d6"
  )),
  title = c("Breakfast", "Burger", "Camera", "Coffee", "Hats", "Honey", "Basketball", "Fern",
            "Mushrooms", "Tomato basil", "Sea star", "Bike"),
  author = c("@bkristastucchio", "@rollelflex_graphy726", "@helloimnik", "@nolanissac", "@hjrc33",
             "@arwinneil", "@tjdragotta", "@katie_wasserman", "@silverdalex", "@shelleypauls",
             "@peterlaster", "@southside_customs"),
  rows = c(2, 1, 1, 1, 1, 2, 1, 1, 2, 1, 1, 1),
  cols = c(2, 1, 1, 2, 2, 2, 1, 1, 2, 1, 1, 2),
  featured = c(TRUE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)
)

# An <img> tag with the resizing query string of the Unsplash CDN
photo <- function(img, title, query) {
  tags$img(
    srcSet = sprintf("%s?%s&auto=format&dpr=2 2x", img, query),
    src = sprintf("%s?%s&auto=format", img, query),
    alt = title,
    loading = "lazy"
  )
}
```

## Standard image list

Standard image lists are best for items of equal importance. They have a uniform container size, ratio, and spacing.

```{r}
muiMaterialPage(
  CssBaseline(),
  ImageList(
    sx = list(width = 500, maxWidth = "100%", height = 450),
    cols = 3,
    rowHeight = 164,
    lapply(seq_len(nrow(photos)), function(i) {
      ImageListItem(photo(photos$img[i], photos$title[i], "w=164&h=164&fit=crop"))
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';

export default function StandardImageList() {
  return (
    <ImageList sx={{ width: 500, height: 450 }} cols={3} rowHeight={164}>
      {itemData.map((item) => (
        <ImageListItem key={item.img}>
          <img
            srcSet={`${item.img}?w=164&h=164&fit=crop&auto=format&dpr=2 2x`}
            src={`${item.img}?w=164&h=164&fit=crop&auto=format`}
            alt={item.title}
            loading="lazy"
          />
        </ImageListItem>
      ))}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1551963831-b3b1ca40c98e',
    title: 'Breakfast',
  },
  {
    img: 'https://images.unsplash.com/photo-1551782450-a2132b4ba21d',
    title: 'Burger',
  },
  {
    img: 'https://images.unsplash.com/photo-1522770179533-24471fcdba45',
    title: 'Camera',
  },
  {
    img: 'https://images.unsplash.com/photo-1444418776041-9c7e33cc5a9c',
    title: 'Coffee',
  },
  {
    img: 'https://images.unsplash.com/photo-1533827432537-70133748f5c8',
    title: 'Hats',
  },
  {
    img: 'https://images.unsplash.com/photo-1558642452-9d2a7deb7f62',
    title: 'Honey',
  },
  {
    img: 'https://images.unsplash.com/photo-1516802273409-68526ee1bdd6',
    title: 'Basketball',
  },
  {
    img: 'https://images.unsplash.com/photo-1518756131217-31eb79b20e8f',
    title: 'Fern',
  },
  {
    img: 'https://images.unsplash.com/photo-1597645587822-e99fa5d45d25',
    title: 'Mushrooms',
  },
  {
    img: 'https://images.unsplash.com/photo-1567306301408-9b74779a11af',
    title: 'Tomato basil',
  },
  {
    img: 'https://images.unsplash.com/photo-1471357674240-e1a485acb3e1',
    title: 'Sea star',
  },
  {
    img: 'https://images.unsplash.com/photo-1589118949245-7d38baf380d6',
    title: 'Bike',
  },
];
```
</details>

## Quilted image list

Quilted image lists emphasize certain items over others in a collection. They create hierarchy using varied container sizes and ratios.

```{r}
size <- 121

muiMaterialPage(
  CssBaseline(),
  ImageList(
    sx = list(width = 500, maxWidth = "100%", height = 450),
    variant = "quilted",
    cols = 4,
    rowHeight = size,
    lapply(seq_len(nrow(photos)), function(i) {
      ImageListItem(
        cols = photos$cols[i],
        rows = photos$rows[i],
        photo(photos$img[i], photos$title[i], sprintf("w=%s&h=%s&fit=crop", size * photos$cols[i], size * photos$rows[i]))
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';

function srcset(image, size, rows = 1, cols = 1) {
  return {
    src: `${image}?w=${size * cols}&h=${size * rows}&fit=crop&auto=format`,
    srcSet: `${image}?w=${size * cols}&h=${
      size * rows
    }&fit=crop&auto=format&dpr=2 2x`,
  };
}

export default function QuiltedImageList() {
  return (
    <ImageList
      sx={{ width: 500, height: 450 }}
      variant="quilted"
      cols={4}
      rowHeight={121}
    >
      {itemData.map((item) => (
        <ImageListItem key={item.img} cols={item.cols || 1} rows={item.rows || 1}>
          <img
            {...srcset(item.img, 121, item.rows, item.cols)}
            alt={item.title}
            loading="lazy"
          />
        </ImageListItem>
      ))}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1551963831-b3b1ca40c98e',
    title: 'Breakfast',
    rows: 2,
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1551782450-a2132b4ba21d',
    title: 'Burger',
  },
  {
    img: 'https://images.unsplash.com/photo-1522770179533-24471fcdba45',
    title: 'Camera',
  },
  {
    img: 'https://images.unsplash.com/photo-1444418776041-9c7e33cc5a9c',
    title: 'Coffee',
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1533827432537-70133748f5c8',
    title: 'Hats',
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1558642452-9d2a7deb7f62',
    title: 'Honey',
    author: '@arwinneil',
    rows: 2,
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1516802273409-68526ee1bdd6',
    title: 'Basketball',
  },
  {
    img: 'https://images.unsplash.com/photo-1518756131217-31eb79b20e8f',
    title: 'Fern',
  },
  {
    img: 'https://images.unsplash.com/photo-1597645587822-e99fa5d45d25',
    title: 'Mushrooms',
    rows: 2,
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1567306301408-9b74779a11af',
    title: 'Tomato basil',
  },
  {
    img: 'https://images.unsplash.com/photo-1471357674240-e1a485acb3e1',
    title: 'Sea star',
  },
  {
    img: 'https://images.unsplash.com/photo-1589118949245-7d38baf380d6',
    title: 'Bike',
    cols: 2,
  },
];
```
</details>

## Woven image list

Woven image lists use alternating container ratios to create a rhythmic layout. A woven image list is best for browsing peer content.

```{r}
muiMaterialPage(
  CssBaseline(),
  ImageList(
    sx = list(width = 500, maxWidth = "100%", height = 450),
    variant = "woven",
    cols = 3,
    gap = 8,
    lapply(seq_len(nrow(photos)), function(i) ImageListItem(photo(photos$img[i], photos$title[i], "w=161&fit=crop")))
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';

export default function WovenImageList() {
  return (
    <ImageList sx={{ width: 500, height: 450 }} variant="woven" cols={3} gap={8}>
      {itemData.map((item) => (
        <ImageListItem key={item.img}>
          <img
            srcSet={`${item.img}?w=161&fit=crop&auto=format&dpr=2 2x`}
            src={`${item.img}?w=161&fit=crop&auto=format`}
            alt={item.title}
            loading="lazy"
          />
        </ImageListItem>
      ))}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1549388604-817d15aa0110',
    title: 'Bed',
  },
  {
    img: 'https://images.unsplash.com/photo-1563298723-dcfebaa392e3',
    title: 'Kitchen',
  },
  {
    img: 'https://images.unsplash.com/photo-1523413651479-597eb2da0ad6',
    title: 'Sink',
  },
  {
    img: 'https://images.unsplash.com/photo-1525097487452-6278ff080c31',
    title: 'Books',
  },
  {
    img: 'https://images.unsplash.com/photo-1574180045827-681f8a1a9622',
    title: 'Chairs',
  },
  {
    img: 'https://images.unsplash.com/photo-1597262975002-c5c3b14bbd62',
    title: 'Candle',
  },
  {
    img: 'https://images.unsplash.com/photo-1530731141654-5993c3016c77',
    title: 'Laptop',
  },
  {
    img: 'https://images.unsplash.com/photo-1481277542470-605612bd2d61',
    title: 'Doors',
  },
  {
    img: 'https://images.unsplash.com/photo-1517487881594-2787fef5ebf7',
    title: 'Coffee',
  },
  {
    img: 'https://images.unsplash.com/photo-1516455207990-7a41ce80f7ee',
    title: 'Storage',
  },
  {
    img: 'https://images.unsplash.com/photo-1519710164239-da123dc03ef4',
    title: 'Coffee table',
  },
  {
    img: 'https://images.unsplash.com/photo-1588436706487-9d55d73a39e3',
    title: 'Blinds',
  },
];
```
</details>

## Masonry image list

Masonry image lists use dynamically sized container heights that reflect the aspect ratio of each image. This image list is best used for browsing uncropped peer content.

```{r}
interiors <- data.frame(
  img = paste0(unsplash, c(
    "1549388604-817d15aa0110", "1525097487452-6278ff080c31", "1523413651479-597eb2da0ad6",
    "1563298723-dcfebaa392e3", "1588436706487-9d55d73a39e3", "1574180045827-681f8a1a9622",
    "1530731141654-5993c3016c77", "1481277542470-605612bd2d61", "1517487881594-2787fef5ebf7",
    "1516455207990-7a41ce80f7ee", "1597262975002-c5c3b14bbd62", "1519710164239-da123dc03ef4"
  )),
  title = c("Bed", "Books", "Sink", "Kitchen", "Blinds", "Chairs", "Laptop", "Doors", "Coffee",
            "Storage", "Candle", "Coffee table"),
  author = c("swabdesign", "Pavel Nekoranec", "Charles Deluvio", "Christian Mackie", "Darren Richardson",
             "Taylor Simpson", "Ben Kolde", "Philipp Berndt", "Jen P.", "Douglas Sheppard",
             "Fi Bell", "Hutomo Abrianto")
)

muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = 500, maxWidth = "100%", height = 450, overflowY = "scroll"),
    ImageList(
      variant = "masonry",
      cols = 3,
      gap = 8,
      lapply(seq_len(nrow(interiors)), function(i) ImageListItem(photo(interiors$img[i], interiors$title[i], "w=248&fit=crop")))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Box from '@mui/material/Box';
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';

export default function MasonryImageList() {
  return (
    <Box sx={{ width: 500, height: 450, overflowY: 'scroll' }}>
      <ImageList variant="masonry" cols={3} gap={8}>
        {itemData.map((item) => (
          <ImageListItem key={item.img}>
            <img
              srcSet={`${item.img}?w=248&fit=crop&auto=format&dpr=2 2x`}
              src={`${item.img}?w=248&fit=crop&auto=format`}
              alt={item.title}
              loading="lazy"
            />
          </ImageListItem>
        ))}
      </ImageList>
    </Box>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1549388604-817d15aa0110',
    title: 'Bed',
  },
  {
    img: 'https://images.unsplash.com/photo-1525097487452-6278ff080c31',
    title: 'Books',
  },
  {
    img: 'https://images.unsplash.com/photo-1523413651479-597eb2da0ad6',
    title: 'Sink',
  },
  {
    img: 'https://images.unsplash.com/photo-1563298723-dcfebaa392e3',
    title: 'Kitchen',
  },
  {
    img: 'https://images.unsplash.com/photo-1588436706487-9d55d73a39e3',
    title: 'Blinds',
  },
  {
    img: 'https://images.unsplash.com/photo-1574180045827-681f8a1a9622',
    title: 'Chairs',
  },
  {
    img: 'https://images.unsplash.com/photo-1530731141654-5993c3016c77',
    title: 'Laptop',
  },
  {
    img: 'https://images.unsplash.com/photo-1481277542470-605612bd2d61',
    title: 'Doors',
  },
  {
    img: 'https://images.unsplash.com/photo-1517487881594-2787fef5ebf7',
    title: 'Coffee',
  },
  {
    img: 'https://images.unsplash.com/photo-1516455207990-7a41ce80f7ee',
    title: 'Storage',
  },
  {
    img: 'https://images.unsplash.com/photo-1597262975002-c5c3b14bbd62',
    title: 'Candle',
  },
  {
    img: 'https://images.unsplash.com/photo-1519710164239-da123dc03ef4',
    title: 'Coffee table',
  },
];
```
</details>

## Image list with title bars

This example demonstrates the use of the `ImageListItemBar` to add an overlay to each item.
The overlay can accommodate a `title`, `subtitle` and secondary action. In this example, an `IconButton`.

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  ImageList(
    sx = list(width = 500, maxWidth = "100%", height = 450),
    ImageListItem(cols = 2, ListSubheader(component = "div", "December")),
    lapply(seq_len(nrow(photos)), function(i) {
      ImageListItem(
        photo(photos$img[i], photos$title[i], "w=248&fit=crop"),
        ImageListItemBar(
          title = photos$title[i],
          subtitle = photos$author[i],
          actionIcon = IconButton(
            sx = list(color = "rgba(255, 255, 255, 0.54)"),
            `aria-label` = paste("info about", photos$title[i]),
            Icon("info")
          )
        )
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';
import ImageListItemBar from '@mui/material/ImageListItemBar';
import ListSubheader from '@mui/material/ListSubheader';
import IconButton from '@mui/material/IconButton';
import InfoIcon from '@mui/icons-material/Info';

export default function TitlebarImageList() {
  return (
    <ImageList sx={{ width: 500, height: 450 }}>
      <ImageListItem key="Subheader" cols={2}>
        <ListSubheader component="div">December</ListSubheader>
      </ImageListItem>
      {itemData.map((item) => (
        <ImageListItem key={item.img}>
          <img
            srcSet={`${item.img}?w=248&fit=crop&auto=format&dpr=2 2x`}
            src={`${item.img}?w=248&fit=crop&auto=format`}
            alt={item.title}
            loading="lazy"
          />
          <ImageListItemBar
            title={item.title}
            subtitle={item.author}
            actionIcon={
              <IconButton
                sx={{ color: 'rgba(255, 255, 255, 0.54)' }}
                aria-label={`info about ${item.title}`}
              >
                <InfoIcon />
              </IconButton>
            }
          />
        </ImageListItem>
      ))}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1551963831-b3b1ca40c98e',
    title: 'Breakfast',
    author: '@bkristastucchio',
    rows: 2,
    cols: 2,
    featured: true,
  },
  {
    img: 'https://images.unsplash.com/photo-1551782450-a2132b4ba21d',
    title: 'Burger',
    author: '@rollelflex_graphy726',
  },
  {
    img: 'https://images.unsplash.com/photo-1522770179533-24471fcdba45',
    title: 'Camera',
    author: '@helloimnik',
  },
  {
    img: 'https://images.unsplash.com/photo-1444418776041-9c7e33cc5a9c',
    title: 'Coffee',
    author: '@nolanissac',
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1533827432537-70133748f5c8',
    title: 'Hats',
    author: '@hjrc33',
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1558642452-9d2a7deb7f62',
    title: 'Honey',
    author: '@arwinneil',
    rows: 2,
    cols: 2,
    featured: true,
  },
  {
    img: 'https://images.unsplash.com/photo-1516802273409-68526ee1bdd6',
    title: 'Basketball',
    author: '@tjdragotta',
  },
  {
    img: 'https://images.unsplash.com/photo-1518756131217-31eb79b20e8f',
    title: 'Fern',
    author: '@katie_wasserman',
  },
  {
    img: 'https://images.unsplash.com/photo-1597645587822-e99fa5d45d25',
    title: 'Mushrooms',
    author: '@silverdalex',
    rows: 2,
    cols: 2,
  },
  {
    img: 'https://images.unsplash.com/photo-1567306301408-9b74779a11af',
    title: 'Tomato basil',
    author: '@shelleypauls',
  },
  {
    img: 'https://images.unsplash.com/photo-1471357674240-e1a485acb3e1',
    title: 'Sea star',
    author: '@peterlaster',
  },
  {
    img: 'https://images.unsplash.com/photo-1589118949245-7d38baf380d6',
    title: 'Bike',
    author: '@southside_customs',
    cols: 2,
  },
];
```
</details>

### Title bar below image (standard)

The title bar can be placed below the image.

```{r}
muiMaterialPage(
  CssBaseline(),
  ImageList(
    sx = list(width = 500, maxWidth = "100%", height = 450),
    lapply(seq_len(nrow(photos)), function(i) {
      ImageListItem(
        photo(photos$img[i], photos$title[i], "w=248&fit=crop"),
        ImageListItemBar(title = photos$title[i], subtitle = tags$span(paste("by:", photos$author[i])), position = "below")
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';
import ImageListItemBar from '@mui/material/ImageListItemBar';

export default function TitlebarBelowImageList() {
  return (
    <ImageList sx={{ width: 500, height: 450 }}>
      {itemData.map((item) => (
        <ImageListItem key={item.img}>
          <img
            srcSet={`${item.img}?w=248&fit=crop&auto=format&dpr=2 2x`}
            src={`${item.img}?w=248&fit=crop&auto=format`}
            alt={item.title}
            loading="lazy"
          />
          <ImageListItemBar
            title={item.title}
            subtitle={<span>by: {item.author}</span>}
            position="below"
          />
        </ImageListItem>
      ))}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1551963831-b3b1ca40c98e',
    title: 'Breakfast',
    author: '@bkristastucchio',
  },
  {
    img: 'https://images.unsplash.com/photo-1551782450-a2132b4ba21d',
    title: 'Burger',
    author: '@rollelflex_graphy726',
  },
  {
    img: 'https://images.unsplash.com/photo-1522770179533-24471fcdba45',
    title: 'Camera',
    author: '@helloimnik',
  },
  {
    img: 'https://images.unsplash.com/photo-1444418776041-9c7e33cc5a9c',
    title: 'Coffee',
    author: '@nolanissac',
  },
  {
    img: 'https://images.unsplash.com/photo-1533827432537-70133748f5c8',
    title: 'Hats',
    author: '@hjrc33',
  },
  {
    img: 'https://images.unsplash.com/photo-1558642452-9d2a7deb7f62',
    title: 'Honey',
    author: '@arwinneil',
  },
  {
    img: 'https://images.unsplash.com/photo-1516802273409-68526ee1bdd6',
    title: 'Basketball',
    author: '@tjdragotta',
  },
  {
    img: 'https://images.unsplash.com/photo-1518756131217-31eb79b20e8f',
    title: 'Fern',
    author: '@katie_wasserman',
  },
  {
    img: 'https://images.unsplash.com/photo-1597645587822-e99fa5d45d25',
    title: 'Mushrooms',
    author: '@silverdalex',
  },
  {
    img: 'https://images.unsplash.com/photo-1567306301408-9b74779a11af',
    title: 'Tomato basil',
    author: '@shelleypauls',
  },
  {
    img: 'https://images.unsplash.com/photo-1471357674240-e1a485acb3e1',
    title: 'Sea star',
    author: '@peterlaster',
  },
  {
    img: 'https://images.unsplash.com/photo-1589118949245-7d38baf380d6',
    title: 'Bike',
    author: '@southside_customs',
  },
];
```
</details>

### Title bar below image (masonry)

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = 500, maxWidth = "100%", height = 450, overflowY = "scroll"),
    ImageList(
      variant = "masonry",
      cols = 3,
      gap = 8,
      lapply(seq_len(nrow(interiors)), function(i) {
        ImageListItem(
          photo(interiors$img[i], interiors$title[i], "w=248&fit=crop"),
          ImageListItemBar(position = "below", title = interiors$author[i])
        )
      })
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Box from '@mui/material/Box';
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';
import ImageListItemBar from '@mui/material/ImageListItemBar';

export default function TitlebarBelowMasonryImageList() {
  return (
    <Box sx={{ width: 500, height: 450, overflowY: 'scroll' }}>
      <ImageList variant="masonry" cols={3} gap={8}>
        {itemData.map((item) => (
          <ImageListItem key={item.img}>
            <img
              srcSet={`${item.img}?w=248&fit=crop&auto=format&dpr=2 2x`}
              src={`${item.img}?w=248&fit=crop&auto=format`}
              alt={item.title}
              loading="lazy"
            />
            <ImageListItemBar position="below" title={item.author} />
          </ImageListItem>
        ))}
      </ImageList>
    </Box>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1549388604-817d15aa0110',
    title: 'Bed',
    author: 'swabdesign',
  },
  {
    img: 'https://images.unsplash.com/photo-1525097487452-6278ff080c31',
    title: 'Books',
    author: 'Pavel Nekoranec',
  },
  {
    img: 'https://images.unsplash.com/photo-1523413651479-597eb2da0ad6',
    title: 'Sink',
    author: 'Charles Deluvio',
  },
  {
    img: 'https://images.unsplash.com/photo-1563298723-dcfebaa392e3',
    title: 'Kitchen',
    author: 'Christian Mackie',
  },
  {
    img: 'https://images.unsplash.com/photo-1588436706487-9d55d73a39e3',
    title: 'Blinds',
    author: 'Darren Richardson',
  },
  {
    img: 'https://images.unsplash.com/photo-1574180045827-681f8a1a9622',
    title: 'Chairs',
    author: 'Taylor Simpson',
  },
  {
    img: 'https://images.unsplash.com/photo-1530731141654-5993c3016c77',
    title: 'Laptop',
    author: 'Ben Kolde',
  },
  {
    img: 'https://images.unsplash.com/photo-1481277542470-605612bd2d61',
    title: 'Doors',
    author: 'Philipp Berndt',
  },
  {
    img: 'https://images.unsplash.com/photo-1517487881594-2787fef5ebf7',
    title: 'Coffee',
    author: 'Jen P.',
  },
  {
    img: 'https://images.unsplash.com/photo-1516455207990-7a41ce80f7ee',
    title: 'Storage',
    author: 'Douglas Sheppard',
  },
  {
    img: 'https://images.unsplash.com/photo-1597262975002-c5c3b14bbd62',
    title: 'Candle',
    author: 'Fi Bell',
  },
  {
    img: 'https://images.unsplash.com/photo-1519710164239-da123dc03ef4',
    title: 'Coffee table',
    author: 'Hutomo Abrianto',
  },
];
```
</details>

## Custom image list

In this example the items have a customized titlebar, positioned at the top and with a custom gradient `titleBackground`.
The secondary action `IconButton` is positioned on the left. The `gap` prop is used to adjust the gap between items.

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  ImageList(
    sx = list(
      width = 500, maxWidth = "100%", height = 450,
      # Promote the list into its own layer in Chrome. This costs memory, but helps keeping high FPS.
      transform = "translateZ(0)"
    ),
    rowHeight = 200,
    gap = 1,
    lapply(seq_len(nrow(photos)), function(i) {
      featured <- photos$featured[i]
      span <- if (featured) 2 else 1
      ImageListItem(
        cols = span,
        rows = span,
        photo(photos$img[i], photos$title[i], sprintf("w=%s&h=%s&fit=crop", 250 * span, 200 * span)),
        ImageListItemBar(
          sx = list(background = "linear-gradient(to bottom, rgba(0,0,0,0.7) 0%, rgba(0,0,0,0.3) 70%, rgba(0,0,0,0) 100%)"),
          title = photos$title[i],
          position = "top",
          actionIcon = IconButton(sx = list(color = "white"), `aria-label` = paste("star", photos$title[i]), Icon("star_border")),
          actionPosition = "left"
        )
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import ImageList from '@mui/material/ImageList';
import ImageListItem from '@mui/material/ImageListItem';
import ImageListItemBar from '@mui/material/ImageListItemBar';
import IconButton from '@mui/material/IconButton';
import StarBorderIcon from '@mui/icons-material/StarBorder';

function srcset(image, width, height, rows = 1, cols = 1) {
  return {
    src: `${image}?w=${width * cols}&h=${height * rows}&fit=crop&auto=format`,
    srcSet: `${image}?w=${width * cols}&h=${
      height * rows
    }&fit=crop&auto=format&dpr=2 2x`,
  };
}

export default function CustomImageList() {
  return (
    <ImageList
      sx={{
        width: 500,
        height: 450,
        // Promote the list into its own layer in Chrome. This costs memory, but helps keeping high FPS.
        transform: 'translateZ(0)',
      }}
      rowHeight={200}
      gap={1}
    >
      {itemData.map((item) => {
        const cols = item.featured ? 2 : 1;
        const rows = item.featured ? 2 : 1;

        return (
          <ImageListItem key={item.img} cols={cols} rows={rows}>
            <img
              {...srcset(item.img, 250, 200, rows, cols)}
              alt={item.title}
              loading="lazy"
            />
            <ImageListItemBar
              sx={{
                background:
                  'linear-gradient(to bottom, rgba(0,0,0,0.7) 0%, ' +
                  'rgba(0,0,0,0.3) 70%, rgba(0,0,0,0) 100%)',
              }}
              title={item.title}
              position="top"
              actionIcon={
                <IconButton
                  sx={{ color: 'white' }}
                  aria-label={`star ${item.title}`}
                >
                  <StarBorderIcon />
                </IconButton>
              }
              actionPosition="left"
            />
          </ImageListItem>
        );
      })}
    </ImageList>
  );
}

const itemData = [
  {
    img: 'https://images.unsplash.com/photo-1551963831-b3b1ca40c98e',
    title: 'Breakfast',
    author: '@bkristastucchio',
    featured: true,
  },
  {
    img: 'https://images.unsplash.com/photo-1551782450-a2132b4ba21d',
    title: 'Burger',
    author: '@rollelflex_graphy726',
  },
  {
    img: 'https://images.unsplash.com/photo-1522770179533-24471fcdba45',
    title: 'Camera',
    author: '@helloimnik',
  },
  {
    img: 'https://images.unsplash.com/photo-1444418776041-9c7e33cc5a9c',
    title: 'Coffee',
    author: '@nolanissac',
  },
  {
    img: 'https://images.unsplash.com/photo-1533827432537-70133748f5c8',
    title: 'Hats',
    author: '@hjrc33',
  },
  {
    img: 'https://images.unsplash.com/photo-1558642452-9d2a7deb7f62',
    title: 'Honey',
    author: '@arwinneil',
    featured: true,
  },
  {
    img: 'https://images.unsplash.com/photo-1516802273409-68526ee1bdd6',
    title: 'Basketball',
    author: '@tjdragotta',
  },
  {
    img: 'https://images.unsplash.com/photo-1518756131217-31eb79b20e8f',
    title: 'Fern',
    author: '@katie_wasserman',
  },
  {
    img: 'https://images.unsplash.com/photo-1597645587822-e99fa5d45d25',
    title: 'Mushrooms',
    author: '@silverdalex',
  },
  {
    img: 'https://images.unsplash.com/photo-1567306301408-9b74779a11af',
    title: 'Tomato basil',
    author: '@shelleypauls',
  },
  {
    img: 'https://images.unsplash.com/photo-1471357674240-e1a485acb3e1',
    title: 'Sea star',
    author: '@peterlaster',
  },
  {
    img: 'https://images.unsplash.com/photo-1589118949245-7d38baf380d6',
    title: 'Bike',
    author: '@southside_customs',
  },
];
```
</details>
