---
title: "Link"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Link}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-link/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
```

# Links

The Link component allows you to easily customize anchor elements with your theme colors and typography styles.

## Basic links

The Link component is built on top of the [Typography](typography.html) component, meaning that you can use its props.
As in the MUI documentation, the demos prevent the navigation with an `onClick`
handler on the wrapper.

```{r}
preventDefault <- JS("(event) => event.preventDefault()")

muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(typography = "body1", "& > :not(style) ~ :not(style)" = list(ml = 2)),
    onClick = preventDefault,
    Link(href = "#", "Link"),
    Link(href = "#", color = "inherit", 'color="inherit"'),
    Link(href = "#", variant = "body2", 'variant="body2"')
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Link from '@mui/material/Link';

const preventDefault = (event) => event.preventDefault();

export default function Links() {
  return (
    <Box
      sx={{
        typography: 'body1',
        '& > :not(style) ~ :not(style)': {
          ml: 2,
        },
      }}
      onClick={preventDefault}
    >
      <Link href="#">Link</Link>
      <Link href="#" color="inherit">
        {'color="inherit"'}
      </Link>
      <Link href="#" variant="body2">
        {'variant="body2"'}
      </Link>
    </Box>
  );
}
```
</details>

However, the Link component has some different default props than the Typography component:

- `color = "primary"` as the link needs to stand out.
- `variant = "inherit"` as the link will, most of the time, be used as a child of a Typography component.

## Underline

The `underline` prop can be used to set the underline behavior. The default is `always`.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(
      display = "flex", flexWrap = "wrap", justifyContent = "center", typography = "body1",
      "& > :not(style) ~ :not(style)" = list(ml = 2)
    ),
    onClick = preventDefault,
    Link(href = "#", underline = "none", 'underline="none"'),
    Link(href = "#", underline = "hover", 'underline="hover"'),
    Link(href = "#", underline = "always", 'underline="always"')
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Link from '@mui/material/Link';

const preventDefault = (event) => event.preventDefault();

export default function UnderlineLink() {
  return (
    <Box
      sx={{
        display: 'flex',
        flexWrap: 'wrap',
        justifyContent: 'center',
        typography: 'body1',
        '& > :not(style) ~ :not(style)': {
          ml: 2,
        },
      }}
      onClick={preventDefault}
    >
      <Link href="#" underline="none">
        {'underline="none"'}
      </Link>
      <Link href="#" underline="hover">
        {'underline="hover"'}
      </Link>
      <Link href="#" underline="always">
        {'underline="always"'}
      </Link>
    </Box>
  );
}
```
</details>

## Third-party routing library

One frequent use case is to perform navigation on the client only, without an HTTP round-trip to the server.
With a hash router such as `createHashRouter()` of
[reactRouter](https://felixluginbuhl.com/reactRouter/), a regular `href` that
starts with `#/` is enough: `Link(href = "#/about", "About")` navigates to the
`about` route. See [Using a router](routing.html).

<div class="alert alert-warning">
reactRouter also exports a `Link()` function, for its own links. When both
packages are attached, the last one attached masks the other: call
`muiMaterial::Link()` explicitly, or attach muiMaterial after reactRouter.
</div>

## Accessibility

(WAI-ARIA: <https://www.w3.org/WAI/ARIA/apg/patterns/link/>)

- When providing the content for the link, avoid generic descriptions like "click here" or "go to".
  Instead, use [specific descriptions](https://developers.google.com/web/tools/lighthouse/audits/descriptive-link-text).
- For the best user experience, links should stand out from the text on the page. For instance, you can keep the default `underline = "always"` behavior.
- If a link doesn't have a meaningful href, [it should be rendered using a `<button>` element](https://github.com/jsx-eslint/eslint-plugin-jsx-a11y/blob/HEAD/docs/rules/anchor-is-valid.md).
  The demo below illustrates how to properly link with a `<button>`:

```{r}
muiMaterialPage(
  CssBaseline(),
  Link(component = "button", variant = "body2", onClick = JS("() => { console.info(\"I'm a button.\"); }"), "Button Link")
)
```

<details><summary>JS code</summary>
```jsx
import Link from '@mui/material/Link';

export default function ButtonLink() {
  return (
    <Link
      component="button"
      variant="body2"
      onClick={() => {
        console.info("I'm a button.");
      }}
    >
      Button Link
    </Link>
  );
}
```
</details>

In a Shiny app, send the click to the server with
`onClick = triggerEvent("link_clicked")`.

### Keyboard accessibility

- Interactive elements should receive focus in a coherent order when the user presses the <kbd class="key">Tab</kbd> key.
- Users should be able to open a link by pressing <kbd class="key">Enter</kbd>.

### Screen reader accessibility

- When a link receives focus, screen readers should announce a descriptive link name.
  If the link opens in a new window or browser tab, add an [`aria-label`](https://www.w3.org/WAI/WCAG22/Techniques/aria/ARIA8) to inform screen reader users, for example, "To learn more, visit the About page which opens in a new window."
