---
title: "Modal"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Modal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-modal/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Modal

The modal component provides a solid foundation for creating dialogs, popovers, lightboxes, or whatever else.

The component renders its `children` node in front of a backdrop component.
The `Modal` offers important features:

- 💄 Manages modal stacking when one-at-a-time just isn't enough.
- 🔐 Creates a backdrop, for disabling interaction below the modal.
- 🔐 It disables scrolling of the page content while open.
- ♿️ It properly manages focus; moving to the modal content,
  and keeping it there until the modal is closed.
- ♿️ Adds the appropriate ARIA roles automatically.

<div class="alert alert-info">
The term "modal" is sometimes used to mean "dialog", but this is a misnomer.
A modal window describes parts of a UI.
An element is considered modal if [it blocks interaction with the rest of the application](https://en.wikipedia.org/wiki/Modal_window).
</div>

If you are creating a modal dialog, you probably want to use the [Dialog](dialog.html) component rather than directly using Modal.
Modal is a lower-level construct that is leveraged by the following components:

- [Dialog](dialog.html)
- [Drawer](drawer.html)
- [Menu](menu.html)
- [Popover](popover.html)

As on the [Dialog](dialog.html) page, the live demos keep the open state in the
URL with [reactRouter](https://felixluginbuhl.com/reactRouter/): each modal is
a route, a link opens it, and `onClose` navigates back to `#/`.

```{r}
closeModal <- JS("() => { window.location.hash = '/'; }")

modalStyle <- list(
  position = "absolute",
  top = "50%",
  left = "50%",
  transform = "translate(-50%, -50%)",
  width = 400,
  bgcolor = "background.paper",
  border = "2px solid #000",
  boxShadow = 24,
  p = 4
)
```

## Basic modal

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/basic-modal", "Open modal"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "basic-modal",
        element = Modal(
          open = TRUE,
          onClose = closeModal,
          `aria-labelledby` = "modal-modal-title",
          `aria-describedby` = "modal-modal-description",
          Box(
            sx = modalStyle,
            Typography(id = "modal-modal-title", variant = "h6", component = "h2", "Text in a modal"),
            Typography(
              id = "modal-modal-description",
              sx = list(mt = 2),
              "Duis mollis, est non commodo luctus, nisi erat porttitor ligula."
            )
          )
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Button from '@mui/material/Button';
import Typography from '@mui/material/Typography';
import Modal from '@mui/material/Modal';

const style = {
  position: 'absolute',
  top: '50%',
  left: '50%',
  transform: 'translate(-50%, -50%)',
  width: 400,
  bgcolor: 'background.paper',
  border: '2px solid #000',
  boxShadow: 24,
  p: 4,
};

export default function BasicModal() {
  const [open, setOpen] = React.useState(false);
  const handleOpen = () => setOpen(true);
  const handleClose = () => setOpen(false);

  return (
    <div>
      <Button onClick={handleOpen}>Open modal</Button>
      <Modal
        open={open}
        onClose={handleClose}
        aria-labelledby="modal-modal-title"
        aria-describedby="modal-modal-description"
      >
        <Box sx={style}>
          <Typography id="modal-modal-title" variant="h6" component="h2">
            Text in a modal
          </Typography>
          <Typography id="modal-modal-description" sx={{ mt: 2 }}>
            Duis mollis, est non commodo luctus, nisi erat porttitor ligula.
          </Typography>
        </Box>
      </Modal>
    </div>
  );
}
```
</details>

Notice that you can disable the outline (often blue or gold) with the `outline = 0` CSS property.

## Nested modal

Modals can be nested, for example a select within a dialog, but stacking of more than two modals, or any two modals with a backdrop is discouraged.

If you need to nest modals, use the `hideBackdrop` prop on nested Modals to avoid stacking multiple backdrops, which can obscure the active modal or affect interactions.

Nested routes map naturally onto nested modals: the child modal is the route
`#/nested-modal/child`, rendered through `Outlet()` inside the parent modal.

```{r}
nestedStyle <- modifyList(modalStyle, list(p = NULL, pt = 2, px = 4, pb = 3))

muiMaterialPage(
  CssBaseline(),
  Button(href = "#/nested-modal", "Open modal"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "nested-modal",
        element = Modal(
          open = TRUE,
          onClose = closeModal,
          `aria-labelledby` = "parent-modal-title",
          `aria-describedby` = "parent-modal-description",
          Box(
            sx = modifyList(nestedStyle, list(width = 400)),
            tags$h2(id = "parent-modal-title", "Text in a modal"),
            tags$p(
              id = "parent-modal-description",
              "Duis mollis, est non commodo luctus, nisi erat porttitor ligula."
            ),
            Button(href = "#/nested-modal/child", "Open Child Modal"),
            Outlet()
          )
        ),
        Route(
          path = "child",
          element = Modal(
            open = TRUE,
            onClose = JS("() => { window.location.hash = '/nested-modal'; }"),
            `aria-labelledby` = "child-modal-title",
            `aria-describedby` = "child-modal-description",
            Box(
              sx = modifyList(nestedStyle, list(width = 200)),
              tags$h2(id = "child-modal-title", "Text in a child modal"),
              tags$p(
                id = "child-modal-description",
                "Lorem ipsum, dolor sit amet consectetur adipisicing elit."
              ),
              Button(href = "#/nested-modal", "Close Child Modal")
            )
          )
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Modal from '@mui/material/Modal';
import Button from '@mui/material/Button';

const style = {
  position: 'absolute',
  top: '50%',
  left: '50%',
  transform: 'translate(-50%, -50%)',
  width: 400,
  bgcolor: 'background.paper',
  border: '2px solid #000',
  boxShadow: 24,
  pt: 2,
  px: 4,
  pb: 3,
};

function ChildModal() {
  const [open, setOpen] = React.useState(false);
  const handleOpen = () => {
    setOpen(true);
  };
  const handleClose = () => {
    setOpen(false);
  };

  return (
    <React.Fragment>
      <Button onClick={handleOpen}>Open Child Modal</Button>
      <Modal
        open={open}
        onClose={handleClose}
        aria-labelledby="child-modal-title"
        aria-describedby="child-modal-description"
      >
        <Box sx={{ ...style, width: 200 }}>
          <h2 id="child-modal-title">Text in a child modal</h2>
          <p id="child-modal-description">
            Lorem ipsum, dolor sit amet consectetur adipisicing elit.
          </p>
          <Button onClick={handleClose}>Close Child Modal</Button>
        </Box>
      </Modal>
    </React.Fragment>
  );
}

export default function NestedModal() {
  const [open, setOpen] = React.useState(false);
  const handleOpen = () => {
    setOpen(true);
  };
  const handleClose = () => {
    setOpen(false);
  };

  return (
    <div>
      <Button onClick={handleOpen}>Open modal</Button>
      <Modal
        open={open}
        onClose={handleClose}
        aria-labelledby="parent-modal-title"
        aria-describedby="parent-modal-description"
      >
        <Box sx={{ ...style, width: 400 }}>
          <h2 id="parent-modal-title">Text in a modal</h2>
          <p id="parent-modal-description">
            Duis mollis, est non commodo luctus, nisi erat porttitor ligula.
          </p>
          <ChildModal />
        </Box>
      </Modal>
    </div>
  );
}
```
</details>

## Transitions

The open/close state of the modal can be animated with a transition component.
This component should respect the following conditions:

- Be a direct child descendent of the modal.
- Have an `in` prop. This corresponds to the open/close state.
- Call the `onEnter` callback prop when the enter transition starts.
- Call the `onExited` callback prop when the exit transition is completed.
  These two callbacks allow the modal to unmount the child content when closed and fully transitioned.

Modal has built-in support for [react-transition-group](https://github.com/reactjs/react-transition-group).
The transition components of muiMaterial (`Fade()`, `Grow()`, `Slide()`,
`Zoom()`, see [Transitions](transitions.html)) satisfy these conditions.

Leaving a route unmounts its element at once, so in these demos only the
opening is animated.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/transitions-modal", "Open modal"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "transitions-modal",
        element = Modal(
          `aria-labelledby` = "transition-modal-title",
          `aria-describedby` = "transition-modal-description",
          open = TRUE,
          onClose = closeModal,
          closeAfterTransition = TRUE,
          slots = list(backdrop = JS("jsmodule['@mui/material'].Backdrop")),
          slotProps = list(backdrop = list(timeout = 500)),
          Fade(
            `in` = TRUE,
            Box(
              sx = modalStyle,
              Typography(id = "transition-modal-title", variant = "h6", component = "h2", "Text in a modal"),
              Typography(
                id = "transition-modal-description",
                sx = list(mt = 2),
                "Duis mollis, est non commodo luctus, nisi erat porttitor ligula."
              )
            )
          )
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Backdrop from '@mui/material/Backdrop';
import Box from '@mui/material/Box';
import Modal from '@mui/material/Modal';
import Fade from '@mui/material/Fade';
import Button from '@mui/material/Button';
import Typography from '@mui/material/Typography';

const style = {
  position: 'absolute',
  top: '50%',
  left: '50%',
  transform: 'translate(-50%, -50%)',
  width: 400,
  bgcolor: 'background.paper',
  border: '2px solid #000',
  boxShadow: 24,
  p: 4,
};

export default function TransitionsModal() {
  const [open, setOpen] = React.useState(false);
  const handleOpen = () => setOpen(true);
  const handleClose = () => setOpen(false);

  return (
    <div>
      <Button onClick={handleOpen}>Open modal</Button>
      <Modal
        aria-labelledby="transition-modal-title"
        aria-describedby="transition-modal-description"
        open={open}
        onClose={handleClose}
        closeAfterTransition
        slots={{ backdrop: Backdrop }}
        slotProps={{
          backdrop: {
            timeout: 500,
          },
        }}
      >
        <Fade in={open}>
          <Box sx={style}>
            <Typography id="transition-modal-title" variant="h6" component="h2">
              Text in a modal
            </Typography>
            <Typography id="transition-modal-description" sx={{ mt: 2 }}>
              Duis mollis, est non commodo luctus, nisi erat porttitor ligula.
            </Typography>
          </Box>
        </Fade>
      </Modal>
    </div>
  );
}
```
</details>

The MUI documentation also shows an animation made with
[react-spring](https://github.com/pmndrs/react-spring). That library is not
bundled with muiMaterial.

## Performance

The content of modal is unmounted when closed.
If you need to make the content available to search engines or render expensive component trees inside your modal while optimizing for interaction responsiveness
it might be a good idea to change this default behavior by enabling the `keepMounted` prop:

```r
Modal(keepMounted = TRUE)
```

A route unmounts its element when you leave it, so `keepMounted` only helps
when the Modal itself stays in the page. `Modal.triggerId()` does exactly that:
it is always rendered and opens when the element with the matching `id` is
clicked.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(id = "open-keep-mounted-modal", "Open modal"),
  Modal.triggerId(
    triggerId = "open-keep-mounted-modal",
    keepMounted = TRUE,
    `aria-labelledby` = "keep-mounted-modal-title",
    `aria-describedby` = "keep-mounted-modal-description",
    Box(
      sx = modalStyle,
      Typography(id = "keep-mounted-modal-title", variant = "h6", component = "h2", "Text in a modal"),
      Typography(
        id = "keep-mounted-modal-description",
        sx = list(mt = 2),
        "Duis mollis, est non commodo luctus, nisi erat porttitor ligula."
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Modal from '@mui/material/Modal';
import Button from '@mui/material/Button';
import Typography from '@mui/material/Typography';

const style = {
  position: 'absolute',
  top: '50%',
  left: '50%',
  transform: 'translate(-50%, -50%)',
  width: 400,
  bgcolor: 'background.paper',
  border: '2px solid #000',
  boxShadow: 24,
  p: 4,
};

export default function KeepMountedModal() {
  const [open, setOpen] = React.useState(false);
  const handleOpen = () => setOpen(true);
  const handleClose = () => setOpen(false);

  return (
    <div>
      <Button onClick={handleOpen}>Open modal</Button>
      <Modal
        keepMounted
        open={open}
        onClose={handleClose}
        aria-labelledby="keep-mounted-modal-title"
        aria-describedby="keep-mounted-modal-description"
      >
        <Box sx={style}>
          <Typography id="keep-mounted-modal-title" variant="h6" component="h2">
            Text in a modal
          </Typography>
          <Typography id="keep-mounted-modal-description" sx={{ mt: 2 }}>
            Duis mollis, est non commodo luctus, nisi erat porttitor ligula.
          </Typography>
        </Box>
      </Modal>
    </div>
  );
}
```
</details>

As with any performance optimization, this is not a silver bullet.
Be sure to identify bottlenecks first, and then try out these optimization strategies.

## Server-side modal

React [doesn't support](https://github.com/facebook/react/issues/13097) the [`createPortal()`](https://react.dev/reference/react-dom/createPortal) API on the server.
In order to display the modal, you need to disable the portal feature with the `disablePortal` prop.
The modal is then rendered in place, inside its parent:

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(height = 300, flexGrow = 1, minWidth = 300, transform = "translateZ(0)"),
    Modal(
      disablePortal = TRUE,
      disableEnforceFocus = TRUE,
      disableAutoFocus = TRUE,
      open = TRUE,
      `aria-labelledby` = "server-modal-title",
      `aria-describedby` = "server-modal-description",
      sx = list(display = "flex", p = 1, alignItems = "center", justifyContent = "center"),
      Box(
        sx = list(
          position = "relative",
          width = 400,
          bgcolor = "background.paper",
          border = "2px solid #000",
          boxShadow = 5,
          p = 4
        ),
        Typography(id = "server-modal-title", variant = "h6", component = "h2", "Server-side modal"),
        Typography(
          id = "server-modal-description",
          sx = list(pt = 2),
          "If you disable JavaScript, you will still see me."
        )
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Modal from '@mui/material/Modal';
import Typography from '@mui/material/Typography';
import Box from '@mui/material/Box';

export default function ServerModal() {
  const rootRef = React.useRef(null);

  return (
    <Box
      sx={{
        height: 300,
        flexGrow: 1,
        minWidth: 300,
        transform: 'translateZ(0)',
      }}
      ref={rootRef}
    >
      <Modal
        disablePortal
        disableEnforceFocus
        disableAutoFocus
        open
        aria-labelledby="server-modal-title"
        aria-describedby="server-modal-description"
        sx={{
          display: 'flex',
          p: 1,
          alignItems: 'center',
          justifyContent: 'center',
        }}
        container={() => rootRef.current}
      >
        <Box
          sx={(theme) => ({
            position: 'relative',
            width: 400,
            bgcolor: 'background.paper',
            border: '2px solid #000',
            boxShadow: theme.shadows[5],
            p: 4,
          })}
        >
          <Typography id="server-modal-title" variant="h6" component="h2">
            Server-side modal
          </Typography>
          <Typography id="server-modal-description" sx={{ pt: 2 }}>
            If you disable JavaScript, you will still see me.
          </Typography>
        </Box>
      </Modal>
    </Box>
  );
}
```
</details>

## Limitations

### Focus trap

The modal moves the focus back to the body of the component if the focus tries to escape it.

This is done for accessibility purposes. However, it might create issues.
In the event the users need to interact with another part of the page, for example with a chatbot window, you can disable the behavior:

```r
Modal(disableEnforceFocus = TRUE)
```

## Accessibility

(WAI-ARIA: <https://www.w3.org/WAI/ARIA/apg/patterns/dialog-modal/>)

- Be sure to add `aria-labelledby = "id..."`, referencing the modal title, to the `Modal`.
  Additionally, you may give a description of your modal with the `aria-describedby = "id..."` prop on the `Modal`.

  ```r
  Modal(
    `aria-labelledby` = "modal-title",
    `aria-describedby` = "modal-description",
    tags$h2(id = "modal-title", "My Title"),
    tags$p(id = "modal-description", "My Description")
  )
  ```

- The [WAI-ARIA Authoring Practices](https://www.w3.org/WAI/ARIA/apg/patterns/dialog-modal/examples/dialog/) can help you set the initial focus on the most relevant element, based on your modal content.
- Keep in mind that a "modal window" overlays on either the primary window or another modal window. Windows under a modal are **inert**. That is, users cannot interact with content outside an active modal window. This might create [conflicting behaviors](#focus-trap).

## Modals in Shiny apps

The router pattern above works unchanged in Shiny. muiMaterial also has:

- `Modal.triggerId()`, shown in [Performance](#performance), where the browser
  handles the open/close state. See [Overlays with `.triggerId`](triggerid.html).
- `Modal.shinyInput()`, where the server opens and closes the modal with
  `updateModal.shinyInput()`:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Button.shinyInput("open", "Open modal"),
  Modal.shinyInput(
    inputId = "modal",
    open = FALSE,
    onClose = triggerEvent("modal_close"),
    Box(
      sx = list(
        position = "absolute", top = "50%", left = "50%",
        transform = "translate(-50%, -50%)", width = 400,
        bgcolor = "background.paper", boxShadow = 24, p = 4
      ),
      Typography("Text in a modal", variant = "h6", component = "h2")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$open, updateModal.shinyInput(inputId = "modal", open = TRUE))
  observeEvent(input$modal_close, updateModal.shinyInput(inputId = "modal", open = FALSE))
}

shinyApp(ui, server)
```

Run the bundled version with `muiMaterialExample("Modal")`.
