---
title: "Pagination"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Pagination}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-pagination/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Pagination

The Pagination component enables the user to select a specific page from a range of pages.

## Basic pagination

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 10),
    Pagination(count = 10, color = "primary"),
    Pagination(count = 10, color = "secondary"),
    Pagination(count = 10, disabled = TRUE)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function BasicPagination() {
  return (
    <Stack spacing={2}>
      <Pagination count={10} />
      <Pagination count={10} color="primary" />
      <Pagination count={10} color="secondary" />
      <Pagination count={10} disabled />
    </Stack>
  );
}
```
</details>

## Outlined pagination

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 10, variant = "outlined"),
    Pagination(count = 10, variant = "outlined", color = "primary"),
    Pagination(count = 10, variant = "outlined", color = "secondary"),
    Pagination(count = 10, variant = "outlined", disabled = TRUE)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationOutlined() {
  return (
    <Stack spacing={2}>
      <Pagination count={10} variant="outlined" />
      <Pagination count={10} variant="outlined" color="primary" />
      <Pagination count={10} variant="outlined" color="secondary" />
      <Pagination count={10} variant="outlined" disabled />
    </Stack>
  );
}
```
</details>

## Rounded pagination

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 10, shape = "rounded"),
    Pagination(count = 10, variant = "outlined", shape = "rounded")
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationRounded() {
  return (
    <Stack spacing={2}>
      <Pagination count={10} shape="rounded" />
      <Pagination count={10} variant="outlined" shape="rounded" />
    </Stack>
  );
}
```
</details>

## Pagination size

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 10, size = "small"),
    Pagination(count = 10),
    Pagination(count = 10, size = "large")
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationSize() {
  return (
    <Stack spacing={2}>
      <Pagination count={10} size="small" />
      <Pagination count={10} />
      <Pagination count={10} size="large" />
    </Stack>
  );
}
```
</details>

## Buttons

You can optionally enable first-page and last-page buttons, or disable the previous-page and next-page buttons.

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 10, showFirstButton = TRUE, showLastButton = TRUE),
    Pagination(count = 10, hidePrevButton = TRUE, hideNextButton = TRUE)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationButtons() {
  return (
    <Stack spacing={2}>
      <Pagination count={10} showFirstButton showLastButton />
      <Pagination count={10} hidePrevButton hideNextButton />
    </Stack>
  );
}
```
</details>

## Custom icons

It's possible to customize the control icons.
`renderItem` is a function that returns an element for each item, so it is
written in JavaScript with `JS()`. `jsmodule` gives access to the React and MUI
modules bundled with muiMaterial:

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(
      count = 10,
      renderItem = JS(
        "(item) => {
          const h = jsmodule['react'].createElement;
          const { PaginationItem, Icon } = jsmodule['@mui/material'];
          return h(PaginationItem, {
            ...item,
            slots: {
              previous: () => h(Icon, null, 'arrow_back'),
              next: () => h(Icon, null, 'arrow_forward')
            }
          });
        }"
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import PaginationItem from '@mui/material/PaginationItem';
import Stack from '@mui/material/Stack';
import ArrowBackIcon from '@mui/icons-material/ArrowBack';
import ArrowForwardIcon from '@mui/icons-material/ArrowForward';

export default function CustomIcons() {
  return (
    <Stack spacing={2}>
      <Pagination
        count={10}
        renderItem={(item) => (
          <PaginationItem
            slots={{ previous: ArrowBackIcon, next: ArrowForwardIcon }}
            {...item}
          />
        )}
      />
    </Stack>
  );
}
```
</details>

## Pagination ranges

You can specify how many digits to display either side of current page with the `siblingCount` prop, and adjacent to the start and end page number with the `boundaryCount` prop.

```{r}
muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    Pagination(count = 11, defaultPage = 6, siblingCount = 0),
    Pagination(count = 11, defaultPage = 6), # Default ranges
    Pagination(count = 11, defaultPage = 6, siblingCount = 0, boundaryCount = 2),
    Pagination(count = 11, defaultPage = 6, boundaryCount = 2)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationRanges() {
  return (
    <Stack spacing={2}>
      <Pagination count={11} defaultPage={6} siblingCount={0} />
      <Pagination count={11} defaultPage={6} /> {/* Default ranges */}
      <Pagination count={11} defaultPage={6} siblingCount={0} boundaryCount={2} />
      <Pagination count={11} defaultPage={6} boundaryCount={2} />
    </Stack>
  );
}
```
</details>

## Controlled pagination

In a Shiny app, `Pagination.shinyInput()` sends the selected page to
`input[[inputId]]`, and `updatePagination.shinyInput()` changes it from the
server:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    sx = list(p = 2),
    uiOutput("page"),
    Pagination.shinyInput("pagination", count = 10, value = 1)
  )
)

server <- function(input, output, session) {
  output$page <- renderUI(Typography(paste("Page:", input$pagination)))
}

shinyApp(ui, server)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Typography from '@mui/material/Typography';
import Pagination from '@mui/material/Pagination';
import Stack from '@mui/material/Stack';

export default function PaginationControlled() {
  const [page, setPage] = React.useState(1);
  const handleChange = (event, value) => {
    setPage(value);
  };

  return (
    <Stack spacing={2}>
      <Typography>Page: {page}</Typography>
      <Pagination count={10} page={page} onChange={handleChange} />
    </Stack>
  );
}
```
</details>

Run the bundled version with `muiMaterialExample("Pagination")`.

## Router integration

Pagination can also be kept in the URL, so that each page can be bookmarked.
The MUI demo uses React Router; in R, the same is done with
[reactRouter](https://felixluginbuhl.com/reactRouter/). Each page is a route
(`#/inbox/<page>`), and `onChange` navigates to the selected page:

```{r}
inbox <- function(page) {
  Stack(
    spacing = 2,
    Typography(paste("Page:", page)),
    Pagination(
      page = page,
      count = 10,
      onChange = JS("(event, value) => { window.location.hash = '/inbox/' + value; }")
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(1:10, function(page) Route(path = paste0("inbox/", page), element = inbox(page))),
      Route(path = "*", element = inbox(1))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import { Link, MemoryRouter, Route, Routes, useLocation } from 'react-router';
import Pagination from '@mui/material/Pagination';
import PaginationItem from '@mui/material/PaginationItem';

function Content() {
  const location = useLocation();
  const query = new URLSearchParams(location.search);
  const page = parseInt(query.get('page') || '1', 10);
  return (
    <Pagination
      page={page}
      count={10}
      renderItem={(item) => (
        <PaginationItem
          component={Link}
          to={`/inbox${item.page === 1 ? '' : `?page=${item.page}`}`}
          {...item}
        />
      )}
    />
  );
}

export default function PaginationLink() {
  return (
    <MemoryRouter initialEntries={['/inbox']} initialIndex={0}>
      <Routes>
        <Route path="*" element={<Content />} />
      </Routes>
    </MemoryRouter>
  );
}
```
</details>

## `usePagination`

For advanced customization use cases, React exposes a headless `usePagination()`
hook. It accepts almost the same options as the Pagination component minus all
the props related to the rendering of JSX. Hooks are not available from R; use
`renderItem` (see [Custom icons](#custom-icons)) to customize the items instead.

## Table pagination

The `Pagination` component was designed to paginate a list of arbitrary items when infinite loading isn't used.
It's preferred in contexts where SEO is important, for instance, a blog.

For the pagination of a large set of tabular data, you should use the `TablePagination` component.

```{r}
muiMaterialPage(
  CssBaseline(),
  TablePagination(component = "div", count = 100, page = 2, rowsPerPage = 10, onPageChange = JS("() => {}"))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import TablePagination from '@mui/material/TablePagination';

export default function TablePaginationDemo() {
  const [page, setPage] = React.useState(2);
  const [rowsPerPage, setRowsPerPage] = React.useState(10);

  const handleChangePage = (event, newPage) => {
    setPage(newPage);
  };

  const handleChangeRowsPerPage = (event) => {
    setRowsPerPage(parseInt(event.target.value, 10));
    setPage(0);
  };

  return (
    <TablePagination
      component="div"
      count={100}
      page={page}
      onPageChange={handleChangePage}
      rowsPerPage={rowsPerPage}
      onRowsPerPageChange={handleChangeRowsPerPage}
    />
  );
}
```
</details>

<div class="alert alert-warning">
Note that the `Pagination` page prop starts at 1 to match the requirement of including the value in the URL, while the `TablePagination` page prop starts at 0 to match the requirement of zero-based JavaScript arrays that come with rendering a lot of tabular data.
</div>

For a working paginated table, see the [sticky header table](table.html#sticky-header).

## Accessibility

### ARIA

The root node has a role of "navigation" and aria-label "pagination navigation" by default. The page items have an aria-label that identifies the purpose of the item ("go to first page", "go to previous page", "go to page 1" etc.).
You can override these using the `getItemAriaLabel` prop.

### Keyboard

The pagination items are in tab order, with a tabindex of "0".
