---
title: "Popover"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Popover}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-popover/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Popover

A Popover can be used to display some content on top of another.

Things to know when using the `Popover` component:

- The component is built on top of the [`Modal`](modal.html) component.
- `Popover` blocks scrolling and dismisses on click-away by default, unlike [`Popper`](popper.html).

In React, the element a popover is anchored to is kept in a `useState()` hook.
`Popover.triggerId()` does this for you: give the button an `id` and pass the
same id as `triggerId`. The popover is anchored to that button, and closes on a
click outside or on Escape. See [Overlays with `.triggerId`](triggerid.html).

## Basic Popover

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(id = "simple-popover-button", `aria-describedby` = "simple-popover", variant = "contained", "Open Popover"),
  Popover.triggerId(
    triggerId = "simple-popover-button",
    id = "simple-popover",
    anchorOrigin = list(vertical = "bottom", horizontal = "left"),
    Typography(sx = list(p = 2), "The content of the Popover.")
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Popover from '@mui/material/Popover';
import Typography from '@mui/material/Typography';
import Button from '@mui/material/Button';

export default function BasicPopover() {
  const [anchorEl, setAnchorEl] = React.useState(null);

  const handleClick = (event) => {
    setAnchorEl(event.currentTarget);
  };

  const handleClose = () => {
    setAnchorEl(null);
  };

  const open = Boolean(anchorEl);
  const id = open ? 'simple-popover' : undefined;

  return (
    <div>
      <Button aria-describedby={id} variant="contained" onClick={handleClick}>
        Open Popover
      </Button>
      <Popover
        id={id}
        open={open}
        anchorEl={anchorEl}
        onClose={handleClose}
        anchorOrigin={{
          vertical: 'bottom',
          horizontal: 'left',
        }}
      >
        <Typography sx={{ p: 2 }}>The content of the Popover.</Typography>
      </Popover>
    </div>
  );
}
```
</details>

## Anchor playground

`anchorOrigin` sets the point of the anchor (the button) where the popover
attaches, and `transformOrigin` sets the point of the popover that is
attached there. Each takes a `vertical` (`"top"`, `"center"`, `"bottom"`) and a
`horizontal` (`"left"`, `"center"`, `"right"`) position.

The MUI documentation has an interactive playground with radio buttons. Below,
each button opens a popover with a different combination:

```{r}
origins <- list(
  "below, aligned left" = list(
    anchorOrigin = list(vertical = "bottom", horizontal = "left"),
    transformOrigin = list(vertical = "top", horizontal = "left")
  ),
  "below, centered" = list(
    anchorOrigin = list(vertical = "bottom", horizontal = "center"),
    transformOrigin = list(vertical = "top", horizontal = "center")
  ),
  "on the right" = list(
    anchorOrigin = list(vertical = "center", horizontal = "right"),
    transformOrigin = list(vertical = "center", horizontal = "left")
  ),
  "above, aligned right" = list(
    anchorOrigin = list(vertical = "top", horizontal = "right"),
    transformOrigin = list(vertical = "bottom", horizontal = "right")
  )
)

muiMaterialPage(
  CssBaseline(),
  Stack(
    direction = "row",
    spacing = 2,
    sx = list(py = 8, justifyContent = "center", flexWrap = "wrap"),
    lapply(seq_along(origins), function(i) {
      tagList(
        Button(id = paste0("anchor-button-", i), variant = "outlined", names(origins)[i]),
        Popover.triggerId(
          triggerId = paste0("anchor-button-", i),
          anchorOrigin = origins[[i]]$anchorOrigin,
          transformOrigin = origins[[i]]$transformOrigin,
          Typography(sx = list(p = 2), "The content of the Popover.")
        )
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';
import FormControlLabel from '@mui/material/FormControlLabel';
import RadioGroup from '@mui/material/RadioGroup';
import Radio from '@mui/material/Radio';
import Grid from '@mui/material/Grid';
import { green } from '@mui/material/colors';
import Typography from '@mui/material/Typography';
import Box from '@mui/material/Box';
import Button from '@mui/material/Button';
import Popover from '@mui/material/Popover';
import Input from '@mui/material/Input';
import InputLabel from '@mui/material/InputLabel';
import { HighlightedCode } from '@mui/internal-core-docs/HighlightedCode';

const inlineStyles = {
  anchorVertical: {
    top: {
      top: -5,
    },
    center: {
      top: 'calc(50% - 5px)',
    },
    bottom: {
      bottom: -5,
    },
  },
  anchorHorizontal: {
    left: {
      left: -5,
    },
    center: {
      left: 'calc(50% - 5px)',
    },
    right: {
      right: -5,
    },
  },
};

function AnchorPlayground() {
  const anchorRef = React.useRef();

  const [state, setState] = React.useState({
    open: false,
    anchorOriginVertical: 'top',
    anchorOriginHorizontal: 'left',
    transformOriginVertical: 'top',
    transformOriginHorizontal: 'left',
    positionTop: 200, // Just so the popover can be spotted more easily
    positionLeft: 400, // Same as above
    anchorReference: 'anchorEl',
  });

  const {
    open,
    anchorOriginVertical,
    anchorOriginHorizontal,
    transformOriginVertical,
    transformOriginHorizontal,
    positionTop,
    positionLeft,
    anchorReference,
  } = state;

  const handleChange = (event) => {
    setState({
      ...state,
      [event.target.name]: event.target.value,
    });
  };

  const handleNumberInputChange = (key) => (event) => {
    setState({
      ...state,
      [key]: parseInt(event.target.value, 10),
    });
  };

  const handleClickButton = () => {
    setState({
      ...state,
      open: true,
    });
  };

  const handleClose = () => {
    setState({
      ...state,
      open: false,
    });
  };

  let mode = '';

  if (anchorReference === 'anchorPosition') {
    mode = `
  anchorReference="${anchorReference}"
  anchorPosition={{ top: ${positionTop}, left: ${positionLeft} }}`;
  }

  const jsx = `
<Popover ${mode}
  anchorOrigin={{
    vertical: '${anchorOriginVertical}',
    horizontal: '${anchorOriginHorizontal}',
  }}
  transformOrigin={{
    vertical: '${transformOriginVertical}',
    horizontal: '${transformOriginHorizontal}',
  }}
>
  The content of the Popover.
</Popover>
`;

  const radioAnchorClasses = {
    color: green[600],
    '&.Mui-checked': {
      color: green[500],
    },
  };

  return (
    <div>
      <Grid container sx={{ justifyContent: 'center' }}>
        <Grid sx={{ position: 'relative', mb: 4 }}>
          <Button ref={anchorRef} variant="contained" onClick={handleClickButton}>
            Open Popover
          </Button>
          {anchorReference === 'anchorEl' && (
            <Box
              sx={{
                bgcolor: green[500],
                width: 10,
                height: 10,
                borderRadius: '50%',
                position: 'absolute',
              }}
              style={{
                ...inlineStyles.anchorVertical[anchorOriginVertical],
                ...inlineStyles.anchorHorizontal[anchorOriginHorizontal],
              }}
            />
          )}
        </Grid>
      </Grid>
      <Popover
        open={open}
        anchorEl={anchorRef.current}
        anchorReference={anchorReference}
        anchorPosition={{
          top: positionTop,
          left: positionLeft,
        }}
        onClose={handleClose}
        anchorOrigin={{
          vertical: anchorOriginVertical,
          horizontal: anchorOriginHorizontal,
        }}
        transformOrigin={{
          vertical: transformOriginVertical,
          horizontal: transformOriginHorizontal,
        }}
      >
        <Typography sx={{ m: 2 }}>The content of the Popover.</Typography>
      </Popover>
      <Grid container spacing={2}>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl component="fieldset">
            <FormLabel component="legend">anchorReference</FormLabel>
            <RadioGroup
              row
              aria-label="anchor reference"
              name="anchorReference"
              value={anchorReference}
              onChange={handleChange}
            >
              <FormControlLabel
                value="anchorEl"
                control={<Radio />}
                label="anchorEl"
              />
              <FormControlLabel
                value="anchorPosition"
                control={<Radio />}
                label="anchorPosition"
              />
            </RadioGroup>
          </FormControl>
        </Grid>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl variant="standard">
            <InputLabel htmlFor="position-top">anchorPosition.top</InputLabel>
            <Input
              id="position-top"
              type="number"
              value={positionTop}
              onChange={handleNumberInputChange('positionTop')}
            />
          </FormControl>
          &nbsp;
          <FormControl variant="standard">
            <InputLabel htmlFor="position-left">anchorPosition.left</InputLabel>
            <Input
              id="position-left"
              type="number"
              value={positionLeft}
              onChange={handleNumberInputChange('positionLeft')}
            />
          </FormControl>
        </Grid>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl component="fieldset">
            <FormLabel component="legend">anchorOrigin.vertical</FormLabel>
            <RadioGroup
              aria-label="anchor origin vertical"
              name="anchorOriginVertical"
              value={anchorOriginVertical}
              onChange={handleChange}
            >
              <FormControlLabel
                value="top"
                control={<Radio sx={radioAnchorClasses} />}
                label="Top"
              />
              <FormControlLabel
                value="center"
                control={<Radio sx={radioAnchorClasses} />}
                label="Center"
              />
              <FormControlLabel
                value="bottom"
                control={<Radio sx={radioAnchorClasses} />}
                label="Bottom"
              />
            </RadioGroup>
          </FormControl>
        </Grid>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl component="fieldset">
            <FormLabel component="legend">transformOrigin.vertical</FormLabel>
            <RadioGroup
              aria-label="transform origin vertical"
              name="transformOriginVertical"
              value={transformOriginVertical}
              onChange={handleChange}
            >
              <FormControlLabel value="top" control={<Radio />} label="Top" />
              <FormControlLabel
                value="center"
                control={<Radio color="primary" />}
                label="Center"
              />
              <FormControlLabel
                value="bottom"
                control={<Radio color="primary" />}
                label="Bottom"
              />
            </RadioGroup>
          </FormControl>
        </Grid>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl component="fieldset">
            <FormLabel component="legend">anchorOrigin.horizontal</FormLabel>
            <RadioGroup
              row
              aria-label="anchor origin horizontal"
              name="anchorOriginHorizontal"
              value={anchorOriginHorizontal}
              onChange={handleChange}
            >
              <FormControlLabel
                value="left"
                control={<Radio sx={radioAnchorClasses} />}
                label="Left"
              />
              <FormControlLabel
                value="center"
                control={<Radio sx={radioAnchorClasses} />}
                label="Center"
              />
              <FormControlLabel
                value="right"
                control={<Radio sx={radioAnchorClasses} />}
                label="Right"
              />
            </RadioGroup>
          </FormControl>
        </Grid>
        <Grid
          size={{
            xs: 12,
            sm: 6,
          }}
        >
          <FormControl component="fieldset">
            <FormLabel component="legend">transformOrigin.horizontal</FormLabel>
            <RadioGroup
              row
              aria-label="transform origin horizontal"
              name="transformOriginHorizontal"
              value={transformOriginHorizontal}
              onChange={handleChange}
            >
              <FormControlLabel
                value="left"
                control={<Radio color="primary" />}
                label="Left"
              />
              <FormControlLabel
                value="center"
                control={<Radio color="primary" />}
                label="Center"
              />
              <FormControlLabel
                value="right"
                control={<Radio color="primary" />}
                label="Right"
              />
            </RadioGroup>
          </FormControl>
        </Grid>
      </Grid>
      <HighlightedCode code={jsx} language="jsx" />
    </div>
  );
}

export default AnchorPlayground;
```
</details>

You can also set `anchorReference = "anchorPosition"`. The component then
ignores the anchor element and uses the `anchorPosition` prop, which gives the
position of the popover in pixels from the top left corner of the window. As
nothing is anchored, a [reactRouter](https://felixluginbuhl.com/reactRouter/)
route opens it:

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/anchor-position-popover", "Open at (top: 200, left: 400)"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "anchor-position-popover",
        element = Popover(
          open = TRUE,
          onClose = JS("() => { window.location.hash = '/'; }"),
          anchorReference = "anchorPosition",
          anchorPosition = list(top = 200, left = 400),
          Typography(sx = list(p = 2), "The content of the Popover.")
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

## Mouse hover interaction

The MUI documentation shows a popover that opens with `mouseenter` and
`mouseleave` events. `Popover.triggerId()` opens on click. For content shown on
hover, use a [Tooltip](tooltip.html), which is designed for it.

## Virtual element

In React, the value of the `anchorEl` prop can be a reference to a fake DOM
element, for example to show a popover next to selected text. This needs
custom JavaScript and has no R equivalent.

## Transitions

Popover uses [Grow](transitions.html#grow) by default.
Use `slots$transition` and `slotProps$transition` to replace it with another transition or to pass transition props.
A component is expected, so it is passed as a JavaScript reference with `JS()`:

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(id = "fade-popover-button", variant = "contained", "Open Popover with Fade"),
  Popover.triggerId(
    triggerId = "fade-popover-button",
    anchorOrigin = list(vertical = "bottom", horizontal = "left"),
    slots = list(transition = JS("jsmodule['@mui/material'].Fade")),
    slotProps = list(transition = list(timeout = 600)),
    Typography(sx = list(p = 2), "The content of the Popover.")
  )
)
```

## Popovers in Shiny apps

`Popover.triggerId()` works unchanged in Shiny apps. A caller-supplied `onClose`
runs after the popover has closed itself, for example
`onClose = triggerEvent("popover_closed")` to notify the server.
