---
title: "Popper"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Popper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-popper/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Popper

A Popper can be used to display some content on top of another. It's an alternative to react-popper.

Some important features of the Popper component:

- 🕷 Popper relies on the 3rd party library ([Popper.js](https://popper.js.org/docs/v2/)) for perfect positioning.
- 💄 It's an alternative API to react-popper. It aims for simplicity.
- Its child element is a [Portal](https://mui.com/material-ui/react-portal/) on the body of the document to avoid rendering problems.
  You can disable this behavior with `disablePortal`.
- The scroll isn't blocked like with the [Popover](popover.html) component.
  The placement of the popper updates with the available area in the viewport.
- Clicking away does not hide the Popper component.
  If you need this behavior, you can use the [Click-Away Listener](click-away-listener.html).
- The `anchorEl` is passed as the reference object to create a new `Popper.js` instance.

<div class="alert alert-secondary">
**How the live demos on this page work.** The open state is kept in the URL
with [reactRouter](https://felixluginbuhl.com/reactRouter/): the route
`#/simple-popper` renders the popper, any other URL does not. The toggle button
is part of each route and links to the other state. `anchorEl` accepts a
function, so the anchor is looked up by `id` with `JS()`.
</div>

## Basic Popper

```{r}
anchorById <- function(id) JS(sprintf("() => document.getElementById('%s')", id))

simplePopper <- function(open) {
  tagList(
    Button(
      id = "simple-popper-button",
      variant = "outlined",
      `aria-describedby` = "simple-popper",
      href = if (open) "#/" else "#/simple-popper",
      "Toggle Popper"
    ),
    if (open) Popper(
      id = "simple-popper",
      open = TRUE,
      anchorEl = anchorById("simple-popper-button"),
      Box(sx = list(border = 1, p = 1, bgcolor = "background.paper"), "The content of the Popper.")
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      Route(path = "simple-popper", element = simplePopper(TRUE)),
      Route(path = "*", element = simplePopper(FALSE))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Popper from '@mui/material/Popper';

export default function SimplePopper() {
  const [anchorEl, setAnchorEl] = React.useState(null);

  const handleClick = (event) => {
    setAnchorEl(anchorEl ? null : event.currentTarget);
  };

  const open = Boolean(anchorEl);
  const id = open ? 'simple-popper' : undefined;

  return (
    <div>
      <button aria-describedby={id} type="button" onClick={handleClick}>
        Toggle Popper
      </button>
      <Popper id={id} open={open} anchorEl={anchorEl}>
        <Box sx={{ border: 1, p: 1, bgcolor: 'background.paper' }}>
          The content of the Popper.
        </Box>
      </Popper>
    </div>
  );
}
```
</details>

## Transitions

The open/close state of the popper can be animated with a render prop child and a transition component.
This component should respect the following conditions:

- Be a direct child descendent of the popper.
- Call the `onEnter` callback prop when the enter transition starts.
- Call the `onExited` callback prop when the exit transition is completed.
  These two callbacks allow the popper to unmount the child content when closed and fully transitioned.

Popper has built-in support for [react-transition-group](https://github.com/reactjs/react-transition-group).

A render prop (a function child) cannot be written in R. Wrapping the content
in a transition such as `Fade(in = TRUE)` animates the opening, as the content
mounts when the route becomes active:

```{r}
transitionsPopper <- function(open) {
  tagList(
    Button(
      id = "transitions-popper-button",
      variant = "outlined",
      href = if (open) "#/" else "#/transitions-popper",
      "Toggle Popper"
    ),
    if (open) Popper(
      open = TRUE,
      anchorEl = anchorById("transitions-popper-button"),
      Fade(
        `in` = TRUE,
        timeout = 350,
        Box(sx = list(border = 1, p = 1, bgcolor = "background.paper"), "The content of the Popper.")
      )
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      Route(path = "transitions-popper", element = transitionsPopper(TRUE)),
      Route(path = "*", element = transitionsPopper(FALSE))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Popper from '@mui/material/Popper';
import Fade from '@mui/material/Fade';

export default function TransitionsPopper() {
  const [open, setOpen] = React.useState(false);
  const [anchorEl, setAnchorEl] = React.useState(null);

  const handleClick = (event) => {
    setAnchorEl(event.currentTarget);
    setOpen((previousOpen) => !previousOpen);
  };

  const canBeOpen = open && Boolean(anchorEl);
  const id = canBeOpen ? 'transition-popper' : undefined;

  return (
    <div>
      <button aria-describedby={id} type="button" onClick={handleClick}>
        Toggle Popper
      </button>
      <Popper id={id} open={open} anchorEl={anchorEl} transition>
        {({ TransitionProps }) => (
          <Fade {...TransitionProps} timeout={350}>
            <Box sx={{ border: 1, p: 1, bgcolor: 'background.paper' }}>
              The content of the Popper.
            </Box>
          </Fade>
        )}
      </Popper>
    </div>
  );
}
```
</details>

## Positioned popper

The `placement` prop sets where the popper appears relative to its anchor.
Here each placement is a route (`#/positioned-popper/<placement>`). The
buttons link to their own placement, or back to `#/` when their popper is open.

```{r}
placements <- list(
  top = c("top-start", "top", "top-end"),
  left = c("left-start", "left", "left-end"),
  right = c("right-start", "right", "right-end"),
  bottom = c("bottom-start", "bottom", "bottom-end")
)

positionedPopper <- function(active = NULL) {
  placementButton <- function(placement) {
    Button(
      id = paste0("placement-", placement),
      href = if (identical(placement, active)) "#/" else paste0("#/positioned-popper/", placement),
      placement
    )
  }
  Box(
    sx = list(width = 500, maxWidth = "100%"),
    if (!is.null(active)) Popper(
      sx = list(zIndex = 1200),
      open = TRUE,
      anchorEl = anchorById(paste0("placement-", active)),
      placement = active,
      Fade(`in` = TRUE, timeout = 350, Paper(Typography(sx = list(p = 2), "The content of the Popper.")))
    ),
    Stack(direction = "row", sx = list(justifyContent = "center"), lapply(placements$top, placementButton)),
    Box(
      sx = list(display = "flex", justifyContent = "space-between"),
      Stack(direction = "column", sx = list(alignItems = "flex-start"), lapply(placements$left, placementButton)),
      Stack(direction = "column", sx = list(alignItems = "flex-end"), lapply(placements$right, placementButton))
    ),
    Stack(direction = "row", sx = list(justifyContent = "center"), lapply(placements$bottom, placementButton))
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(unlist(placements, use.names = FALSE), function(placement) {
        Route(path = paste0("positioned-popper/", placement), element = positionedPopper(placement))
      }),
      Route(path = "*", element = positionedPopper())
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Popper from '@mui/material/Popper';
import Typography from '@mui/material/Typography';
import Stack from '@mui/material/Stack';
import Button from '@mui/material/Button';
import Fade from '@mui/material/Fade';
import Paper from '@mui/material/Paper';

export default function PositionedPopper() {
  const [anchorEl, setAnchorEl] = React.useState(null);
  const [open, setOpen] = React.useState(false);
  const [placement, setPlacement] = React.useState();

  const handleClick = (newPlacement) => (event) => {
    setAnchorEl(event.currentTarget);
    setOpen((prev) => placement !== newPlacement || !prev);
    setPlacement(newPlacement);
  };

  return (
    <Box sx={{ width: 500 }}>
      <Popper
        // Note: The following zIndex style is specifically for documentation purposes and may not be necessary in your application.
        sx={{ zIndex: 1200 }}
        open={open}
        anchorEl={anchorEl}
        placement={placement}
        transition
      >
        {({ TransitionProps }) => (
          <Fade {...TransitionProps} timeout={350}>
            <Paper>
              <Typography sx={{ p: 2 }}>The content of the Popper.</Typography>
            </Paper>
          </Fade>
        )}
      </Popper>
      <Stack direction="row" sx={{ justifyContent: 'center' }}>
        <Button onClick={handleClick('top-start')}>top-start</Button>
        <Button onClick={handleClick('top')}>top</Button>
        <Button onClick={handleClick('top-end')}>top-end</Button>
      </Stack>
      <Box sx={{ display: 'flex', justifyContent: 'space-between' }}>
        <Stack direction="column" sx={{ alignItems: 'flex-start' }}>
          <Button onClick={handleClick('left-start')}>left-start</Button>
          <Button onClick={handleClick('left')}>left</Button>
          <Button onClick={handleClick('left-end')}>left-end</Button>
        </Stack>
        <Stack direction="column" sx={{ alignItems: 'flex-end' }}>
          <Button onClick={handleClick('right-start')}>right-start</Button>
          <Button onClick={handleClick('right')}>right</Button>
          <Button onClick={handleClick('right-end')}>right-end</Button>
        </Stack>
      </Box>
      <Stack direction="row" sx={{ justifyContent: 'center' }}>
        <Button onClick={handleClick('bottom-start')}>bottom-start</Button>
        <Button onClick={handleClick('bottom')}>bottom</Button>
        <Button onClick={handleClick('bottom-end')}>bottom-end</Button>
      </Stack>
    </Box>
  );
}
```
</details>

## Scroll playground

The MUI documentation has an interactive playground for the `flip` and
`preventOverflow` modifiers of Popper.js. The modifiers are passed with the
`modifiers` prop, for example:

```r
Popper(
  modifiers = list(
    list(name = "flip", enabled = TRUE, options = list(padding = 8)),
    list(name = "preventOverflow", enabled = TRUE, options = list(tether = TRUE, padding = 8))
  )
)
```

## Virtual element

In React, the value of the `anchorEl` prop can be a fake DOM element shaped
like a Popper.js [`VirtualElement`](https://popper.js.org/docs/v2/virtual-elements/),
for example to show a popper next to selected text. This needs custom
JavaScript and has no R equivalent.

## Popover or Popper?

For most "open on click" cases, [Popover](popover.html) is simpler:
`Popover.triggerId()` manages the open state and closes on click-away. Use
`Popper()` when the page must stay scrollable and interactive while the content
is shown.
