---
title: "Quarto and R Markdown"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Quarto and R Markdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE}
library(muiMaterial)
```

muiMaterial components render in [Quarto](https://quarto.org/) documents and
R Markdown HTML documents, without a Shiny server. Every page of this website
is an R Markdown document: all the live demos are rendered this way.

## Rendering components

A code chunk that returns a muiMaterial component renders it in the document:

````markdown
```{r}
library(muiMaterial)

Stack(
  direction = "row",
  spacing = 2,
  Button(variant = "contained", "Contained"),
  Chip(label = "A chip", color = "primary"),
  Rating(defaultValue = 3)
)
```
````

```{r, echo = FALSE}
Stack(
  direction = "row",
  spacing = 2,
  Button(variant = "contained", "Contained"),
  Chip(label = "A chip", color = "primary"),
  Rating(defaultValue = 3)
)
```

Each chunk is its own React tree, so a `ThemeProvider()` only applies to the
components of its chunk.

## What works without a server

In a document, everything runs in the browser:

- **Uncontrolled components** keep their own state: `Rating(defaultValue = 3)`,
  `Accordion()`, `Checkbox(defaultChecked = TRUE)`, `TextField()`, `Slider(defaultValue = 30)`, ...
- **Tabs**: `TabContext.static()`, `TabList.static()` and `TabPanel()` (see [Tabs](tabs.html)).
- **Overlays**: the `.triggerId()` wrappers open dialogs, drawers, menus and
  popovers (see [Overlays with `.triggerId`](triggerid.html)).
- **State in the URL**: [reactRouter](https://felixluginbuhl.com/reactRouter/)
  routes hold the state of controlled components (see [Dialog](dialog.html) and
  [Using a router](routing.html)).
- **JavaScript callbacks** written with `JS()`.

The `.shinyInput()` wrappers render too, but they only report their value to a
Shiny server. For inputs that talk to R, use a Quarto document with
[`server: shiny`](https://quarto.org/docs/interactive/shiny/) or a Shiny app.

## `muiMaterialPage()` in documents

`muiMaterialPage()` is not required in a document, but it is the simplest way
to load the Google fonts:

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  Stack(direction = "row", spacing = 2, Icon("home"), Icon("favorite", color = "error"), Icon("star", color = "primary"))
)
```

The fonts (`useFontRoboto`, `useMaterialIcons*`) are added to the document head,
once even if several chunks request them.

`muiMaterialPage()` also removes Bootstrap by default (`suppressBootstrap = TRUE`):

- In **Quarto** and in **pkgdown** websites, the page theme is part of the page
  template, so it is not affected.
- In an **R Markdown `html_document`**, the Bootstrap theme is an HTML
  dependency, so it is removed and the document loses its styling. Use
  `muiMaterialPage(suppressBootstrap = FALSE, ...)` in R Markdown documents.

## `CssBaseline()` in documents

`CssBaseline()` sets global styles on the page `<body>` (margin, background,
font). In a document, you may prefer `ScopedCssBaseline()`, which only applies
them to its children (see [CSS Baseline](css-baseline.html)).

## Quarto example

````markdown
---
title: "My report"
format: html
---

```{r}
#| echo: false
library(muiMaterial)

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  Card(
    sx = list(maxWidth = 400),
    CardHeader(avatar = Avatar(Icon("insights")), title = "Summary", subheader = "Monthly report"),
    CardContent(Typography(variant = "body2", "All indicators are up this month.")),
    CardActions(Button(id = "details-button", size = "small", "Details"))
  ),
  Dialog.triggerId(
    triggerId = "details-button",
    DialogTitle("Details"),
    DialogContent(DialogContentText("Details rendered without a server."))
  )
)
```
````
