---
title: "Radio Group"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Radio Group}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-radio-button/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# Radio Group

The Radio Group allows the user to select one option from a set.

Use radio buttons when the user needs to see all available options.
If available options can be collapsed, consider using a [Select component](select.html) because it uses less space.

Radio buttons should have the most commonly used option selected by default.

## Radio group

`RadioGroup` is a helpful wrapper used to group `Radio` components that provides an easier API, and proper keyboard accessibility to the group.

```{r}
genderOptions <- tagList(
  FormControlLabel(value = "female", control = Radio(), label = "Female"),
  FormControlLabel(value = "male", control = Radio(), label = "Male"),
  FormControlLabel(value = "other", control = Radio(), label = "Other")
)

muiMaterialPage(
  CssBaseline(),
  FormControl(
    FormLabel(id = "radio-buttons-group-label", "Gender"),
    RadioGroup(
      `aria-labelledby` = "radio-buttons-group-label",
      defaultValue = "female",
      name = "radio-buttons-group",
      genderOptions
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';

export default function RadioButtonsGroup() {
  const id = React.useId();
  return (
    <FormControl>
      <FormLabel id={`${id}-label`}>Gender</FormLabel>
      <RadioGroup
        aria-labelledby={`${id}-label`}
        defaultValue="female"
        name="radio-buttons-group"
      >
        <FormControlLabel value="female" control={<Radio />} label="Female" />
        <FormControlLabel value="male" control={<Radio />} label="Male" />
        <FormControlLabel value="other" control={<Radio />} label="Other" />
      </RadioGroup>
    </FormControl>
  );
}
```
</details>

### Direction

To lay out the buttons horizontally, set the `row` prop:

```{r}
muiMaterialPage(
  CssBaseline(),
  FormControl(
    FormLabel(id = "row-radio-buttons-group-label", "Gender"),
    RadioGroup(
      row = TRUE,
      `aria-labelledby` = "row-radio-buttons-group-label",
      name = "row-radio-buttons-group",
      genderOptions,
      FormControlLabel(value = "disabled", disabled = TRUE, control = Radio(), label = "other")
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';

export default function RowRadioButtonsGroup() {
  const id = React.useId();
  return (
    <FormControl>
      <FormLabel id={`${id}-label`}>Gender</FormLabel>
      <RadioGroup row aria-labelledby={`${id}-label`} name="row-radio-buttons-group">
        <FormControlLabel value="female" control={<Radio />} label="Female" />
        <FormControlLabel value="male" control={<Radio />} label="Male" />
        <FormControlLabel value="other" control={<Radio />} label="Other" />
        <FormControlLabel
          value="disabled"
          disabled
          control={<Radio />}
          label="other"
        />
      </RadioGroup>
    </FormControl>
  );
}
```
</details>

### Controlled

In React, you control the radio group with the `value` and `onChange` props.
In a Shiny app, use `RadioGroup.shinyInput()`: the selected `value` is sent to
`input[[inputId]]`, and `updateRadioGroup.shinyInput()` changes it from the
server.

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(p = 2),
    FormControl(
      FormLabel(id = "controlled-radio-buttons-group", "Gender"),
      RadioGroup.shinyInput(
        inputId = "gender",
        `aria-labelledby` = "controlled-radio-buttons-group",
        name = "controlled-radio-buttons-group",
        value = "female",
        FormControlLabel(value = "female", control = Radio(), label = "Female"),
        FormControlLabel(value = "male", control = Radio(), label = "Male")
      )
    ),
    Button.shinyInput("reset", "Reset to female"),
    verbatimTextOutput("selected")
  )
)

server <- function(input, output, session) {
  output$selected <- renderPrint(input$gender)
  observeEvent(input$reset, updateRadioGroup.shinyInput(inputId = "gender", value = "female"))
}

shinyApp(ui, server)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';

export default function ControlledRadioButtonsGroup() {
  const id = React.useId();
  const [value, setValue] = React.useState('female');

  const handleChange = (event) => {
    setValue(event.target.value);
  };

  return (
    <FormControl>
      <FormLabel id={`${id}-label`}>Gender</FormLabel>
      <RadioGroup
        aria-labelledby={`${id}-label`}
        name="controlled-radio-buttons-group"
        value={value}
        onChange={handleChange}
      >
        <FormControlLabel value="female" control={<Radio />} label="Female" />
        <FormControlLabel value="male" control={<Radio />} label="Male" />
      </RadioGroup>
    </FormControl>
  );
}
```
</details>

Run the bundled version with `muiMaterialExample("RadioGroup")`.

## Standalone radio buttons

`Radio` can also be used standalone, without labels. In React, standalone radios
are controlled with the `checked` prop. Without React state, keep them in a
`RadioGroup()` (here with `row = TRUE`): the group keeps a single radio
checked.

A standalone `Radio.shinyInput()` reports a boolean (`checked`) to the server,
not the selected value: prefer `RadioGroup.shinyInput()` for a set of options.

```{r}
muiMaterialPage(
  CssBaseline(),
  RadioGroup(
    row = TRUE,
    name = "radio-buttons",
    defaultValue = "a",
    Radio(value = "a", slotProps = list(input = list(`aria-label` = "A"))),
    Radio(value = "b", slotProps = list(input = list(`aria-label` = "B")))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';

export default function RadioButtons() {
  const [selectedValue, setSelectedValue] = React.useState('a');

  const handleChange = (event) => {
    setSelectedValue(event.target.value);
  };

  return (
    <div>
      <Radio
        checked={selectedValue === 'a'}
        onChange={handleChange}
        value="a"
        name="radio-buttons"
        slotProps={{ input: { 'aria-label': 'A' } }}
      />
      <Radio
        checked={selectedValue === 'b'}
        onChange={handleChange}
        value="b"
        name="radio-buttons"
        slotProps={{ input: { 'aria-label': 'B' } }}
      />
    </div>
  );
}
```
</details>

## Size

Use the `size` prop or customize the font size of the svg icons to change the size of the radios.

```{r}
muiMaterialPage(
  CssBaseline(),
  RadioGroup(
    row = TRUE,
    name = "size-radio-button-demo",
    defaultValue = "a",
    Radio(value = "a", size = "small", inputProps = list(`aria-label` = "a")),
    Radio(value = "b", inputProps = list(`aria-label` = "b")),
    Radio(value = "c", inputProps = list(`aria-label` = "c"), sx = list("& .MuiSvgIcon-root" = list(fontSize = 28)))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';

export default function SizeRadioButtons() {
  const [selectedValue, setSelectedValue] = React.useState('a');
  const handleChange = (event) => {
    setSelectedValue(event.target.value);
  };

  const controlProps = (item) => ({
    checked: selectedValue === item,
    onChange: handleChange,
    value: item,
    name: 'size-radio-button-demo',
    inputProps: { 'aria-label': item },
  });

  return (
    <div>
      <Radio {...controlProps('a')} size="small" />
      <Radio {...controlProps('b')} />
      <Radio
        {...controlProps('c')}
        sx={{
          '& .MuiSvgIcon-root': {
            fontSize: 28,
          },
        }}
      />
    </div>
  );
}
```
</details>

## Color

```{r}
muiMaterialPage(
  CssBaseline(),
  RadioGroup(
    row = TRUE,
    name = "color-radio-button-demo",
    defaultValue = "a",
    Radio(value = "a", inputProps = list(`aria-label` = "a")),
    Radio(value = "b", color = "secondary", inputProps = list(`aria-label` = "b")),
    Radio(value = "c", color = "success", inputProps = list(`aria-label` = "c")),
    Radio(value = "d", color = "default", inputProps = list(`aria-label` = "d")),
    Radio(
      value = "e",
      inputProps = list(`aria-label` = "e"),
      # pink[800] and pink[600] from the MUI color palette
      sx = list(color = "#ad1457", "&.Mui-checked" = list(color = "#d81b60"))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { pink } from '@mui/material/colors';
import Radio from '@mui/material/Radio';

export default function ColorRadioButtons() {
  const [selectedValue, setSelectedValue] = React.useState('a');

  const handleChange = (event) => {
    setSelectedValue(event.target.value);
  };

  const controlProps = (item) => ({
    checked: selectedValue === item,
    onChange: handleChange,
    value: item,
    name: 'color-radio-button-demo',
    inputProps: { 'aria-label': item },
  });

  return (
    <div>
      <Radio {...controlProps('a')} />
      <Radio {...controlProps('b')} color="secondary" />
      <Radio {...controlProps('c')} color="success" />
      <Radio {...controlProps('d')} color="default" />
      <Radio
        {...controlProps('e')}
        sx={{
          color: pink[800],
          '&.Mui-checked': {
            color: pink[600],
          },
        }}
      />
    </div>
  );
}
```
</details>

## Label placement

You can change the placement of the label with the `FormControlLabel` component's `labelPlacement` prop:

```{r}
muiMaterialPage(
  CssBaseline(),
  FormControl(
    FormLabel(id = "label-placement-label", "Label placement"),
    RadioGroup(
      row = TRUE,
      `aria-labelledby` = "label-placement-label",
      name = "position",
      defaultValue = "top",
      FormControlLabel(value = "bottom", control = Radio(), label = "Bottom", labelPlacement = "bottom"),
      FormControlLabel(value = "end", control = Radio(), label = "End")
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';

export default function FormControlLabelPlacement() {
  const id = React.useId();
  return (
    <FormControl>
      <FormLabel id={`${id}-label`}>Label placement</FormLabel>
      <RadioGroup
        row
        aria-labelledby={`${id}-label`}
        name="position"
        defaultValue="top"
      >
        <FormControlLabel
          value="bottom"
          control={<Radio />}
          label="Bottom"
          labelPlacement="bottom"
        />
        <FormControlLabel value="end" control={<Radio />} label="End" />
      </RadioGroup>
    </FormControl>
  );
}
```
</details>

## Show error

In general, radio buttons should have a value selected by default. If this is not the case, you can display an error if no value is selected when the form is submitted:

The answer is kept in the URL with [reactRouter](https://felixluginbuhl.com/reactRouter/):
the form's `onSubmit` navigates to `#/quiz/<answer>`, and each route renders the
form with the matching `error` state and helper text.

```{r, message = FALSE}
library(reactRouter)

quizForm <- function(answer = NULL) {
  checked <- !is.null(answer)
  error <- checked && answer != "best"
  helperText <- if (!checked) {
    "Choose wisely"
  } else if (answer == "best") {
    "You got it!"
  } else if (answer == "worst") {
    "Sorry, wrong answer!"
  } else {
    "Please select an option."
  }
  Box(
    component = "form",
    onSubmit = JS(
      "(event) => {
        event.preventDefault();
        const answer = new FormData(event.currentTarget).get('quiz') || 'none';
        window.location.hash = '/quiz/' + answer;
      }"
    ),
    FormControl(
      sx = list(m = 3),
      error = error,
      variant = "standard",
      FormLabel(id = "quiz-label", "Pop quiz: MUI is…"),
      RadioGroup(
        `aria-labelledby` = "quiz-label",
        name = "quiz",
        defaultValue = if (checked && answer != "none") answer else "",
        FormControlLabel(value = "best", control = Radio(), label = "The best!"),
        FormControlLabel(value = "worst", control = Radio(), label = "The worst.")
      ),
      FormHelperText(helperText),
      Button(sx = list(mt = 1, mr = 1), type = "submit", variant = "outlined", "Check Answer")
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(c("best", "worst", "none"), function(answer) Route(path = paste0("quiz/", answer), element = quizForm(answer))),
      Route(path = "*", element = quizForm())
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormHelperText from '@mui/material/FormHelperText';
import FormLabel from '@mui/material/FormLabel';
import Button from '@mui/material/Button';

export default function ErrorRadios() {
  const id = React.useId();
  const [value, setValue] = React.useState('');
  const [error, setError] = React.useState(false);
  const [helperText, setHelperText] = React.useState('Choose wisely');

  const handleRadioChange = (event) => {
    setValue(event.target.value);
    setError(false);
    setHelperText('Choose wisely');
  };

  const handleSubmit = (event) => {
    event.preventDefault();

    if (value === 'best') {
      setHelperText('You got it!');
      setError(false);
    } else if (value === 'worst') {
      setHelperText('Sorry, wrong answer!');
      setError(true);
    } else {
      setHelperText('Please select an option.');
      setError(true);
    }
  };

  return (
    <form onSubmit={handleSubmit}>
      <FormControl sx={{ m: 3 }} error={error} variant="standard">
        <FormLabel id={`${id}-label`}>Pop quiz: MUI is…</FormLabel>
        <RadioGroup
          aria-labelledby={`${id}-label`}
          name="quiz"
          value={value}
          onChange={handleRadioChange}
        >
          <FormControlLabel value="best" control={<Radio />} label="The best!" />
          <FormControlLabel value="worst" control={<Radio />} label="The worst." />
        </RadioGroup>
        <FormHelperText>{helperText}</FormHelperText>
        <Button sx={{ mt: 1, mr: 1 }} type="submit" variant="outlined">
          Check Answer
        </Button>
      </FormControl>
    </form>
  );
}
```
</details>

## Customization

Here is an example of customizing the component.
You can learn more about this in the [overrides documentation page](https://mui.com/material-ui/customization/how-to-customize/).
The `styled()` icons of the original demo become `Box()` elements styled with
`sx`, passed to the `icon` and `checkedIcon` props:

```{r}
bpIconSx <- list(
  borderRadius = "50%",
  width = 16,
  height = 16,
  boxShadow = "inset 0 0 0 1px rgba(16,22,26,.2), inset 0 -1px 0 rgba(16,22,26,.1)",
  backgroundColor = "#f5f8fa",
  backgroundImage = "linear-gradient(180deg,hsla(0,0%,100%,.8),hsla(0,0%,100%,0))",
  ".Mui-focusVisible &" = list(outline = "2px auto rgba(19,124,189,.6)", outlineOffset = 2),
  "input:hover ~ &" = list(backgroundColor = "#ebf1f5"),
  "input:disabled ~ &" = list(boxShadow = "none", background = "rgba(206,217,224,.5)")
)
bpCheckedIconSx <- modifyList(bpIconSx, list(
  backgroundColor = "#137cbd",
  backgroundImage = "linear-gradient(180deg,hsla(0,0%,100%,.1),hsla(0,0%,100%,0))",
  "&::before" = list(
    display = "block", width = 16, height = 16,
    backgroundImage = "radial-gradient(#fff,#fff 28%,transparent 32%)", content = '""'
  ),
  "input:hover ~ &" = list(backgroundColor = "#106ba3")
))

# Inspired by blueprintjs
BpRadio <- function(...) {
  Radio(
    disableRipple = TRUE,
    color = "default",
    checkedIcon = Box(component = "span", sx = bpCheckedIconSx),
    icon = Box(component = "span", sx = bpIconSx),
    ...
  )
}

muiMaterialPage(
  CssBaseline(),
  FormControl(
    FormLabel(id = "customized-radios-label", "Gender"),
    RadioGroup(
      defaultValue = "female",
      `aria-labelledby` = "customized-radios-label",
      name = "customized-radios",
      FormControlLabel(value = "female", control = BpRadio(), label = "Female"),
      FormControlLabel(value = "male", control = BpRadio(), label = "Male"),
      FormControlLabel(value = "other", control = BpRadio(), label = "Other"),
      FormControlLabel(value = "disabled", disabled = TRUE, control = BpRadio(), label = "(Disabled option)")
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import Radio from '@mui/material/Radio';
import RadioGroup from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import FormControl from '@mui/material/FormControl';
import FormLabel from '@mui/material/FormLabel';

const BpIcon = styled('span')(({ theme }) => ({
  borderRadius: '50%',
  width: 16,
  height: 16,
  boxShadow: 'inset 0 0 0 1px rgba(16,22,26,.2), inset 0 -1px 0 rgba(16,22,26,.1)',
  backgroundColor: '#f5f8fa',
  backgroundImage: 'linear-gradient(180deg,hsla(0,0%,100%,.8),hsla(0,0%,100%,0))',
  '.Mui-focusVisible &': {
    outline: '2px auto rgba(19,124,189,.6)',
    outlineOffset: 2,
  },
  'input:hover ~ &': {
    backgroundColor: '#ebf1f5',
    ...theme.applyStyles('dark', {
      backgroundColor: '#30404d',
    }),
  },
  'input:disabled ~ &': {
    boxShadow: 'none',
    background: 'rgba(206,217,224,.5)',
    ...theme.applyStyles('dark', {
      background: 'rgba(57,75,89,.5)',
    }),
    '@media (forced-colors: active)': {
      outline: '1px solid GrayText',
    },
  },
  '@media (forced-colors: active)': {
    outline: '1px solid ButtonText',
  },
  ...theme.applyStyles('dark', {
    boxShadow: '0 0 0 1px rgb(16 22 26 / 40%)',
    backgroundColor: '#394b59',
    backgroundImage: 'linear-gradient(180deg,hsla(0,0%,100%,.05),hsla(0,0%,100%,0))',
  }),
}));

const BpCheckedIcon = styled(BpIcon)({
  backgroundColor: '#137cbd',
  backgroundImage: 'linear-gradient(180deg,hsla(0,0%,100%,.1),hsla(0,0%,100%,0))',
  '&::before': {
    display: 'block',
    width: 16,
    height: 16,
    backgroundImage: 'radial-gradient(#fff,#fff 28%,transparent 32%)',
    content: '""',
    '@media (forced-colors: active)': {
      backgroundImage: 'none',
      backgroundColor: 'ButtonText',
      borderRadius: '50%',
      width: 8,
      height: 8,
      margin: 4,
    },
  },
  'input:hover ~ &': {
    backgroundColor: '#106ba3',
  },
  '@media (forced-colors: active)': {
    outline: '2px solid ButtonText',
  },
});

// Inspired by blueprintjs
function BpRadio(props) {
  return (
    <Radio
      disableRipple
      color="default"
      checkedIcon={<BpCheckedIcon />}
      icon={<BpIcon />}
      {...props}
    />
  );
}

export default function CustomizedRadios() {
  const id = React.useId();
  return (
    <FormControl>
      <FormLabel id={`${id}-label`}>Gender</FormLabel>
      <RadioGroup
        defaultValue="female"
        aria-labelledby={`${id}-label`}
        name="customized-radios"
      >
        <FormControlLabel value="female" control={<BpRadio />} label="Female" />
        <FormControlLabel value="male" control={<BpRadio />} label="Male" />
        <FormControlLabel value="other" control={<BpRadio />} label="Other" />
        <FormControlLabel
          value="disabled"
          disabled
          control={<BpRadio />}
          label="(Disabled option)"
        />
      </RadioGroup>
    </FormControl>
  );
}
```
</details>

## `useRadioGroup`

For advanced customization use cases, React exposes a `useRadioGroup()` hook
that returns the context value of the parent radio group. Hooks are not
available from R. The MUI demo uses it to color the label of the checked option,
which a CSS `:has()` selector also does:

```{r}
muiMaterialPage(
  CssBaseline(),
  RadioGroup(
    name = "use-radio-group",
    defaultValue = "first",
    sx = list(
      "& .MuiFormControlLabel-root:has(.Mui-checked) .MuiFormControlLabel-label" = list(color = "primary.main")
    ),
    FormControlLabel(value = "first", label = "First", control = Radio()),
    FormControlLabel(value = "second", label = "Second", control = Radio())
  )
)
```

<details><summary>JS code</summary>
```jsx
import { styled } from '@mui/material/styles';
import PropTypes from 'prop-types';
import RadioGroup, { useRadioGroup } from '@mui/material/RadioGroup';
import FormControlLabel from '@mui/material/FormControlLabel';
import Radio from '@mui/material/Radio';

const StyledFormControlLabel = styled((props) => <FormControlLabel {...props} />)(
  ({ theme }) => ({
    variants: [
      {
        props: { checked: true },
        style: {
          '.MuiFormControlLabel-label': {
            color: theme.palette.primary.main,
          },
        },
      },
    ],
  }),
);

function MyFormControlLabel(props) {
  const radioGroup = useRadioGroup();

  let checked = false;

  if (radioGroup) {
    checked = radioGroup.value === props.value;
  }

  return <StyledFormControlLabel checked={checked} {...props} />;
}

MyFormControlLabel.propTypes = {
  /**
   * The value of the component.
   */
  value: PropTypes.any,
};

export default function UseRadioGroup() {
  return (
    <RadioGroup name="use-radio-group" defaultValue="first">
      <MyFormControlLabel value="first" label="First" control={<Radio />} />
      <MyFormControlLabel value="second" label="Second" control={<Radio />} />
    </RadioGroup>
  );
}
```
</details>

## When to use

- [Checkboxes vs. Radio Buttons](https://www.nngroup.com/articles/checkboxes-vs-radio-buttons/)

## Accessibility

(WAI-ARIA: <https://www.w3.org/WAI/ARIA/apg/patterns/radio/>)

- All form controls should have labels, and this includes radio buttons, checkboxes, and switches. In most cases, this is done by using the `<label>` element ([FormControlLabel](https://mui.com/material-ui/api/form-control-label/)).
- When a label can't be used, it's necessary to add an attribute directly to the input component.
  In this case, you can apply the additional attribute (for example `aria-label`, `aria-labelledby`, `title`) via the `slotProps$input` property.

```r
Radio(value = "radioA", slotProps = list(input = list(`aria-label` = "Radio A")))
```
