---
title: "Rating"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rating}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-rating/).</div>

Ratings provide insight regarding others' opinions and experiences, and can allow the user to submit a rating of their own.

```{r setup, include=FALSE}
library(shinyMaterialUI)
library(shiny)
```

## Controlled

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Rating from '@mui/material/Rating';
import Typography from '@mui/material/Typography';

export default function BasicRating() {
  const [value, setValue] = React.useState<number | null>(2);

  return (
    <Box sx={{ '& > legend': { mt: 2 } }}>
      <Typography component="legend">Controlled</Typography>
      <Rating
        name="simple-controlled"
        value={value}
        onChange={(event, newValue) => {
          setValue(newValue);
        }}
      />
      <Typography component="legend">Uncontrolled</Typography>
      <Rating
        name="simple-uncontrolled"
        onChange={(event, newValue) => {
          console.log(newValue);
        }}
        defaultValue={2}
      />
      <Typography component="legend">Read only</Typography>
      <Rating name="read-only" value={value} readOnly />
      <Typography component="legend">Disabled</Typography>
      <Rating name="disabled" value={value} disabled />
      <Typography component="legend">No rating given</Typography>
      <Rating name="no-value" value={null} />
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shinyMaterialUI)
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    sx = list('& > legend' = list(mt = 2)),
    
    Typography(component = "legend", "Controlled"),
    Rating.shinyInput(
      inputId = "simple_controlled",
      value = 2
    ),
    
    Typography(component = "legend", "Uncontrolled"),
    Rating.shinyInput(
      inputId = "simple_uncontrolled",
      value = 2
    ),
    
    Typography(component = "legend", "Read only"),
    Rating(
      name = "read-only", 
      value = 2, 
      readOnly = TRUE
    ),
    
    Typography(component = "legend", "Disabled"),
    Rating(
      name = "disabled", 
      value = 2, 
      disabled = TRUE
    ),
    
    Typography(component = "legend", "No rating given"),
    Rating(
      name = "no-value", 
      value = NULL
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$simple_controlled, {
    # Handle the rating value change
  })
  
  observeEvent(input$simple_uncontrolled, {
    # Print the new value to console
    print(input$simple_uncontrolled)
  })
}

shinyApp(ui, server)
```

## Rating precision

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Rating from '@mui/material/Rating';
import Stack from '@mui/material/Stack';

export default function HalfRating() {
  return (
    <Stack spacing={1}>
      <Rating name="half-rating" defaultValue={2.5} precision={0.5} />
      <Rating name="half-rating-read" defaultValue={2.5} precision={0.5} readOnly />
    </Stack>
  );
}
```
</details>

```{r}
library(shinyMaterialUI)

CssBaseline(
  Stack(
    spacing = 1,
    Rating(
      value = 2.5,
      precision = 0.5
    ),
    Rating(
      name = "half-rating-read",
      value = 2.5,
      precision = 0.5,
      readOnly = TRUE
    )
  )
)
```

## Hover feedback

You can display a label on hover to help the user pick the correct rating value. The demo uses the onChangeActive prop.

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Rating from '@mui/material/Rating';
import Box from '@mui/material/Box';
import StarIcon from '@mui/icons-material/Star';

const labels: { [index: string]: string } = {
  0.5: 'Useless',
  1: 'Useless+',
  1.5: 'Poor',
  2: 'Poor+',
  2.5: 'Ok',
  3: 'Ok+',
  3.5: 'Good',
  4: 'Good+',
  4.5: 'Excellent',
  5: 'Excellent+',
};

function getLabelText(value: number) {
  return `${value} Star${value !== 1 ? 's' : ''}, ${labels[value]}`;
}

export default function HoverRating() {
  const [value, setValue] = React.useState<number | null>(2);
  const [hover, setHover] = React.useState(-1);

  return (
    <Box sx={{ width: 200, display: 'flex', alignItems: 'center' }}>
      <Rating
        name="hover-feedback"
        value={value}
        precision={0.5}
        getLabelText={getLabelText}
        onChange={(event, newValue) => {
          setValue(newValue);
        }}
        onChangeActive={(event, newHover) => {
          setHover(newHover);
        }}
        emptyIcon={<StarIcon style={{ opacity: 0.55 }} fontSize="inherit" />}
      />
      {value !== null && (
        <Box sx={{ ml: 2 }}>{labels[hover !== -1 ? hover : value]}</Box>
      )}
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
# Hover Feedback Example
labels <- list(
  "0.5" = "Useless",
  "1" = "Useless+",
  "1.5" = "Poor",
  "2" = "Poor+",
  "2.5" = "Ok",
  "3" = "Ok+",
  "3.5" = "Good",
  "4" = "Good+",
  "4.5" = "Excellent",
  "5" = "Excellent+"
)

ui <- shinyMaterialUIPage(
  div(
    Rating.shinyInput(
      inputId = "hover_feedback",
      value = 2,
      precision = 0.5
    ),
    # Note: Hover feedback is more complex to replicate exactly in Shiny
    textOutput("hover_label")
  )
)

server <- function(input, output, session) {
  output$hover_label <- renderText({
    req(input$hover_feedback)
    labels[[as.character(input$hover_feedback)]]
  })
}

shinyApp(ui, server)
```

## Sizes

For larger or smaller ratings use the size prop.

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Rating from '@mui/material/Rating';
import Stack from '@mui/material/Stack';

export default function RatingSize() {
  return (
    <Stack spacing={1}>
      <Rating name="size-small" defaultValue={2} size="small" />
      <Rating name="size-medium" defaultValue={2} />
      <Rating name="size-large" defaultValue={2} size="large" />
    </Stack>
  );
}
```
</details>

```{r}
library(shinyMaterialUI)

CssBaseline(
  Stack(
    spacing = 1,
    Rating(
      value = 2,
      size = "small"
    ),
    Rating(
      value = 2
    ),
    Rating(
      value = 2,
      size = "large"
    )
  )
)
```

## Customization

Here are some examples of customizing the component. You can learn more about this in the overrides documentation page.

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import Box from '@mui/material/Box';
import Rating from '@mui/material/Rating';
import FavoriteIcon from '@mui/icons-material/Favorite';
import FavoriteBorderIcon from '@mui/icons-material/FavoriteBorder';
import Typography from '@mui/material/Typography';

const StyledRating = styled(Rating)({
  '& .MuiRating-iconFilled': {
    color: '#ff6d75',
  },
  '& .MuiRating-iconHover': {
    color: '#ff3d47',
  },
});

export default function CustomizedRating() {
  return (
    <Box sx={{ '& > legend': { mt: 2 } }}>
      <Typography component="legend">Custom icon and color</Typography>
      <StyledRating
        name="customized-color"
        defaultValue={2}
        getLabelText={(value: number) => `${value} Heart${value !== 1 ? 's' : ''}`}
        precision={0.5}
        icon={<FavoriteIcon fontSize="inherit" />}
        emptyIcon={<FavoriteBorderIcon fontSize="inherit" />}
      />
      <Typography component="legend">10 stars</Typography>
      <Rating name="customized-10" defaultValue={2} max={10} />
    </Box>
  );
}
```
</details>

```{r}
shinyMaterialUIPage(
  CssBaseline(
    ThemeProvider(
      theme = list(
        components = list(
          MuiRating = list(
            styleOverrides = list(
              iconFilled = list(
                color = "#ff6d75"
              ),
              iconHover = list(
                color = "#ff3d47"
              )
            )
          )
        )
      ),
      Box(
        sx = list('& > legend' = list(mt = 2)),
        Typography("Custom icon and color", component = "legend"),
        Rating(
          name = "customized_color",
          defaultValue = 2,
          precision = 0.5,
          icon = shiny::icon("heart"),
          emptyIcon = shiny::icon("heart", class = "far")
        ),
        Typography("10 stars", component = "legend"),
        Rating(name = "customized_10", defaultValue = 2, max = 10)
      )
    )
  )
)
```

## Radio group

The rating is implemented with a radio group, set highlightSelectedOnly to restore the natural behavior.

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { styled } from '@mui/material/styles';
import Rating, { IconContainerProps } from '@mui/material/Rating';
import SentimentVeryDissatisfiedIcon from '@mui/icons-material/SentimentVeryDissatisfied';
import SentimentDissatisfiedIcon from '@mui/icons-material/SentimentDissatisfied';
import SentimentSatisfiedIcon from '@mui/icons-material/SentimentSatisfied';
import SentimentSatisfiedAltIcon from '@mui/icons-material/SentimentSatisfiedAltOutlined';
import SentimentVerySatisfiedIcon from '@mui/icons-material/SentimentVerySatisfied';

const StyledRating = styled(Rating)(({ theme }) => ({
  '& .MuiRating-iconEmpty .MuiSvgIcon-root': {
    color: theme.palette.action.disabled,
  },
}));

const customIcons: {
  [index: string]: {
    icon: React.ReactElement<unknown>;
    label: string;
  };
} = {
  1: {
    icon: <SentimentVeryDissatisfiedIcon color="error" />,
    label: 'Very Dissatisfied',
  },
  2: {
    icon: <SentimentDissatisfiedIcon color="error" />,
    label: 'Dissatisfied',
  },
  3: {
    icon: <SentimentSatisfiedIcon color="warning" />,
    label: 'Neutral',
  },
  4: {
    icon: <SentimentSatisfiedAltIcon color="success" />,
    label: 'Satisfied',
  },
  5: {
    icon: <SentimentVerySatisfiedIcon color="success" />,
    label: 'Very Satisfied',
  },
};

function IconContainer(props: IconContainerProps) {
  const { value, ...other } = props;
  return <span {...other}>{customIcons[value].icon}</span>;
}

export default function RadioGroupRating() {
  return (
    <StyledRating
      name="highlight-selected-only"
      defaultValue={2}
      IconContainerComponent={IconContainer}
      getLabelText={(value: number) => customIcons[value].label}
      highlightSelectedOnly
    />
  );
}
```
</details>

```{r eval=FALSE}
ui <- shinyMaterialUIPage(
  CssBaseline(
    ThemeProvider(
      theme = list(
        components = list(
          MuiRating = list(
            styleOverrides = list(
              iconEmpty = list(
                '& .MuiSvgIcon-root' = list(
                  color = "rgba(0, 0, 0, 0.26)"
                )
              )
            )
          )
        )
      ),
      Rating(
        name = "highlight_selected_only",
        defaultValue = 2,
        highlightSelectedOnly = TRUE,
        # Custom icons would need to be implemented with custom IconContainerComponent
        # For simplicity, using standard emoticon icons
        icon = shiny::icon("smile"),
        emptyIcon = shiny::icon("meh")
      )
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$radio_group_rating, {
    # Handle rating change
  })
}

shinyApp(ui, server)
```
