---
title: "Shiny inputs and server updates"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Shiny inputs and server updates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This guide explains how muiMaterial components talk to a Shiny server: the
`.shinyInput()` wrappers, the `update*.shinyInput()` functions, the reserved
props, and the lower-level helpers `triggerEvent()`, `setInput()`, `JS()`,
`renderReact()` and `reactOutput()`.

## `.shinyInput()` wrappers

A plain component such as `Slider()` renders in the browser only. Its
`.shinyInput()` variant, `Slider.shinyInput(inputId, ...)`, also sends its
value to the server as `input[[inputId]]`:

```{r}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(p = 2, width = 300),
    Slider.shinyInput("size", value = 30, min = 0, max = 100, valueLabelDisplay = "auto"),
    verbatimTextOutput("value")
  )
)

server <- function(input, output, session) {
  output$value <- renderPrint(input$size)
}

shinyApp(ui, server)
```

- The first argument is the `inputId`; every other argument is a prop of the
  MUI component, exactly as with the plain component.
- `value` sets the **initial** value.

There are two kinds of `.shinyInput()` wrappers.

**Value inputs** report their current value:

| Function | `input[[inputId]]` |
|---|---|
| `Autocomplete.shinyInput()` | the selected option, or a vector with `multiple = TRUE` |
| `BottomNavigation.shinyInput()` | the `value` of the selected action |
| `Checkbox.shinyInput()`, `Switch.shinyInput()`, `Radio.shinyInput()`, `FormControlLabel.shinyInput()` | `TRUE` or `FALSE` |
| `TextField.shinyInput()`, `Input.shinyInput()`, `OutlinedInput.shinyInput()`, `FilledInput.shinyInput()` | the text (sent 250ms after the last keystroke) |
| `Select.shinyInput()`, `NativeSelect.shinyInput()` | the selected value, or a vector with `multiple = TRUE` |
| `RadioGroup.shinyInput()` | the `value` of the selected radio |
| `Slider.shinyInput()` | a number, or a vector of two for a range (sent 250ms after the last move) |
| `Rating.shinyInput()` | a number |
| `Pagination.shinyInput()` | the page number (starting at 1) |
| `Tabs.shinyInput()`, `TabList.shinyInput()` | the `value` of the selected tab |
| `ToggleButtonGroup.shinyInput()` | the selected value, or a vector without `exclusive = TRUE` |

**Action inputs** report a click counter, like `shiny::actionButton()`:
`Button.shinyInput()`, `IconButton.shinyInput()`, `Fab.shinyInput()`,
`LoadingButton.shinyInput()`, `ListItemButton.shinyInput()`,
`MenuItem.shinyInput()`, `StepButton.shinyInput()`, `ToggleButton.shinyInput()`.
Use them with `observeEvent()`.

The overlays `Dialog.shinyInput()`, `Drawer.shinyInput()`, `Menu.shinyInput()`,
`Modal.shinyInput()` and `Snackbar.shinyInput()` are also action inputs: they
count clicks inside the surface. Their `open` state is set from the server; see
[Dialog](dialog.html#dialogs-in-shiny-apps) and
[Snackbar](snackbar.html#snackbars-in-shiny-apps). For overlays opened by a
button, [`.triggerId`](triggerid.html) is simpler.

The bundled showcase app lists all of them:

```{r}
muiMaterialExample("showcase")
```

## Updating inputs from the server

Each `.shinyInput()` has an `update*.shinyInput()` function. It changes **any
prop** of the component, not only its value:

```{r}
ui <- muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    sx = list(p = 2, width = 300),
    TextField.shinyInput("name", label = "Name", value = ""),
    Button.shinyInput("reset", "Reset", variant = "outlined"),
    Button.shinyInput("submit", "Submit", variant = "contained", disabled = TRUE)
  )
)

server <- function(input, output, session) {
  # Enable the submit button once a name is typed
  observe({
    updateButton.shinyInput(inputId = "submit", disabled = !nzchar(input$name))
  })
  # Clear the field and change its label
  observeEvent(input$reset, {
    updateTextField.shinyInput(inputId = "name", value = "", label = "Name (cleared)")
  })
}

shinyApp(ui, server)
```

`session` is the first argument of the update functions and defaults to the
current session, so `inputId` must be named.

## Reserved props

A `.shinyInput()` wrapper owns the props that carry its value:

- value inputs own `value` (or `checked` / `page`) and `onChange`,
- action inputs own `onClick`.

Passing one of them through `...` replaces the wrapper's own handler, and
`input[[inputId]]` stops updating. muiMaterial warns when this happens:

```{r}
Switch.shinyInput("dark", checked = TRUE)
#> Warning: Switch.shinyInput(): `checked` is a reserved prop -- a caller-supplied
#> value overrides the .shinyInput wiring and `input[[inputId]]` may stop updating. ...
```

Set the initial state with the `value` argument instead:
`Switch.shinyInput("dark", value = TRUE)`. For a custom browser-to-server signal
on a component, use a plain component with `onClick = triggerEvent("name")`
(see below).

Overriding the wiring on purpose is a legitimate advanced pattern, for example
when the server is the single source of truth and pushes `checked` down with
`renderReact()`, as in the [Transfer List](transfer-list.html#transfer-lists-in-shiny-apps)
example. Silence the warning with:

```{r}
options(muiMaterial.warnReservedProps = FALSE)
```

## Events without an input: `triggerEvent()` and `setInput()`

Any callback prop of any component (`onClick`, `onChange`, `onClose`, ...) can
send a value to the server, without a `.shinyInput()` wrapper:

- `triggerEvent("name")` sets `input$name` each time the callback runs. Use it
  with `observeEvent(input$name, ...)`.
- `setInput("name", accessor)` sets `input$name` to one of the arguments of the
  callback: `setInput("name")` is the first argument, `setInput("name", 1)` the
  second one, and `setInput("name", "[0].target.value")` a JavaScript accessor.

```{r}
ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(p = 2),
    Chip(label = "Delete me", onDelete = triggerEvent("chip_deleted")),
    Accordion(
      onChange = setInput("details_expanded", 1), # onChange(event, expanded)
      AccordionSummary("Details"),
      AccordionDetails("Content")
    ),
    verbatimTextOutput("state")
  )
)

server <- function(input, output, session) {
  output$state <- renderPrint(list(deleted = input$chip_deleted, expanded = input$details_expanded))
}

shinyApp(ui, server)
```

## JavaScript callbacks: `JS()`

Props that expect a JavaScript function (callbacks, render functions, `sx`
functions of the theme) are written as strings marked with `JS()`. The function
runs in the browser:

```{r}
Button(onClick = JS("() => alert('Hello from the browser')"), "Say hello")

# Send a custom value with Shiny's JavaScript API
MenuItem(onClick = JS("() => Shiny.setInputValue('action', 'edit', {priority: 'event'})"), "Edit")

# sx as a function of the theme
Box(sx = JS("(theme) => ({ color: theme.palette.primary.main })"), "Themed text")
```

`{priority: 'event'}` makes Shiny send the value even when it did not change,
so `observeEvent()` runs on every click. When a component prop expects a
component (not an element), pass it by reference, for example
`slots = list(transition = JS("jsmodule['@mui/material'].Fade"))`.

## Server-side rendering: `renderReact()` and `reactOutput()`

Render muiMaterial components from the server with either:

- `shiny::renderUI()` and `shiny::uiOutput()`, or
- `renderReact()` and `reactOutput()`, re-exported from shiny.react.

`renderReact()` updates the React elements already on the page instead of
replacing them. Transitions run, and the state of components that did not change is
kept. Prefer it for components that change often, such as progress bars or
[transitions](transitions.html#transitions-in-shiny-apps):

```{r}
ui <- muiMaterialPage(
  CssBaseline(),
  Box(sx = list(p = 2, width = 300), Slider.shinyInput("value", value = 40), reactOutput("progress"))
)

server <- function(input, output, session) {
  output$progress <- renderReact({
    LinearProgress(variant = "determinate", value = input$value)
  })
}

shinyApp(ui, server)
```

## Bookmarking

Shiny's [bookmarking](https://shiny.posit.co/r/articles/share/bookmarking-state/)
saves `input` values in the URL, but restores them only for Shiny's own input
bindings. Restore `.shinyInput()` values in `onRestore()` with their update
function:

```{r}
ui <- function(request) {
  muiMaterialPage(
    CssBaseline(),
    Box(
      sx = list(p = 2),
      TextField.shinyInput("txt", label = "Enter text", value = "initial"),
      verbatimTextOutput("show_txt")
    )
  )
}

server <- function(input, output, session) {
  output$show_txt <- renderPrint(input$txt)

  onRestore(function(state) {
    updateTextField.shinyInput(session, "txt", value = state$input$txt)
  })

  # Update the URL on every change
  observe({
    reactiveValuesToList(input)
    session$doBookmark()
  })
  onBookmarked(function(url) updateQueryString(url))
}

shinyApp(ui, server, enableBookmarking = "url")
```

`shiny::bookmarkButton()` needs Bootstrap, which `muiMaterialPage()` removes.
Updating the URL automatically, as above, avoids the button. Run this example
with `muiMaterialExample("bookmarking")`.

For state kept in the URL on the client side (no server), see
[Using a router](routing.html).
