---
title: "Snackbar"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Snackbar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-snackbar/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Snackbar

Snackbars (also known as toasts) are used for brief notifications of processes that have been or will be performed.

<div class="alert alert-secondary">
**How the live demos on this page work.** As on the [Dialog](dialog.html)
page, the open state is kept in the URL with
[reactRouter](https://felixluginbuhl.com/reactRouter/): each snackbar is a
route rendered with `open = TRUE`, a link opens it, and `onClose` navigates back
to `#/`. In a Shiny app, you would rather open snackbars from the server; see
[Snackbars in Shiny apps](#snackbars-in-shiny-apps).
</div>

```{r}
# Close on auto-hide, Escape or the close button, but not on a click elsewhere
closeSnackbar <- JS(
  "(event, reason) => {
    if (reason === 'clickaway') return;
    window.location.hash = '/';
  }"
)
```

## Introduction

The Snackbar component appears temporarily and floats above the UI to provide users with (non-critical) updates on an app's processes.
The demo below, inspired by Google Keep, shows a basic Snackbar with a text element and two actions:

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Button(href = "#/simple-snackbar", "Open Snackbar"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "simple-snackbar",
        element = Snackbar(
          open = TRUE,
          autoHideDuration = 6000,
          onClose = closeSnackbar,
          message = "Note archived",
          action = tagList(
            Button(color = "secondary", size = "small", href = "#/", "UNDO"),
            IconButton(
              size = "small",
              `aria-label` = "close",
              color = "inherit",
              href = "#/",
              Icon("close", fontSize = "small")
            )
          )
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';
import IconButton from '@mui/material/IconButton';
import CloseIcon from '@mui/icons-material/Close';

export default function SimpleSnackbar() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen(true);
  };

  const handleClose = (event, reason) => {
    if (reason === 'clickaway') {
      return;
    }

    setOpen(false);
  };

  const action = (
    <React.Fragment>
      <Button color="secondary" size="small" onClick={handleClose}>
        UNDO
      </Button>
      <IconButton
        size="small"
        aria-label="close"
        color="inherit"
        onClick={handleClose}
      >
        <CloseIcon fontSize="small" />
      </IconButton>
    </React.Fragment>
  );

  return (
    <div>
      <Button onClick={handleClick}>Open Snackbar</Button>
      <Snackbar
        open={open}
        autoHideDuration={6000}
        onClose={handleClose}
        message="Note archived"
        action={action}
      />
    </div>
  );
}
```
</details>

### Usage

Snackbars differ from [Alerts](alert.html) in that Snackbars have a fixed position and a high z-index, so they're intended to break out of the document flow; Alerts, on the other hand, are usually part of the flow, except when they're [used as children of a Snackbar](#use-with-alerts).

Snackbars also differ from [Dialogs](dialog.html) in that Snackbars are not intended to convey _critical_ information or block the user from interacting with the rest of the app; Dialogs, by contrast, require input from the user in order to be dismissed.

## Basics

### Position

Use the `anchorOrigin` prop to control the Snackbar's position on the screen.

```{r}
positions <- list(
  "top-center" = list(vertical = "top", horizontal = "center"),
  "top-left" = list(vertical = "top", horizontal = "left"),
  "top-right" = list(vertical = "top", horizontal = "right"),
  "bottom-center" = list(vertical = "bottom", horizontal = "center"),
  "bottom-left" = list(vertical = "bottom", horizontal = "left"),
  "bottom-right" = list(vertical = "bottom", horizontal = "right")
)
positionButton <- function(position) {
  Button(href = paste0("#/positioned-snackbar/", position), gsub("-", " ", position))
}

muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = 500, maxWidth = "100%"),
    Box(sx = list(display = "flex", justifyContent = "center"), positionButton("top-center")),
    Grid(
      container = TRUE,
      sx = list(justifyContent = "center"),
      Grid(size = 6, positionButton("top-left")),
      Grid(sx = list(textAlign = "right"), size = 6, positionButton("top-right")),
      Grid(size = 6, positionButton("bottom-left")),
      Grid(sx = list(textAlign = "right"), size = 6, positionButton("bottom-right"))
    ),
    Box(sx = list(display = "flex", justifyContent = "center"), positionButton("bottom-center")),
    RouterProvider(
      router = createHashRouter(
        lapply(names(positions), function(position) {
          Route(
            path = paste0("positioned-snackbar/", position),
            element = Snackbar(
              anchorOrigin = positions[[position]],
              open = TRUE,
              onClose = closeSnackbar,
              message = "I love snacks"
            )
          )
        }),
        Route(path = "*", element = NULL)
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Grid from '@mui/material/Grid';
import Box from '@mui/material/Box';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';

export default function PositionedSnackbar() {
  const [state, setState] = React.useState({
    open: false,
    vertical: 'top',
    horizontal: 'center',
  });
  const { vertical, horizontal, open } = state;

  const handleClick = (newState) => () => {
    setState({ ...newState, open: true });
  };

  const handleClose = () => {
    setState({ ...state, open: false });
  };

  const buttons = (
    <React.Fragment>
      <Box sx={{ display: 'flex', justifyContent: 'center' }}>
        <Button onClick={handleClick({ vertical: 'top', horizontal: 'center' })}>
          Top-Center
        </Button>
      </Box>
      <Grid container sx={{ justifyContent: 'center' }}>
        <Grid size={6}>
          <Button onClick={handleClick({ vertical: 'top', horizontal: 'left' })}>
            Top-Left
          </Button>
        </Grid>
        <Grid sx={{ textAlign: 'right' }} size={6}>
          <Button onClick={handleClick({ vertical: 'top', horizontal: 'right' })}>
            Top-Right
          </Button>
        </Grid>
        <Grid size={6}>
          <Button onClick={handleClick({ vertical: 'bottom', horizontal: 'left' })}>
            Bottom-Left
          </Button>
        </Grid>
        <Grid sx={{ textAlign: 'right' }} size={6}>
          <Button onClick={handleClick({ vertical: 'bottom', horizontal: 'right' })}>
            Bottom-Right
          </Button>
        </Grid>
      </Grid>
      <Box sx={{ display: 'flex', justifyContent: 'center' }}>
        <Button onClick={handleClick({ vertical: 'bottom', horizontal: 'center' })}>
          Bottom-Center
        </Button>
      </Box>
    </React.Fragment>
  );

  return (
    <Box sx={{ width: 500 }}>
      {buttons}
      <Snackbar
        anchorOrigin={{ vertical, horizontal }}
        open={open}
        onClose={handleClose}
        message="I love snacks"
        key={vertical + horizontal}
      />
    </Box>
  );
}
```
</details>

### Content

Use the `SnackbarContent()` component to add text and actions to the Snackbar.

```{r}
action <- Button(color = "secondary", size = "small", "lorem ipsum dolorem")

muiMaterialPage(
  CssBaseline(),
  Stack(
    spacing = 2,
    sx = list(maxWidth = 600),
    SnackbarContent(message = "I love snacks.", action = action),
    SnackbarContent(
      message = "I love candy. I love cookies. I love cupcakes. I love cheesecake. I love chocolate."
    ),
    SnackbarContent(message = "I love candy. I love cookies. I love cupcakes.", action = action),
    SnackbarContent(
      message = "I love candy. I love cookies. I love cupcakes. I love cheesecake. I love chocolate.",
      action = action
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Button from '@mui/material/Button';
import Stack from '@mui/material/Stack';
import SnackbarContent from '@mui/material/SnackbarContent';

const action = (
  <Button color="secondary" size="small">
    lorem ipsum dolorem
  </Button>
);

export default function LongTextSnackbar() {
  return (
    <Stack spacing={2} sx={{ maxWidth: 600 }}>
      <SnackbarContent message="I love snacks." action={action} />
      <SnackbarContent
        message={
          'I love candy. I love cookies. I love cupcakes. \
          I love cheesecake. I love chocolate.'
        }
      />
      <SnackbarContent
        message="I love candy. I love cookies. I love cupcakes."
        action={action}
      />
      <SnackbarContent
        message={
          'I love candy. I love cookies. I love cupcakes. \
          I love cheesecake. I love chocolate.'
        }
        action={action}
      />
    </Stack>
  );
}
```
</details>

### Automatic dismiss

Use the `autoHideDuration` prop to automatically trigger the Snackbar's `onClose` function after a set period of time (in milliseconds).

Make sure to [provide sufficient time](https://www.w3.org/WAI/WCAG22/Understanding/timing-adjustable.html) for the user to process the information displayed on it.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/autohide-snackbar", "Open Snackbar"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "autohide-snackbar",
        element = Snackbar(
          open = TRUE,
          autoHideDuration = 5000,
          onClose = closeSnackbar,
          message = "This Snackbar will be dismissed in 5 seconds."
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';

export default function AutohideSnackbar() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen(true);
  };

  const handleClose = (event, reason) => {
    if (reason === 'clickaway') {
      return;
    }

    setOpen(false);
  };

  return (
    <div>
      <Button onClick={handleClick}>Open Snackbar</Button>
      <Snackbar
        open={open}
        autoHideDuration={5000}
        onClose={handleClose}
        message="This Snackbar will be dismissed in 5 seconds."
      />
    </div>
  );
}
```
</details>

### Transitions

Use `slots$transition` and `slotProps$transition` to change the transition of
the Snackbar from [Grow](transitions.html#grow) (the default) to others such as
[Slide](transitions.html#slide). A component is expected, so it is passed as a
JavaScript reference with `JS()`.

```{r}
transitions <- list(
  grow = list(slots = list(transition = JS("jsmodule['@mui/material'].Grow"))),
  fade = list(slots = list(transition = JS("jsmodule['@mui/material'].Fade"))),
  slide = list(
    slots = list(transition = JS("jsmodule['@mui/material'].Slide")),
    slotProps = list(transition = list(direction = "up"))
  )
)

muiMaterialPage(
  CssBaseline(),
  Button(href = "#/transition-snackbar/grow", "Grow Transition"),
  Button(href = "#/transition-snackbar/fade", "Fade Transition"),
  Button(href = "#/transition-snackbar/slide", "Slide Transition"),
  RouterProvider(
    router = createHashRouter(
      lapply(names(transitions), function(name) {
        Route(
          path = paste0("transition-snackbar/", name),
          element = do.call(Snackbar, c(
            list(open = TRUE, onClose = closeSnackbar, message = "I love snacks", autoHideDuration = 1200),
            transitions[[name]]
          ))
        )
      }),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';
import Fade from '@mui/material/Fade';
import Slide from '@mui/material/Slide';
import Grow from '@mui/material/Grow';

function SlideTransition(props) {
  return <Slide {...props} direction="up" />;
}

function GrowTransition(props) {
  return <Grow {...props} />;
}

export default function TransitionsSnackbar() {
  const [state, setState] = React.useState({
    open: false,
    Transition: Fade,
  });

  const handleClick = (Transition) => () => {
    setState({
      open: true,
      Transition,
    });
  };

  const handleClose = () => {
    setState({
      ...state,
      open: false,
    });
  };

  return (
    <div>
      <Button onClick={handleClick(GrowTransition)}>Grow Transition</Button>
      <Button onClick={handleClick(Fade)}>Fade Transition</Button>
      <Button onClick={handleClick(SlideTransition)}>Slide Transition</Button>
      <Snackbar
        open={state.open}
        onClose={handleClose}
        slots={{ transition: state.Transition }}
        message="I love snacks"
        key={state.Transition.name}
        autoHideDuration={1200}
      />
    </div>
  );
}
```
</details>

#### Control Slide direction

You can change the direction of the [Slide](transitions.html#slide) transition, as in the following example:

```{r}
# The snackbar is at the bottom left, so "Left" slides it in from the right side
directions <- c(Up = "up", Left = "right", Right = "left", Down = "down")
directionButton <- function(label) Button(href = paste0("#/direction-snackbar/", directions[[label]]), label)

muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = 300),
    Box(sx = list(display = "flex", justifyContent = "center"), directionButton("Up")),
    Grid(
      container = TRUE,
      sx = list(justifyContent = "center"),
      Grid(size = 6, directionButton("Left")),
      Grid(sx = list(textAlign = "right"), size = 6, directionButton("Right"))
    ),
    Box(sx = list(display = "flex", justifyContent = "center"), directionButton("Down")),
    RouterProvider(
      router = createHashRouter(
        lapply(unname(directions), function(direction) {
          Route(
            path = paste0("direction-snackbar/", direction),
            element = Snackbar(
              open = TRUE,
              onClose = closeSnackbar,
              slots = list(transition = JS("jsmodule['@mui/material'].Slide")),
              slotProps = list(transition = list(direction = direction)),
              message = "I love snacks"
            )
          )
        }),
        Route(path = "*", element = NULL)
      )
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Grid from '@mui/material/Grid';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';
import Slide from '@mui/material/Slide';

function TransitionLeft(props) {
  return <Slide {...props} direction="left" />;
}

function TransitionUp(props) {
  return <Slide {...props} direction="up" />;
}

function TransitionRight(props) {
  return <Slide {...props} direction="right" />;
}

function TransitionDown(props) {
  return <Slide {...props} direction="down" />;
}

export default function DirectionSnackbar() {
  const [open, setOpen] = React.useState(false);
  const [transition, setTransition] = React.useState(undefined);

  const handleClick = (Transition) => () => {
    setTransition(() => Transition);
    setOpen(true);
  };

  const handleClose = () => {
    setOpen(false);
  };

  return (
    <Box sx={{ width: 300 }}>
      <Box sx={{ display: 'flex', justifyContent: 'center' }}>
        <Button onClick={handleClick(TransitionUp)}>Up</Button>
      </Box>
      <Grid container sx={{ justifyContent: 'center' }}>
        <Grid size={6}>
          <Button onClick={handleClick(TransitionRight)}>Left</Button>
        </Grid>
        <Grid sx={{ textAlign: 'right' }} size={6}>
          <Button onClick={handleClick(TransitionLeft)}>Right</Button>
        </Grid>
      </Grid>
      <Box sx={{ display: 'flex', justifyContent: 'center' }}>
        <Button onClick={handleClick(TransitionDown)}>Down</Button>
      </Box>
      <Snackbar
        open={open}
        onClose={handleClose}
        slots={{ transition }}
        message="I love snacks"
        key={transition ? transition.name : ''}
      />
    </Box>
  );
}
```
</details>

## Customization

### Preventing default click away event

If you would like to prevent the default `onClickAway` behavior, you can set the event's `defaultMuiPrevented` property to `true`:

```r
Snackbar(
  slotProps = list(
    clickAwayListener = list(
      # Prevents the default 'onClickAway' behavior
      onClickAway = JS("(event) => { event.defaultMuiPrevented = true; }")
    )
  )
)
```

### Use with Alerts

Use an [Alert](alert.html) inside a Snackbar for messages that communicate a certain severity.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(href = "#/alert-snackbar", "Open Snackbar"),
  RouterProvider(
    router = createHashRouter(
      Route(
        path = "alert-snackbar",
        element = Snackbar(
          open = TRUE,
          autoHideDuration = 6000,
          onClose = closeSnackbar,
          Alert(
            onClose = closeSnackbar,
            severity = "success",
            variant = "filled",
            sx = list(width = "100%"),
            "This is a success Alert inside a Snackbar!"
          )
        )
      ),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';
import Alert from '@mui/material/Alert';

export default function CustomizedSnackbars() {
  const [open, setOpen] = React.useState(false);

  const handleClick = () => {
    setOpen(true);
  };

  const handleClose = (event, reason) => {
    if (reason === 'clickaway') {
      return;
    }

    setOpen(false);
  };

  return (
    <div>
      <Button onClick={handleClick}>Open Snackbar</Button>
      <Snackbar open={open} autoHideDuration={6000} onClose={handleClose}>
        <Alert
          onClose={handleClose}
          severity="success"
          variant="filled"
          sx={{ width: '100%' }}
        >
          This is a success Alert inside a Snackbar!
        </Alert>
      </Snackbar>
    </div>
  );
}
```
</details>

### Use with Floating Action Buttons

If you're using a [Floating Action Button](floating-action-button.html) on mobile, Material Design recommends positioning snackbars directly above it, as shown in the demo below.
The demo is rendered in a framed box whose `transform` keeps the fixed-position
snackbar inside it:

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Box(
    sx = list(
      position = "relative", height = 300, maxWidth = 400, overflow = "hidden",
      transform = "translateZ(0)", border = 1, borderColor = "divider", bgcolor = "background.paper"
    ),
    AppBar(
      position = "static",
      color = "primary",
      Toolbar(
        IconButton(edge = "start", sx = list(mr = 2), color = "inherit", `aria-label` = "menu", Icon("menu")),
        Typography(variant = "h6", component = "div", sx = list(color = "inherit"), "App bar")
      )
    ),
    Fab(color = "secondary", sx = list(position = "absolute", bottom = 16, right = 16), Icon("add")),
    Snackbar(
      open = TRUE,
      message = "Archived",
      action = Button(color = "inherit", size = "small", "Undo"),
      sx = list(bottom = list(xs = 90, sm = 90))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import AppBar from '@mui/material/AppBar';
import CssBaseline from '@mui/material/CssBaseline';
import GlobalStyles from '@mui/material/GlobalStyles';
import Toolbar from '@mui/material/Toolbar';
import IconButton from '@mui/material/IconButton';
import MenuIcon from '@mui/icons-material/Menu';
import Typography from '@mui/material/Typography';
import Button from '@mui/material/Button';
import Fab from '@mui/material/Fab';
import AddIcon from '@mui/icons-material/Add';
import Snackbar from '@mui/material/Snackbar';

export default function FabIntegrationSnackbar() {
  return (
    <React.Fragment>
      <CssBaseline />
      <GlobalStyles
        styles={(theme) => ({
          body: { backgroundColor: theme.palette.background.paper },
        })}
      />
      <div>
        <AppBar position="static" color="primary">
          <Toolbar>
            <IconButton
              edge="start"
              sx={{ mr: 2 }}
              color="inherit"
              aria-label="menu"
            >
              <MenuIcon />
            </IconButton>
            <Typography
              variant="h6"
              component="div"
              sx={{
                color: 'inherit',
              }}
            >
              App bar
            </Typography>
          </Toolbar>
        </AppBar>
        <Fab
          color="secondary"
          sx={(theme) => ({
            position: 'absolute',
            bottom: theme.spacing(2),
            right: theme.spacing(2),
          })}
        >
          <AddIcon />
        </Fab>
        <Snackbar
          open
          autoHideDuration={6000}
          message="Archived"
          action={
            <Button color="inherit" size="small">
              Undo
            </Button>
          }
          sx={{ bottom: { xs: 90, sm: 0 } }}
        />
      </div>
    </React.Fragment>
  );
}
```
</details>

## Common examples

### Consecutive Snackbars

This demo shows how to display multiple Snackbars without stacking them. Each
message is its own route, so showing message B while message A is open replaces
it.

```{r}
consecutiveSnackbar <- function(message) {
  Snackbar(
    open = TRUE,
    autoHideDuration = 6000,
    onClose = closeSnackbar,
    message = message,
    action = tagList(
      Button(color = "secondary", size = "small", href = "#/", "UNDO"),
      IconButton(`aria-label` = "close", color = "inherit", sx = list(p = 0.5), href = "#/", Icon("close"))
    )
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Button(href = "#/consecutive-snackbar/a", "Show message A"),
  Button(href = "#/consecutive-snackbar/b", "Show message B"),
  RouterProvider(
    router = createHashRouter(
      Route(path = "consecutive-snackbar/a", element = consecutiveSnackbar("Message A")),
      Route(path = "consecutive-snackbar/b", element = consecutiveSnackbar("Message B")),
      Route(path = "*", element = NULL)
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Button from '@mui/material/Button';
import Snackbar from '@mui/material/Snackbar';
import IconButton from '@mui/material/IconButton';
import CloseIcon from '@mui/icons-material/Close';

export default function ConsecutiveSnackbars() {
  const [snackPack, setSnackPack] = React.useState([]);
  const [open, setOpen] = React.useState(false);
  const [messageInfo, setMessageInfo] = React.useState(undefined);

  React.useEffect(() => {
    if (snackPack.length && !messageInfo) {
      // Set a new snack when we don't have an active one
      setMessageInfo({ ...snackPack[0] });
      setSnackPack((prev) => prev.slice(1));
      setOpen(true);
    } else if (snackPack.length && messageInfo && open) {
      // Close an active snack when a new one is added
      setOpen(false);
    }
  }, [snackPack, messageInfo, open]);

  const handleClick = (message) => () => {
    setSnackPack((prev) => [...prev, { message, key: new Date().getTime() }]);
  };

  const handleClose = (event, reason) => {
    if (reason === 'clickaway') {
      return;
    }
    setOpen(false);
  };

  const handleExited = () => {
    setMessageInfo(undefined);
  };

  return (
    <div>
      <Button onClick={handleClick('Message A')}>Show message A</Button>
      <Button onClick={handleClick('Message B')}>Show message B</Button>
      <Snackbar
        key={messageInfo ? messageInfo.key : undefined}
        open={open}
        autoHideDuration={6000}
        onClose={handleClose}
        slotProps={{ transition: { onExited: handleExited } }}
        message={messageInfo ? messageInfo.message : undefined}
        action={
          <React.Fragment>
            <Button color="secondary" size="small" onClick={handleClose}>
              UNDO
            </Button>
            <IconButton
              aria-label="close"
              color="inherit"
              sx={{ p: 0.5 }}
              onClick={handleClose}
            >
              <CloseIcon />
            </IconButton>
          </React.Fragment>
        }
      />
    </div>
  );
}
```
</details>

## Accessibility

The user should be able to dismiss Snackbars by pressing <kbd class="key">Escape</kbd>. If there are multiple instances appearing at the same time and you want <kbd class="key">Escape</kbd> to dismiss only the oldest one that's currently open, call `event.preventDefault` in the `onClose` prop.

## Anatomy

The Snackbar component is composed of a root `<div>` that houses interior elements like the Snackbar Content and other optional components (such as buttons or decorators).

```html
<div role="presentation" class="MuiSnackbar-root">
  <div class="MuiPaper-root MuiSnackbarContent-root" role="alert">
    <div class="MuiSnackbarContent-message">
      <!-- Snackbar content goes here -->
    </div>
  </div>
</div>
```

## Snackbars in Shiny apps

Snackbars often report the result of server work ("Saved", "Upload failed").
With `Snackbar.shinyInput()`, the server opens the snackbar with
`updateSnackbar.shinyInput()`, and can change the message at the same time. The
wrapper does not manage `open`, so close it again from `onClose`:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Button.shinyInput("save", "Save", variant = "contained"),
  Snackbar.shinyInput(
    inputId = "snackbar",
    open = FALSE,
    autoHideDuration = 3000,
    message = "",
    onClose = triggerEvent("snackbar_close"),
    anchorOrigin = list(vertical = "bottom", horizontal = "center")
  )
)

server <- function(input, output, session) {
  observeEvent(input$save, {
    Sys.sleep(1) # some server work
    updateSnackbar.shinyInput(
      inputId = "snackbar",
      open = TRUE,
      message = paste("Saved at", format(Sys.time(), "%H:%M:%S"))
    )
  })
  observeEvent(input$snackbar_close, {
    updateSnackbar.shinyInput(inputId = "snackbar", open = FALSE)
  })
}

shinyApp(ui, server)
```

Run the bundled version with `muiMaterialExample("Snackbar")`.
