---
title: "Stepper"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Stepper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-stepper/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Stepper

Steppers convey progress through numbered steps. It provides a wizard-like workflow.

Steppers display progress through a sequence of logical and numbered steps. They may also be used for navigation.
Steppers may display a transient feedback message after a step is saved.

- **Types of Steps**: Editable, Non-editable, Mobile, Optional
- **Types of Steppers**: Horizontal, Vertical, Linear, Non-linear

<div class="alert alert-info">
This component is no longer documented in the [Material Design guidelines](https://m2.material.io/), but Material UI will continue to support it.
</div>

<div class="alert alert-secondary">
**How the live demos on this page work.** In React, the active step lives in a
`useState()` hook. Here, each state of the stepper is a
[reactRouter](https://felixluginbuhl.com/reactRouter/) route generated in R
(for example `#/linear-stepper/2`), and the Back and Next buttons are links to
the neighbouring states. In a Shiny app, the server usually holds the active
step: see [Steppers in Shiny apps](#steppers-in-shiny-apps).
</div>

## Introduction

The Stepper component displays progress through a sequence of logical and numbered steps.
It supports horizontal and vertical orientation for desktop and mobile viewports.

Steppers are implemented using a collection of related components:

- `Stepper()`: the container for the steps.
- `Step()`: an individual step in the sequence.
- `StepLabel()`: a label for a Step.
- `StepContent()`: optional content for a Step.
- `StepButton()`: optional button for a Step.
- `StepIcon()`: optional icon for a Step.
- `StepConnector()`: optional customized connector between Steps.

## Horizontal stepper

Horizontal steppers are ideal when the contents of one step depend on an earlier step.

Avoid using long step names in horizontal steppers.

### Linear

A linear stepper allows the user to complete the steps in sequence.

The `Stepper` can be controlled by passing the current step index (zero-based) as the `activeStep` prop. `Stepper` orientation is set using the `orientation` prop.

This example also shows the use of an optional step by placing the `optional` prop on the second `Step` component. Note that it's up to you to manage when an optional step is skipped. Once you've determined this for a particular step you must set `completed = FALSE` to signify that even though the active step index has gone beyond the optional step, it's not actually complete.

The route is `#/linear-stepper/<activeStep>`, with a `/skipped` suffix once the
optional step has been skipped.

```{r}
steps <- c("Select campaign settings", "Create an ad group", "Create an ad")
isStepOptional <- function(step) step == 1

linearHref <- function(step, skipped) {
  sprintf("#/linear-stepper/%s%s", step, if (skipped) "/skipped" else "")
}

horizontalLinearStepper <- function(activeStep = 0, skipped = FALSE) {
  Box(
    sx = list(width = "100%"),
    Stepper(
      activeStep = activeStep,
      lapply(seq_along(steps) - 1, function(index) {
        Step(
          completed = if (skipped && index == 1) FALSE else index < activeStep,
          StepLabel(
            optional = if (isStepOptional(index)) Typography(variant = "caption", "Optional") else FALSE,
            steps[index + 1]
          )
        )
      })
    ),
    if (activeStep == length(steps)) {
      tagList(
        Typography(sx = list(mt = 2, mb = 1), "All steps completed - you're finished"),
        Box(
          sx = list(display = "flex", flexDirection = "row", pt = 2),
          Box(sx = list(flex = "1 1 auto")),
          Button(href = "#/", "Reset")
        )
      )
    } else {
      tagList(
        Typography(sx = list(mt = 2, mb = 1), paste("Step", activeStep + 1)),
        Box(
          sx = list(display = "flex", flexDirection = "row", pt = 2),
          Button(
            color = "inherit",
            disabled = activeStep == 0,
            href = linearHref(max(activeStep - 1, 0), skipped && activeStep - 1 > 1),
            sx = list(mr = 1),
            "Back"
          ),
          Box(sx = list(flex = "1 1 auto")),
          if (isStepOptional(activeStep)) {
            Button(color = "inherit", href = linearHref(activeStep + 1, TRUE), sx = list(mr = 1), "Skip")
          },
          Button(
            href = linearHref(activeStep + 1, skipped && activeStep != 1),
            if (activeStep == length(steps) - 1) "Finish" else "Next"
          )
        )
      )
    }
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(1:3, function(step) Route(path = paste0("linear-stepper/", step), element = horizontalLinearStepper(step))),
      lapply(2:3, function(step) {
        Route(path = sprintf("linear-stepper/%s/skipped", step), element = horizontalLinearStepper(step, skipped = TRUE))
      }),
      Route(path = "*", element = horizontalLinearStepper(0))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';
import Button from '@mui/material/Button';
import Typography from '@mui/material/Typography';

const steps = ['Select campaign settings', 'Create an ad group', 'Create an ad'];

export default function HorizontalLinearStepper() {
  const [activeStep, setActiveStep] = React.useState(0);
  const [skipped, setSkipped] = React.useState(new Set());

  const isStepOptional = React.useCallback((step) => {
    return step === 1;
  }, []);

  const isStepSkipped = (step) => {
    return skipped.has(step);
  };

  const handleNext = () => {
    let newSkipped = skipped;
    if (isStepSkipped(activeStep)) {
      newSkipped = new Set(newSkipped.values());
      newSkipped.delete(activeStep);
    }

    setActiveStep((prevActiveStep) => prevActiveStep + 1);
    setSkipped(newSkipped);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const handleSkip = () => {
    if (!isStepOptional(activeStep)) {
      // You probably want to guard against something like this,
      // it should never occur unless someone's actively trying to break something.
      throw new Error("You can't skip a step that isn't optional.");
    }

    setActiveStep((prevActiveStep) => prevActiveStep + 1);
    setSkipped((prevSkipped) => {
      const newSkipped = new Set(prevSkipped.values());
      newSkipped.add(activeStep);
      return newSkipped;
    });
  };

  const handleReset = () => {
    setActiveStep(0);
  };

  const previousActiveStepRef = React.useRef(activeStep);
  const resetButtonRef = React.useRef(null);
  const nextButtonRef = React.useRef(null);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    const previousActiveStep = previousActiveStepRef.current;
    previousActiveStepRef.current = activeStep;

    if (activeStep === steps.length) {
      // If the user has completed all steps and hits "Finish", focus the "Reset" button.
      resetButtonRef.current.focus();
      return;
    }
    if (activeStep === 0 && previousActiveStep === steps.length) {
      // If the user has completed all steps and hits "Reset", focus the "Next" button.
      nextButtonRef.current.focus();
      return;
    }
    if (isStepOptional(previousActiveStep) && !isStepOptional(activeStep)) {
      // If the user hits "Skip" and the next step is not optional, focus the "Next" button.
      nextButtonRef.current.focus();
    }
  }, [activeStep, isStepOptional]);

  return (
    <Box sx={{ width: '100%' }}>
      <Stepper activeStep={activeStep}>
        {steps.map((label, index) => {
          const stepProps = {};
          const labelProps = {};
          if (isStepOptional(index)) {
            labelProps.optional = (
              <Typography variant="caption">Optional</Typography>
            );
          }
          if (isStepSkipped(index)) {
            stepProps.completed = false;
          }
          return (
            <Step key={label} {...stepProps}>
              <StepLabel {...labelProps}>{label}</StepLabel>
            </Step>
          );
        })}
      </Stepper>
      {activeStep === steps.length ? (
        <React.Fragment>
          <Typography sx={{ mt: 2, mb: 1 }}>
            All steps completed - you&apos;re finished
          </Typography>
          <Box sx={{ display: 'flex', flexDirection: 'row', pt: 2 }}>
            <Box sx={{ flex: '1 1 auto' }} />
            <Button onClick={handleReset} ref={resetButtonRef}>
              Reset
            </Button>
          </Box>
        </React.Fragment>
      ) : (
        <React.Fragment>
          <Typography sx={{ mt: 2, mb: 1 }}>Step {activeStep + 1}</Typography>
          <Box sx={{ display: 'flex', flexDirection: 'row', pt: 2 }}>
            <Button
              color="inherit"
              disabled={activeStep === 0}
              onClick={handleBack}
              sx={{ mr: 1 }}
            >
              Back
            </Button>
            <Box sx={{ flex: '1 1 auto' }} />
            {isStepOptional(activeStep) && (
              <Button color="inherit" onClick={handleSkip} sx={{ mr: 1 }}>
                Skip
              </Button>
            )}
            <Button onClick={handleNext} ref={nextButtonRef}>
              {activeStep === steps.length - 1 ? 'Finish' : 'Next'}
            </Button>
          </Box>
        </React.Fragment>
      )}
    </Box>
  );
}
```
</details>

### Non-linear

Non-linear steppers allow the user to enter a multi-step flow at any point.

This example is similar to the regular horizontal stepper, except steps are no longer automatically set to `disabled = TRUE` based on the `activeStep` prop.

The use of the `StepButton` here demonstrates clickable step labels, as well as setting the `completed`
flag. However because steps can be accessed in a non-linear fashion, it's up to your own implementation to
determine when all steps are completed (or even if they need to be completed).

Actionable steps mean that they control the content update of a section. From an accessibility standpoint, this means that each `StepButton` requires an `aria-controls` attribute that points to the `id` of the content area.

The state has two parts, the active step and the set of completed steps. Both
are in the route, `#/nonlinear-stepper/<activeStep>/<completed>`, where
`<completed>` has one digit per step (for example `101`). R generates the 24
routes with `expand.grid()`.

```{r}
nonLinearHref <- function(active, completed) {
  sprintf("#/nonlinear-stepper/%s/%s", active, paste(as.integer(completed), collapse = ""))
}

horizontalNonLinearStepper <- function(activeStep = 0, completed = c(FALSE, FALSE, FALSE)) {
  totalSteps <- length(steps)
  allStepsCompleted <- all(completed)
  isLastStep <- activeStep == totalSteps - 1
  # Next: go to the first uncompleted step when on the last one
  nextStep <- function(done) {
    if (isLastStep && !all(done)) which(!done)[1] - 1 else min(activeStep + 1, totalSteps - 1)
  }
  withCompleted <- replace(completed, activeStep + 1, TRUE)
  Box(
    sx = list(width = "100%"),
    Stepper(
      nonLinear = TRUE,
      activeStep = activeStep,
      lapply(seq_len(totalSteps) - 1, function(index) {
        Step(
          completed = completed[index + 1],
          StepButton(
            `aria-controls` = "stepper-content",
            color = "inherit",
            href = nonLinearHref(index, completed),
            steps[index + 1]
          )
        )
      })
    ),
    Box(
      id = "stepper-content",
      if (allStepsCompleted) {
        tagList(
          Typography(sx = list(mt = 2, mb = 1), "All steps completed - you're finished"),
          Box(sx = list(display = "flex", flexDirection = "row", pt = 2), Box(sx = list(flex = "1 1 auto")), Button(href = "#/", "Reset"))
        )
      } else {
        tagList(
          Typography(sx = list(mt = 2, mb = 1, py = 1), paste("Step", activeStep + 1)),
          Box(
            sx = list(display = "flex", flexDirection = "row", pt = 2),
            Button(
              color = "inherit",
              disabled = activeStep == 0,
              href = nonLinearHref(max(activeStep - 1, 0), completed),
              sx = list(mr = 1),
              "Back"
            ),
            Box(sx = list(flex = "1 1 auto")),
            Button(href = nonLinearHref(nextStep(completed), completed), sx = list(mr = 1), "Next"),
            if (completed[activeStep + 1]) {
              Typography(variant = "caption", sx = list(display = "inline-block"), paste("Step", activeStep + 1, "already completed"))
            } else {
              Button(
                href = nonLinearHref(if (all(withCompleted)) activeStep else nextStep(withCompleted), withCompleted),
                if (sum(completed) == totalSteps - 1) "Finish" else "Complete Step"
              )
            }
          )
        )
      }
    )
  )
}

states <- expand.grid(active = 0:2, c1 = c(FALSE, TRUE), c2 = c(FALSE, TRUE), c3 = c(FALSE, TRUE))

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(
      lapply(seq_len(nrow(states)), function(i) {
        completed <- unlist(states[i, c("c1", "c2", "c3")])
        Route(
          path = sub("^#/", "", nonLinearHref(states$active[i], completed)),
          element = horizontalNonLinearStepper(states$active[i], completed)
        )
      }),
      Route(path = "*", element = horizontalNonLinearStepper())
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepButton from '@mui/material/StepButton';
import Button from '@mui/material/Button';
import Typography from '@mui/material/Typography';

const steps = ['Select campaign settings', 'Create an ad group', 'Create an ad'];

export default function HorizontalNonLinearStepper() {
  const [activeStep, setActiveStep] = React.useState(0);
  const [completed, setCompleted] = React.useState({});

  const totalSteps = steps.length;
  const completedSteps = Object.keys(completed).length;
  const isLastStep = activeStep === totalSteps - 1;
  const allStepsCompleted = completedSteps === totalSteps;

  const handleNext = () => {
    const newActiveStep =
      isLastStep && !allStepsCompleted
        ? // It's the last step, but not all steps have been completed,
          // find the first step that has not been completed
          steps.findIndex((_step, i) => !(i in completed))
        : activeStep + 1;
    setActiveStep(newActiveStep);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const handleStep = (step) => () => {
    setActiveStep(step);
  };

  const handleComplete = () => {
    setCompleted({
      ...completed,
      [activeStep]: true,
    });
    handleNext();
  };

  const handleReset = () => {
    setActiveStep(0);
    setCompleted({});
  };

  const resetButtonRef = React.useRef(null);
  const nextButtonRef = React.useRef(null);
  const previousActiveStepRef = React.useRef(activeStep);
  const previousCompletedRef = React.useRef(completed);

  // Manage focus when the completed steps change.
  React.useEffect(() => {
    const previousCompleted = previousCompletedRef.current;
    previousCompletedRef.current = completed;

    if (allStepsCompleted) {
      // If the user has completed all steps and hits "Finish", focus the "Reset" button.
      resetButtonRef.current.focus();
      return;
    }

    if (
      Object.keys(completed).length === 0 &&
      Object.keys(previousCompleted).length !== 0
    ) {
      // If the user has completed all steps and hits "Reset", focus the "Next" button.
      nextButtonRef.current.focus();
    }
  }, [completed, allStepsCompleted]);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    if (activeStep === 0 && previousActiveStepRef.current === 1) {
      // If the user navigated to first step via "Back" button, focus the "Next" button.
      nextButtonRef.current.focus();
    }

    previousActiveStepRef.current = activeStep;
  }, [activeStep]);

  return (
    <Box sx={{ width: '100%' }}>
      <Stepper nonLinear activeStep={activeStep}>
        {steps.map((label, index) => (
          <Step key={label} completed={completed[index]}>
            <StepButton
              aria-controls="stepper-content"
              color="inherit"
              onClick={handleStep(index)}
            >
              {label}
            </StepButton>
          </Step>
        ))}
      </Stepper>
      <div id="stepper-content">
        {allStepsCompleted ? (
          <React.Fragment>
            <Typography sx={{ mt: 2, mb: 1 }}>
              All steps completed - you&apos;re finished
            </Typography>
            <Box sx={{ display: 'flex', flexDirection: 'row', pt: 2 }}>
              <Box sx={{ flex: '1 1 auto' }} />
              <Button onClick={handleReset} ref={resetButtonRef}>
                Reset
              </Button>
            </Box>
          </React.Fragment>
        ) : (
          <React.Fragment>
            <Typography sx={{ mt: 2, mb: 1, py: 1 }}>
              Step {activeStep + 1}
            </Typography>
            <Box sx={{ display: 'flex', flexDirection: 'row', pt: 2 }}>
              <Button
                color="inherit"
                disabled={activeStep === 0}
                onClick={handleBack}
                sx={{ mr: 1 }}
              >
                Back
              </Button>
              <Box sx={{ flex: '1 1 auto' }} />
              <Button onClick={handleNext} sx={{ mr: 1 }} ref={nextButtonRef}>
                Next
              </Button>
              {activeStep !== steps.length &&
                (completed[activeStep] ? (
                  <Typography variant="caption" sx={{ display: 'inline-block' }}>
                    Step {activeStep + 1} already completed
                  </Typography>
                ) : (
                  <Button onClick={handleComplete}>
                    {completedSteps === totalSteps - 1 ? 'Finish' : 'Complete Step'}
                  </Button>
                ))}
            </Box>
          </React.Fragment>
        )}
      </div>
    </Box>
  );
}
```
</details>

### Alternative label

Labels can be placed below the step icon by setting the `alternativeLabel` prop on the `Stepper` component.

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = "100%"),
    Stepper(activeStep = 1, alternativeLabel = TRUE, lapply(steps, function(label) Step(StepLabel(label))))
  )
)
```

<details><summary>JS code</summary>
```jsx
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';

const steps = [
  'Select master blaster campaign settings',
  'Create an ad group',
  'Create an ad',
];

export default function HorizontalLinearAlternativeLabelStepper() {
  return (
    <Box sx={{ width: '100%' }}>
      <Stepper activeStep={1} alternativeLabel>
        {steps.map((label) => (
          <Step key={label}>
            <StepLabel>{label}</StepLabel>
          </Step>
        ))}
      </Stepper>
    </Box>
  );
}
```
</details>

### Error step

```{r}
muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = "100%"),
    Stepper(
      activeStep = 1,
      lapply(seq_along(steps), function(i) {
        failed <- i == 2
        Step(
          StepLabel(
            optional = if (failed) Typography(variant = "caption", color = "error", "Alert message") else FALSE,
            error = failed,
            steps[i]
          )
        )
      })
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';
import Typography from '@mui/material/Typography';

const steps = ['Select campaign settings', 'Create an ad group', 'Create an ad'];

export default function HorizontalStepperWithError() {
  const isStepFailed = (step) => {
    return step === 1;
  };

  return (
    <Box sx={{ width: '100%' }}>
      <Stepper activeStep={1}>
        {steps.map((label, index) => {
          const labelProps = {};
          if (isStepFailed(index)) {
            labelProps.optional = (
              <Typography variant="caption" color="error">
                Alert message
              </Typography>
            );

            labelProps.error = true;
          }

          return (
            <Step key={label}>
              <StepLabel {...labelProps}>{label}</StepLabel>
            </Step>
          );
        })}
      </Stepper>
    </Box>
  );
}
```
</details>

### Customized horizontal stepper

Here is an example of customizing the component.
You can learn more about this in the [overrides documentation page](https://mui.com/material-ui/customization/how-to-customize/).

In React, the custom step icons are components passed to `slots$stepIcon`.
In R, the `icon` prop of `StepLabel()` accepts an element, which is built for
each step from its state (the active step is known when the page is built).
The `styled()` connectors become `sx` values:

```{r}
activeStep <- 1
gradient <- "linear-gradient( 136deg, rgb(242,113,33) 0%, rgb(233,64,87) 50%, rgb(138,35,135) 100%)"

QontoConnector <- StepConnector(
  sx = list(
    "&.MuiStepConnector-alternativeLabel" = list(top = 10, left = "calc(-50% + 16px)", right = "calc(50% + 16px)"),
    "&.Mui-active .MuiStepConnector-line, &.Mui-completed .MuiStepConnector-line" = list(borderColor = "#784af4"),
    "& .MuiStepConnector-line" = list(borderColor = "#eaeaf0", borderTopWidth = 3, borderRadius = 1)
  )
)
QontoStepIcon <- function(index) {
  Box(
    sx = list(color = if (index == activeStep) "#784af4" else "#eaeaf0", display = "flex", height = 22, alignItems = "center"),
    if (index < activeStep) {
      Icon("check", sx = list(color = "#784af4", zIndex = 1, fontSize = 18))
    } else {
      Box(sx = list(width = 8, height = 8, borderRadius = "50%", backgroundColor = "currentColor"))
    }
  )
}

ColorlibConnector <- StepConnector(
  sx = list(
    "&.MuiStepConnector-alternativeLabel" = list(top = 22),
    "&.Mui-active .MuiStepConnector-line, &.Mui-completed .MuiStepConnector-line" = list(
      backgroundImage = "linear-gradient( 95deg,rgb(242,113,33) 0%,rgb(233,64,87) 50%,rgb(138,35,135) 100%)"
    ),
    "& .MuiStepConnector-line" = list(height = 3, border = 0, backgroundColor = "#eaeaf0", borderRadius = 1)
  )
)
ColorlibStepIcon <- function(index) {
  Box(
    sx = list(
      backgroundColor = "#ccc", zIndex = 1, color = "#fff", width = 50, height = 50, display = "flex",
      borderRadius = "50%", justifyContent = "center", alignItems = "center",
      backgroundImage = if (index <= activeStep) gradient else "none",
      boxShadow = if (index == activeStep) "0 4px 10px 0 rgba(0,0,0,.25)" else "none"
    ),
    Icon(c("settings", "group_add", "video_label")[index + 1])
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Stack(
    sx = list(width = "100%"),
    spacing = 4,
    Stepper(
      alternativeLabel = TRUE,
      activeStep = activeStep,
      connector = QontoConnector,
      lapply(seq_along(steps) - 1, function(i) Step(StepLabel(icon = QontoStepIcon(i), steps[i + 1])))
    ),
    Stepper(
      alternativeLabel = TRUE,
      activeStep = activeStep,
      connector = ColorlibConnector,
      lapply(seq_along(steps) - 1, function(i) Step(StepLabel(icon = ColorlibStepIcon(i), steps[i + 1])))
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import PropTypes from 'prop-types';
import { styled } from '@mui/material/styles';
import Stack from '@mui/material/Stack';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';
import Check from '@mui/icons-material/Check';
import SettingsIcon from '@mui/icons-material/Settings';
import GroupAddIcon from '@mui/icons-material/GroupAdd';
import VideoLabelIcon from '@mui/icons-material/VideoLabel';
import StepConnector, { stepConnectorClasses } from '@mui/material/StepConnector';

const QontoConnector = styled(StepConnector)(({ theme }) => ({
  [`&.${stepConnectorClasses.alternativeLabel}`]: {
    top: 10,
    left: 'calc(-50% + 16px)',
    right: 'calc(50% + 16px)',
  },
  [`&.${stepConnectorClasses.active}`]: {
    [`& .${stepConnectorClasses.line}`]: {
      borderColor: '#784af4',
    },
  },
  [`&.${stepConnectorClasses.completed}`]: {
    [`& .${stepConnectorClasses.line}`]: {
      borderColor: '#784af4',
    },
  },
  [`& .${stepConnectorClasses.line}`]: {
    borderColor: '#eaeaf0',
    borderTopWidth: 3,
    borderRadius: 1,
    ...theme.applyStyles('dark', {
      borderColor: theme.palette.grey[800],
    }),
  },
}));

const QontoStepIconRoot = styled('div')(({ theme }) => ({
  color: '#eaeaf0',
  display: 'flex',
  height: 22,
  alignItems: 'center',
  '& .QontoStepIcon-completedIcon': {
    color: '#784af4',
    zIndex: 1,
    fontSize: 18,
  },
  '& .QontoStepIcon-circle': {
    width: 8,
    height: 8,
    borderRadius: '50%',
    backgroundColor: 'currentColor',
  },
  ...theme.applyStyles('dark', {
    color: theme.palette.grey[700],
  }),
  variants: [
    {
      props: ({ ownerState }) => ownerState.active,
      style: {
        color: '#784af4',
      },
    },
  ],
}));

function QontoStepIcon(props) {
  const { active, completed, className } = props;

  return (
    <QontoStepIconRoot ownerState={{ active }} className={className}>
      {completed ? (
        <Check className="QontoStepIcon-completedIcon" />
      ) : (
        <div className="QontoStepIcon-circle" />
      )}
    </QontoStepIconRoot>
  );
}

QontoStepIcon.propTypes = {
  /**
   * Whether this step is active.
   * @default false
   */
  active: PropTypes.bool,
  className: PropTypes.string,
  /**
   * Mark the step as completed. Is passed to child components.
   * @default false
   */
  completed: PropTypes.bool,
};

const ColorlibConnector = styled(StepConnector)(({ theme }) => ({
  [`&.${stepConnectorClasses.alternativeLabel}`]: {
    top: 22,
  },
  [`&.${stepConnectorClasses.active}`]: {
    [`& .${stepConnectorClasses.line}`]: {
      backgroundImage:
        'linear-gradient( 95deg,rgb(242,113,33) 0%,rgb(233,64,87) 50%,rgb(138,35,135) 100%)',
    },
  },
  [`&.${stepConnectorClasses.completed}`]: {
    [`& .${stepConnectorClasses.line}`]: {
      backgroundImage:
        'linear-gradient( 95deg,rgb(242,113,33) 0%,rgb(233,64,87) 50%,rgb(138,35,135) 100%)',
    },
  },
  [`& .${stepConnectorClasses.line}`]: {
    height: 3,
    border: 0,
    backgroundColor: '#eaeaf0',
    borderRadius: 1,
    ...theme.applyStyles('dark', {
      backgroundColor: theme.palette.grey[800],
    }),
  },
}));

const ColorlibStepIconRoot = styled('div')(({ theme }) => ({
  backgroundColor: '#ccc',
  zIndex: 1,
  color: '#fff',
  width: 50,
  height: 50,
  display: 'flex',
  borderRadius: '50%',
  justifyContent: 'center',
  alignItems: 'center',
  ...theme.applyStyles('dark', {
    backgroundColor: theme.palette.grey[700],
  }),
  variants: [
    {
      props: ({ ownerState }) => ownerState.active,
      style: {
        backgroundImage:
          'linear-gradient( 136deg, rgb(242,113,33) 0%, rgb(233,64,87) 50%, rgb(138,35,135) 100%)',
        boxShadow: '0 4px 10px 0 rgba(0,0,0,.25)',
      },
    },
    {
      props: ({ ownerState }) => ownerState.completed,
      style: {
        backgroundImage:
          'linear-gradient( 136deg, rgb(242,113,33) 0%, rgb(233,64,87) 50%, rgb(138,35,135) 100%)',
      },
    },
  ],
}));

function ColorlibStepIcon(props) {
  const { active, completed, className } = props;

  const icons = {
    1: <SettingsIcon />,
    2: <GroupAddIcon />,
    3: <VideoLabelIcon />,
  };

  return (
    <ColorlibStepIconRoot ownerState={{ completed, active }} className={className}>
      {icons[String(props.icon)]}
    </ColorlibStepIconRoot>
  );
}

ColorlibStepIcon.propTypes = {
  /**
   * Whether this step is active.
   * @default false
   */
  active: PropTypes.bool,
  className: PropTypes.string,
  /**
   * Mark the step as completed. Is passed to child components.
   * @default false
   */
  completed: PropTypes.bool,
  /**
   * The label displayed in the step icon.
   */
  icon: PropTypes.node,
};

const steps = ['Select campaign settings', 'Create an ad group', 'Create an ad'];

export default function CustomizedSteppers() {
  return (
    <Stack sx={{ width: '100%' }} spacing={4}>
      <Stepper alternativeLabel activeStep={1} connector={<QontoConnector />}>
        {steps.map((label) => (
          <Step key={label}>
            <StepLabel slots={{ stepIcon: QontoStepIcon }}>{label}</StepLabel>
          </Step>
        ))}
      </Stepper>
      <Stepper alternativeLabel activeStep={1} connector={<ColorlibConnector />}>
        {steps.map((label) => (
          <Step key={label}>
            <StepLabel slots={{ stepIcon: ColorlibStepIcon }}>{label}</StepLabel>
          </Step>
        ))}
      </Stepper>
    </Stack>
  );
}
```
</details>

## Vertical stepper

Vertical steppers are designed for narrow screen sizes. They are ideal for mobile. All the features of the horizontal stepper can be implemented.

```{r}
verticalSteps <- list(
  list(
    label = "Select campaign settings",
    description = "For each ad campaign that you create, you can control how much
    you're willing to spend on clicks and conversions, which networks
    and geographical locations you want your ads to show on, and more."
  ),
  list(
    label = "Create an ad group",
    description = "An ad group contains one or more ads which target a shared set of keywords."
  ),
  list(
    label = "Create an ad",
    description = "Try out different ad text to see what brings in the most customers,
    and learn how to enhance your ads using features like ad extensions.
    If you run into any problems with your ads, find out how to tell if
    they're running and how to resolve approval issues."
  )
)

verticalLinearStepper <- function(activeStep = 0, alternativeLabel = FALSE, route = "vertical-stepper") {
  Box(
    sx = list(maxWidth = 400),
    Stepper(
      activeStep = activeStep,
      orientation = "vertical",
      alternativeLabel = alternativeLabel,
      lapply(seq_along(verticalSteps) - 1, function(index) {
        step <- verticalSteps[[index + 1]]
        Step(
          StepLabel(
            optional = if (index == length(verticalSteps) - 1) Typography(variant = "caption", "Last step") else FALSE,
            step$label
          ),
          StepContent(
            Typography(step$description),
            Box(
              sx = list(mb = 2),
              Button(
                variant = "contained",
                href = sprintf("#/%s/%s", route, index + 1),
                sx = list(mt = 1, mr = 1),
                if (index == length(verticalSteps) - 1) "Finish" else "Continue"
              ),
              if (index != 0) Button(href = sprintf("#/%s/%s", route, index - 1), sx = list(mt = 1, mr = 1), "Back")
            )
          )
        )
      })
    ),
    if (activeStep == length(verticalSteps)) {
      Paper(
        square = TRUE,
        elevation = 0,
        sx = list(p = 3),
        Typography("All steps completed - you're finished"),
        Button(href = "#/", sx = list(mt = 1, mr = 1), "Reset")
      )
    }
  )
}

stepRoutes <- function(route, n, render) {
  c(
    lapply(seq_len(n), function(step) Route(path = paste0(route, "/", step), element = render(step))),
    list(Route(path = "*", element = render(0)))
  )
}

muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(stepRoutes("vertical-stepper", 3, verticalLinearStepper))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';
import StepContent from '@mui/material/StepContent';
import Button from '@mui/material/Button';
import Paper from '@mui/material/Paper';
import Typography from '@mui/material/Typography';

const steps = [
  {
    label: 'Select campaign settings',
    description: `For each ad campaign that you create, you can control how much
              you're willing to spend on clicks and conversions, which networks
              and geographical locations you want your ads to show on, and more.`,
  },
  {
    label: 'Create an ad group',
    description:
      'An ad group contains one or more ads which target a shared set of keywords.',
  },
  {
    label: 'Create an ad',
    description: `Try out different ad text to see what brings in the most customers,
              and learn how to enhance your ads using features like ad extensions.
              If you run into any problems with your ads, find out how to tell if
              they're running and how to resolve approval issues.`,
  },
];

export default function VerticalLinearStepper() {
  const [activeStep, setActiveStep] = React.useState(0);

  const handleNext = () => {
    setActiveStep((prevActiveStep) => prevActiveStep + 1);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const handleReset = () => {
    setActiveStep(0);
  };

  const previousActiveStepRef = React.useRef(activeStep);
  const continueButtonRef = React.useRef(null);
  const backButtonRef = React.useRef(null);
  const resetButtonRef = React.useRef(null);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    const previousActiveStep = previousActiveStepRef.current;
    previousActiveStepRef.current = activeStep;

    // If the user is going forward.
    if (previousActiveStep < activeStep) {
      if (activeStep === steps.length) {
        // If the user has completed all steps and hits "Finish", focus the "Reset" button.
        resetButtonRef.current.focus();
      } else {
        // Focus the "Continue" button otherwise.
        continueButtonRef.current.focus();
      }
      return;
    }
    // Otherwise, the user is going back.

    if (activeStep === 0) {
      // If the user hit "Back" on the second step, or hit "Reset", focus the "Continue" button.
      continueButtonRef.current.focus();
      return;
    }

    // Focus the "Back" button otherwise.
    backButtonRef.current.focus();
  }, [activeStep]);

  return (
    <Box sx={{ maxWidth: 400 }}>
      <Stepper activeStep={activeStep} orientation="vertical">
        {steps.map((step, index) => (
          <Step key={step.label}>
            <StepLabel
              optional={
                index === steps.length - 1 ? (
                  <Typography variant="caption">Last step</Typography>
                ) : null
              }
            >
              {step.label}
            </StepLabel>
            <StepContent>
              <Typography>{step.description}</Typography>
              <Box sx={{ mb: 2 }}>
                <Button
                  variant="contained"
                  onClick={handleNext}
                  sx={{ mt: 1, mr: 1 }}
                  ref={continueButtonRef}
                >
                  {index === steps.length - 1 ? 'Finish' : 'Continue'}
                </Button>
                {index !== 0 && (
                  <Button
                    onClick={handleBack}
                    sx={{ mt: 1, mr: 1 }}
                    ref={backButtonRef}
                  >
                    Back
                  </Button>
                )}
              </Box>
            </StepContent>
          </Step>
        ))}
      </Stepper>
      {activeStep === steps.length && (
        <Paper square elevation={0} sx={{ p: 3 }}>
          <Typography>All steps completed - you&apos;re finished</Typography>
          <Button onClick={handleReset} sx={{ mt: 1, mr: 1 }} ref={resetButtonRef}>
            Reset
          </Button>
        </Paper>
      )}
    </Box>
  );
}
```
</details>

### Alternative label

Use `alternativeLabel` prop on the vertical `Stepper` component to reverse the placement of the label and content.

```{r}
muiMaterialPage(
  CssBaseline(),
  RouterProvider(
    router = createHashRouter(stepRoutes("vertical-alternative-stepper", 3, function(step) {
      verticalLinearStepper(step, alternativeLabel = TRUE, route = "vertical-alternative-stepper")
    }))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Stepper from '@mui/material/Stepper';
import Step from '@mui/material/Step';
import StepLabel from '@mui/material/StepLabel';
import StepContent from '@mui/material/StepContent';
import Button from '@mui/material/Button';
import Paper from '@mui/material/Paper';
import Typography from '@mui/material/Typography';

const steps = [
  {
    label: 'Select campaign settings',
    description: `For each ad campaign that you create, you can control how much
              you're willing to spend on clicks and conversions, which networks
              and geographical locations you want your ads to show on, and more.`,
  },
  {
    label: 'Create an ad group',
    description:
      'An ad group contains one or more ads which target a shared set of keywords.',
  },
  {
    label: 'Create an ad',
    description: `Try out different ad text to see what brings in the most customers,
              and learn how to enhance your ads using features like ad extensions.
              If you run into any problems with your ads, find out how to tell if
              they're running and how to resolve approval issues.`,
  },
];

export default function VerticalLinearAlternativeLabelStepper() {
  const [activeStep, setActiveStep] = React.useState(0);

  const handleNext = () => {
    setActiveStep((prevActiveStep) => prevActiveStep + 1);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const handleReset = () => {
    setActiveStep(0);
  };

  return (
    <Box sx={{ maxWidth: 400 }}>
      <Stepper activeStep={activeStep} orientation="vertical" alternativeLabel>
        {steps.map((step, index) => (
          <Step key={step.label}>
            <StepLabel
              optional={
                index === steps.length - 1 ? (
                  <Typography variant="caption">Last step</Typography>
                ) : null
              }
            >
              {step.label}
            </StepLabel>
            <StepContent sx={{ textAlign: 'right' }}>
              <Typography>{step.description}</Typography>
              <Box sx={{ mb: 2 }}>
                <Button
                  variant="contained"
                  onClick={handleNext}
                  sx={{ mt: 1, mr: 1 }}
                >
                  {index === steps.length - 1 ? 'Finish' : 'Continue'}
                </Button>
                <Button
                  disabled={index === 0}
                  onClick={handleBack}
                  sx={{ mt: 1, mr: 1 }}
                >
                  Back
                </Button>
              </Box>
            </StepContent>
          </Step>
        ))}
      </Stepper>
      {activeStep === steps.length && (
        <Paper square elevation={0} sx={{ p: 3 }}>
          <Typography>All steps completed - you&apos;re finished</Typography>
          <Button onClick={handleReset} sx={{ mt: 1, mr: 1 }}>
            Reset
          </Button>
        </Paper>
      )}
    </Box>
  );
}
```
</details>

## Transition

`StepContent` uses [Collapse](transitions.html#collapse) by default.
Use `slots$transition` and `slotProps$transition` to replace it with another transition or to pass transition props.

## Performance

The content of a step is unmounted when closed.
If you need to make the content available to search engines or render expensive component trees inside your modal while optimizing for interaction responsiveness it might be a good idea to keep the step mounted with:

```r
StepContent(slotProps = list(transition = list(unmountOnExit = FALSE)))
```

## Mobile stepper

This component implements a compact stepper suitable for a mobile device. It has more limited functionality than the vertical stepper. See [mobile steps](https://m1.material.io/components/steppers.html#steppers-types-of-steps) for its inspiration.

The mobile stepper supports three variants to display progress through the available steps: text, dots, and progress.

### Text

The current step and total number of steps are displayed as text.

```{r}
mobileButtons <- function(route, activeStep, maxSteps) {
  list(
    nextButton = Button(
      size = "small",
      href = sprintf("#/%s/%s", route, activeStep + 1),
      disabled = activeStep == maxSteps - 1,
      "Next", Icon("keyboard_arrow_right")
    ),
    backButton = Button(
      size = "small",
      href = sprintf("#/%s/%s", route, activeStep - 1),
      disabled = activeStep == 0,
      Icon("keyboard_arrow_left"), "Back"
    )
  )
}

textMobileStepper <- function(activeStep) {
  buttons <- mobileButtons("text-mobile-stepper", activeStep, length(verticalSteps))
  Box(
    sx = list(maxWidth = 400, flexGrow = 1),
    Paper(
      square = TRUE,
      elevation = 0,
      sx = list(display = "flex", alignItems = "center", height = 50, pl = 2, bgcolor = "background.default"),
      Typography(verticalSteps[[activeStep + 1]]$label)
    ),
    Box(sx = list(height = 255, maxWidth = 400, width = "100%", p = 2), verticalSteps[[activeStep + 1]]$description),
    MobileStepper(
      variant = "text", steps = length(verticalSteps), position = "static", activeStep = activeStep,
      nextButton = buttons$nextButton, backButton = buttons$backButton
    )
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(stepRoutes("text-mobile-stepper", 2, textMobileStepper)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import { useTheme } from '@mui/material/styles';
import MobileStepper from '@mui/material/MobileStepper';
import Paper from '@mui/material/Paper';
import Typography from '@mui/material/Typography';
import Button from '@mui/material/Button';
import KeyboardArrowLeft from '@mui/icons-material/KeyboardArrowLeft';
import KeyboardArrowRight from '@mui/icons-material/KeyboardArrowRight';

const steps = [
  {
    label: 'Select campaign settings',
    description: `For each ad campaign that you create, you can control how much
              you're willing to spend on clicks and conversions, which networks
              and geographical locations you want your ads to show on, and more.`,
  },
  {
    label: 'Create an ad group',
    description:
      'An ad group contains one or more ads which target a shared set of keywords.',
  },
  {
    label: 'Create an ad',
    description: `Try out different ad text to see what brings in the most customers,
              and learn how to enhance your ads using features like ad extensions.
              If you run into any problems with your ads, find out how to tell if
              they're running and how to resolve approval issues.`,
  },
];

export default function TextMobileStepper() {
  const theme = useTheme();
  const [activeStep, setActiveStep] = React.useState(0);
  const maxSteps = steps.length;

  const handleNext = () => {
    setActiveStep((prevActiveStep) => prevActiveStep + 1);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const nextButtonRef = React.useRef(null);
  const backButtonRef = React.useRef(null);
  const previousActiveStepRef = React.useRef(activeStep);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    const previousActiveStep = previousActiveStepRef.current;
    previousActiveStepRef.current = activeStep;

    if (activeStep === 0 && previousActiveStep === 1) {
      // If the user is going back to the first step, focus the "Next" button.
      nextButtonRef.current.focus();
      return;
    }

    if (activeStep === maxSteps - 1 && previousActiveStep === maxSteps - 2) {
      // If the user is going to the last step, focus the "Back" button.
      backButtonRef.current.focus();
    }
  }, [activeStep, maxSteps]);

  return (
    <Box sx={{ maxWidth: 400, flexGrow: 1 }}>
      <Paper
        square
        elevation={0}
        sx={{
          display: 'flex',
          alignItems: 'center',
          height: 50,
          pl: 2,
          bgcolor: 'background.default',
        }}
      >
        <Typography>{steps[activeStep].label}</Typography>
      </Paper>
      <Box sx={{ height: 255, maxWidth: 400, width: '100%', p: 2 }}>
        {steps[activeStep].description}
      </Box>
      <MobileStepper
        variant="text"
        steps={maxSteps}
        position="static"
        activeStep={activeStep}
        nextButton={
          <Button
            size="small"
            onClick={handleNext}
            disabled={activeStep === maxSteps - 1}
            ref={nextButtonRef}
          >
            Next
            {theme.direction === 'rtl' ? (
              <KeyboardArrowLeft />
            ) : (
              <KeyboardArrowRight />
            )}
          </Button>
        }
        backButton={
          <Button
            size="small"
            onClick={handleBack}
            disabled={activeStep === 0}
            ref={backButtonRef}
          >
            {theme.direction === 'rtl' ? (
              <KeyboardArrowRight />
            ) : (
              <KeyboardArrowLeft />
            )}
            Back
          </Button>
        }
      />
    </Box>
  );
}
```
</details>

### Dots

Use dots when the number of steps is small.

```{r}
dotsMobileStepper <- function(activeStep) {
  buttons <- mobileButtons("dots-mobile-stepper", activeStep, 6)
  MobileStepper(
    variant = "dots", steps = 6, position = "static", activeStep = activeStep,
    sx = list(maxWidth = 400, flexGrow = 1),
    nextButton = buttons$nextButton, backButton = buttons$backButton
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(stepRoutes("dots-mobile-stepper", 5, dotsMobileStepper)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { useTheme } from '@mui/material/styles';
import MobileStepper from '@mui/material/MobileStepper';
import Button from '@mui/material/Button';
import KeyboardArrowLeft from '@mui/icons-material/KeyboardArrowLeft';
import KeyboardArrowRight from '@mui/icons-material/KeyboardArrowRight';

const steps = 6;

export default function DotsMobileStepper() {
  const theme = useTheme();
  const [activeStep, setActiveStep] = React.useState(0);

  const handleNext = () => {
    setActiveStep((prevActiveStep) => prevActiveStep + 1);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const nextButtonRef = React.useRef(null);
  const backButtonRef = React.useRef(null);
  const previousActiveStepRef = React.useRef(activeStep);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    const previousActiveStep = previousActiveStepRef.current;
    previousActiveStepRef.current = activeStep;

    if (activeStep === 0 && previousActiveStep === 1) {
      // If the user is going back to the first step, focus the "Next" button.
      nextButtonRef.current.focus();
      return;
    }
    if (activeStep === steps - 1 && previousActiveStep === steps - 2) {
      // If the user is going to the last step, focus the "Back" button.
      backButtonRef.current.focus();
    }
  }, [activeStep]);

  return (
    <MobileStepper
      variant="dots"
      steps={steps}
      position="static"
      activeStep={activeStep}
      sx={{ maxWidth: 400, flexGrow: 1 }}
      slotProps={{
        progress: {
          'aria-label': 'stepper dotted progress',
        },
      }}
      nextButton={
        <Button
          size="small"
          onClick={handleNext}
          disabled={activeStep === 5}
          ref={nextButtonRef}
        >
          Next
          {theme.direction === 'rtl' ? (
            <KeyboardArrowLeft />
          ) : (
            <KeyboardArrowRight />
          )}
        </Button>
      }
      backButton={
        <Button
          size="small"
          onClick={handleBack}
          disabled={activeStep === 0}
          ref={backButtonRef}
        >
          {theme.direction === 'rtl' ? (
            <KeyboardArrowRight />
          ) : (
            <KeyboardArrowLeft />
          )}
          Back
        </Button>
      }
    />
  );
}
```
</details>

### Progress

Use a progress bar when there are many steps, or if there are steps that need to be inserted during the process (based on responses to earlier steps).

```{r}
progressMobileStepper <- function(activeStep) {
  buttons <- mobileButtons("progress-mobile-stepper", activeStep, 6)
  MobileStepper(
    variant = "progress", steps = 6, position = "static", activeStep = activeStep,
    sx = list(maxWidth = 400, flexGrow = 1),
    nextButton = buttons$nextButton, backButton = buttons$backButton
  )
}

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  RouterProvider(router = createHashRouter(stepRoutes("progress-mobile-stepper", 5, progressMobileStepper)))
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import { useTheme } from '@mui/material/styles';
import MobileStepper from '@mui/material/MobileStepper';
import Button from '@mui/material/Button';
import KeyboardArrowLeft from '@mui/icons-material/KeyboardArrowLeft';
import KeyboardArrowRight from '@mui/icons-material/KeyboardArrowRight';

export default function ProgressMobileStepper() {
  const theme = useTheme();
  const [activeStep, setActiveStep] = React.useState(0);

  const handleNext = () => {
    setActiveStep((prevActiveStep) => prevActiveStep + 1);
  };

  const handleBack = () => {
    setActiveStep((prevActiveStep) => prevActiveStep - 1);
  };

  const nextButtonRef = React.useRef(null);
  const backButtonRef = React.useRef(null);
  const previousActiveStepRef = React.useRef(activeStep);

  // Manage focus when the active step changes.
  React.useEffect(() => {
    const previousActiveStep = previousActiveStepRef.current;

    if (activeStep === 0 && previousActiveStep === 1) {
      // If the user is going back to the first step, focus the "Next" button.
      nextButtonRef.current.focus();
    } else if (activeStep === 5 && previousActiveStep === 4) {
      // If the user is going to the last step, focus the "Back" button.
      backButtonRef.current.focus();
    }

    previousActiveStepRef.current = activeStep;
  }, [activeStep]);

  return (
    <MobileStepper
      variant="progress"
      steps={6}
      position="static"
      activeStep={activeStep}
      sx={{ maxWidth: 400, flexGrow: 1 }}
      slotProps={{
        progress: {
          'aria-label': 'stepper linear progress',
        },
      }}
      nextButton={
        <Button
          size="small"
          onClick={handleNext}
          disabled={activeStep === 5}
          ref={nextButtonRef}
        >
          Next
          {theme.direction === 'rtl' ? (
            <KeyboardArrowLeft />
          ) : (
            <KeyboardArrowRight />
          )}
        </Button>
      }
      backButton={
        <Button
          size="small"
          onClick={handleBack}
          disabled={activeStep === 0}
          ref={backButtonRef}
        >
          {theme.direction === 'rtl' ? (
            <KeyboardArrowRight />
          ) : (
            <KeyboardArrowLeft />
          )}
          Back
        </Button>
      }
    />
  );
}
```
</details>

## Steppers in Shiny apps

In a Shiny app, the server holds the active step, for example to validate each
step before moving on. Render the stepper with `renderUI()` and update the step
from `Button.shinyInput()` clicks:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

steps <- c("Select campaign settings", "Create an ad group", "Create an ad")

ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(width = "100%", p = 2),
    uiOutput("stepper"),
    Box(
      sx = list(display = "flex", flexDirection = "row", pt = 2),
      Button.shinyInput("back", "Back", color = "inherit"),
      Box(sx = list(flex = "1 1 auto")),
      Button.shinyInput("next", "Next")
    )
  )
)

server <- function(input, output, session) {
  activeStep <- reactiveVal(0)
  observeEvent(input$back, activeStep(max(activeStep() - 1, 0)))
  observeEvent(input$`next`, activeStep(min(activeStep() + 1, length(steps))))

  output$stepper <- renderUI({
    Stepper(activeStep = activeStep(), lapply(steps, function(label) Step(StepLabel(label))))
  })
}

shinyApp(ui, server)
```

Run a complete example with `muiMaterialExample("Stepper")`. `StepButton.shinyInput()`
reports clicks on a step label to the server, for non-linear steppers.
