---
title: "Tabs"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Tabs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Tabs organize content into separate views where only one view is visible at a time.

## Basic Usage

Use `TabContext.static()`, `TabList.static()`, and `TabPanel()` to create tabs. These are custom wrappers around [`@mui/lab`](https://mui.com/material-ui/about-the-lab/) tab components that manage tab state internally with `React.useState`, so they work everywhere -- in Shiny apps, Quarto documents, and plain R Markdown -- with no server logic required.

```{r eval=FALSE}
library(muiMaterial)

TabContext.static(
  value = "one",
  Box(
    sx = list(borderBottom = 1, borderColor = "divider"),
    TabList.static(
      Tab(label = "Item One", value = "one"),
      Tab(label = "Item Two", value = "two"),
      Tab(label = "Item Three", value = "three")
    )
  ),
  TabPanel(value = "one", "Content 1"),
  TabPanel(value = "two", "Content 2"),
  TabPanel(value = "three", "Content 3")
)
```

The `value` argument in `TabContext.static()` sets the initially selected tab. Each `Tab` and `TabPanel.static` share a matching `value` string to link headers to content.

### Quarto usage

In a Quarto document (`.qmd`), use the `.static` variants inside an R code chunk. No server function or `runtime: shiny` is needed.

````
---
title: "Tabs Example"
format: html
---

```{r}
library(muiMaterial)

TabContext.static(
  value = "one",
  Box(
    sx = list(borderBottom = 1, borderColor = "divider"),
    TabList.static(
      Tab(label = "Item One", value = "one"),
      Tab(label = "Item Two", value = "two"),
      Tab(label = "Item Three", value = "three")
    )
  ),
  TabPanel.static(value = "one", Typography("First tab content")),
  TabPanel.static(value = "two", Typography("Second tab content")),
  TabPanel.static(value = "three", Typography("Third tab content"))
)
```
````

### Why `@mui/lab`?

MUI provides tab components in two packages:

- **`@mui/material`**: `Tabs` and `Tab` -- base components that handle the tab header and selection state. Always controlled: requires a `value` prop and an `onChange` handler.
- **`@mui/lab`**: `TabContext`, `TabList`, and `TabPanel` -- higher-level wrappers that add automatic coordination between tab headers and content panels.

The `.static` variants wrap `@mui/lab` components and add internal `React.useState` management, which maps naturally to R's declarative UI model -- you declare the tabs and their content, and the switching is handled automatically.

## Advanced: `.shinyInput` Variants

The `.shinyInput` variants (`TabContext.shinyInput()`, `TabList.shinyInput()`, `TabPanel.shinyInput()`) expose the selected tab to the Shiny server as a reactive input. This is useful when:

- **Server-side reactions to tab changes**: You need the server to know which tab is active -- for example, to lazy-load expensive content, trigger data fetches, or log analytics only when a specific tab is viewed.
- **Programmatic tab switching**: The server can change the active tab via `updateTabContext.shinyInput()` in response to other events (e.g., completing a form step, receiving a notification).
- **Conditional server logic**: You want `observeEvent(input$tabList, ...)` to run code when the user switches tabs, such as refreshing a plot or resetting a form.

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  TabContext.shinyInput(
    inputId = "tabContext",
    value = "one",
    Box(
      sx = list(borderBottom = 1, borderColor = "divider"),
      TabList.shinyInput(
        inputId = "tabList",
        value = "one",
        Tab(label = "Item One", value = "one"),
        Tab(label = "Item Two", value = "two"),
        Tab(label = "Item Three", value = "three")
      )
    ),
    TabPanel.shinyInput(inputId = "panel1", value = "one", "Content 1"),
    TabPanel.shinyInput(inputId = "panel2", value = "two", "Content 2"),
    TabPanel.shinyInput(inputId = "panel3", value = "three", "Content 3")
  )
)

server <- function(input, output, session) {
  # Bridge TabList selection to TabContext
  observe({
    updateTabContext.shinyInput(
      inputId = "tabContext",
      value = input$tabList
    )
  })

  # React to tab changes on the server
  observeEvent(input$tabList, {
    message("User switched to tab: ", input$tabList)
  })
}

shinyApp(ui, server)
```

Because `TabContext` and `TabList` are separate Shiny inputs, clicking a tab updates `TabList`'s value but does not automatically propagate to `TabContext`. The `observe()` block bridges them: it watches `input$tabList` and calls `updateTabContext.shinyInput()` to match, which controls which `TabPanel` is visible.

Note that `TabPanel.shinyInput()` requires an `inputId` even though it never produces a value for the server. This is because `.shinyInput()` routes through the package's JavaScript bundle where it shares the same React tree as `TabContext`. The plain `TabPanel()` would resolve in a separate module context, breaking the React context connection.

### Comparison

| | `.static` | `.shinyInput` |
|---|---|---|
| Requires Shiny | No | Yes |
| Works in Quarto | Yes | No |
| Server can read selected tab | No | Yes (`input$tabList`) |
| Server can change selected tab | No | Yes (`updateTabContext.shinyInput`) |
| Requires observer | No | Yes |
| Requires `inputId` | No | Yes |
