---
title: "TextField"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TextField}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-text-field/).</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(muiMaterial)
```

# Text Field

Text Fields let users enter and edit text.

Text fields allow users to enter text into a UI. They typically appear in forms and dialogs.

## Basic TextField

The TextField wrapper component is a complete form control including a label, input, and help text.
It comes with three variants: outlined (default), filled, and standard.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function BasicTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& > :not(style)': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <TextField id="outlined-basic" label="Outlined" variant="outlined" />
      <TextField id="filled-basic" label="Filled" variant="filled" />
      <TextField id="standard-basic" label="Standard" variant="standard" />
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    component = "form",
    sx = list('& > :not(style)' = list(m = 1, width = '25ch')),
    noValidate = TRUE,
    autoComplete = "off",
    TextField.shinyInput(
      inputId = "outlined-basic",
      value = "",
      label = "Outlined",
      variant = "outlined"
    ),
    TextField.shinyInput(
      inputId = "filled-basic",
      value = "",
      label = "Filled",
      variant = "filled"
    ),
    TextField.shinyInput(
      inputId = "standard-basic",
      value = "",
      label = "Standard",
      variant = "standard"
    )
  )
)

server <- function(input, output, session) {
  # Access input values with input$`outlined-basic`, input$`filled-basic`, etc.
}

shinyApp(ui, server)
```

## Form props

Standard form attributes are supported, for example required, disabled, type, etc. as well as a helperText which is used to give context about a field's input, such as how the input will be used.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function FormPropsTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& .MuiTextField-root': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <div>
        <TextField
          required
          id="outlined-required"
          label="Required"
          defaultValue="Hello World"
        />
        <TextField
          disabled
          id="outlined-disabled"
          label="Disabled"
          defaultValue="Hello World"
        />
        <TextField
          id="outlined-password-input"
          label="Password"
          type="password"
          autoComplete="current-password"
        />
        <TextField
          id="outlined-read-only-input"
          label="Read Only"
          defaultValue="Hello World"
          slotProps={{
            input: {
              readOnly: true,
            },
          }}
        />
        <TextField
          id="outlined-number"
          label="Number"
          type="number"
          slotProps={{
            input: {
              step: 300,
            },
          }}
        />
        <TextField id="outlined-search" label="Search field" type="search" />
        <TextField
          id="outlined-helperText"
          label="Helper text"
          defaultValue="Default Value"
          helperText="Some important text"
        />
      </div>
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    component = "form",
    sx = list('& .MuiTextField-root' = list(m = 1, width = '25ch')),
    noValidate = TRUE,
    autoComplete = "off",
    div(
      TextField.shinyInput(
        inputId = "outlined-required",
        value = "Hello World",
        label = "Required",
        required = TRUE
      ),
      TextField.shinyInput(
        inputId = "outlined-disabled",
        value = "Hello World",
        label = "Disabled",
        disabled = TRUE
      ),
      TextField.shinyInput(
        inputId = "outlined-password-input",
        value = "",
        label = "Password",
        type = "password",
        autoComplete = "current-password"
      ),
      TextField.shinyInput(
        inputId = "outlined-read-only-input",
        value = "Hello World",
        label = "Read Only",
        slotProps = list(
          input = list(
            readOnly = TRUE
          )
        )
      ),
      TextField.shinyInput(
        inputId = "outlined-number",
        value = "",
        label = "Number",
        type = "number",
        slotProps = list(
          input = list(
            step = 300
          )
        )
      ),
      TextField.shinyInput(
        inputId = "outlined-search",
        value = "",
        label = "Search field",
        type = "search"
      ),
      TextField.shinyInput(
        inputId = "outlined-helperText",
        value = "Default Value",
        label = "Helper text",
        helperText = "Some important text"
      )
    )
  )
)

server <- function(input, output, session) {
  # Handle input values
}

shinyApp(ui, server)
```

## Controlling the HTML input

Use slotProps.htmlInput to pass attributes to the underlying `<input>` element.

<details>
<summary>JS code</summary>
```jsx
<TextField slotProps={{ htmlInput: { 'data-testid': '…' } }} />
```
</details>

```{r}
library(shiny)

ControlledInput <- function() {
  CssBaseline(
    TextField.shinyInput(
      inputId = "controlled-input",
      value = "",
      slotProps = list(
        htmlInput = list(
          'data-testid' = 'test-input'
        )
      )
    )
  )
}
```

The rendered HTML input will look like this:
```html
<input
  aria-invalid="false"
  class="MuiInputBase-input MuiOutlinedInput-input"
  type="text"
  data-testid="test-input"
/>
```

## Validation

The error prop toggles the error state.
The helperText prop can then be used to provide feedback to the user about the error.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function ValidationTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& .MuiTextField-root': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <div>
        <TextField
          error
          id="outlined-error"
          label="Error"
          defaultValue="Hello World"
        />
        <TextField
          error
          id="outlined-error-helper-text"
          label="Error"
          defaultValue="Hello World"
          helperText="Incorrect entry."
        />
      </div>
      <div>
        <TextField
          error
          id="filled-error"
          label="Error"
          defaultValue="Hello World"
          variant="filled"
        />
        <TextField
          error
          id="filled-error-helper-text"
          label="Error"
          defaultValue="Hello World"
          helperText="Incorrect entry."
          variant="filled"
        />
      </div>
      <div>
        <TextField
          error
          id="standard-error"
          label="Error"
          defaultValue="Hello World"
          variant="standard"
        />
        <TextField
          error
          id="standard-error-helper-text"
          label="Error"
          defaultValue="Hello World"
          helperText="Incorrect entry."
          variant="standard"
        />
      </div>
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    component = "form",
    sx = list('& .MuiTextField-root' = list(m = 1, width = '25ch')),
    noValidate = TRUE,
    autoComplete = "off",
    div(
      TextField.shinyInput(
        inputId = "outlined-error",
        value = "Hello World",
        label = "Error",
        error = TRUE
      ),
      TextField.shinyInput(
        inputId = "outlined-error-helper-text",
        value = "Hello World",
        label = "Error",
        error = TRUE,
        helperText = "Incorrect entry."
      )
    ),
    div(
      TextField.shinyInput(
        inputId = "filled-error",
        value = "Hello World",
        label = "Error",
        error = TRUE,
        variant = "filled"
      ),
      TextField.shinyInput(
        inputId = "filled-error-helper-text",
        value = "Hello World",
        label = "Error",
        error = TRUE,
        helperText = "Incorrect entry.",
        variant = "filled"
      )
    ),
    div(
      TextField.shinyInput(
        inputId = "standard-error",
        value = "Hello World",
        label = "Error",
        error = TRUE,
        variant = "standard"
      ),
      TextField.shinyInput(
        inputId = "standard-error-helper-text",
        value = "Hello World",
        label = "Error",
        error = TRUE,
        helperText = "Incorrect entry.",
        variant = "standard"
      )
    )
  )
)

server <- function(input, output, session) {
  # Handle validation logic
}

shinyApp(ui, server)
```

## Multiline

The multiline prop transforms the Text Field into a MUI Base Textarea Autosize element.
Unless the rows prop is set, the height of the text field dynamically matches its content.
You can use the minRows and maxRows props to bound it.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function MultilineTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& .MuiTextField-root': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <div>
        <TextField
          id="outlined-multiline-flexible"
          label="Multiline"
          multiline
          maxRows={4}
        />
        <TextField
          id="outlined-textarea"
          label="Multiline Placeholder"
          placeholder="Placeholder"
          multiline
        />
        <TextField
          id="outlined-multiline-static"
          label="Multiline"
          multiline
          rows={4}
          defaultValue="Default Value"
        />
      </div>
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    component = "form",
    sx = list('& .MuiTextField-root' = list(m = 1, width = '25ch')),
    noValidate = TRUE,
    autoComplete = "off",
    div(
      TextField.shinyInput(
        inputId = "outlined-multiline-flexible",
        value = "",
        label = "Multiline",
        multiline = TRUE,
        maxRows = 4
      ),
      TextField.shinyInput(
        inputId = "outlined-textarea",
        value = "",
        label = "Multiline Placeholder",
        placeholder = "Placeholder",
        multiline = TRUE
      ),
      TextField.shinyInput(
        inputId = "outlined-multiline-static",
        value = "Default Value",
        label = "Multiline",
        multiline = TRUE,
        rows = 4
      )
    )
  )
)

server <- function(input, output, session) {
  # Handle multiline input
}

shinyApp(ui, server)
```

## Icons

There are multiple ways to display an icon with a text field.

### Input Adornments

The main way is with an InputAdornment.
This can be used to add a prefix, a suffix, or an action to an input.
For instance, you can use an icon button to hide or reveal the password.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import IconButton from '@mui/material/IconButton';
import Input from '@mui/material/Input';
import FilledInput from '@mui/material/FilledInput';
import OutlinedInput from '@mui/material/OutlinedInput';
import InputLabel from '@mui/material/InputLabel';
import InputAdornment from '@mui/material/InputAdornment';
import FormHelperText from '@mui/material/FormHelperText';
import FormControl from '@mui/material/FormControl';
import TextField from '@mui/material/TextField';
import Visibility from '@mui/icons-material/Visibility';
import VisibilityOff from '@mui/icons-material/VisibilityOff';

export default function InputAdornments() {
  const [showPassword, setShowPassword] = React.useState(false);

  const handleClickShowPassword = () => setShowPassword((show) => !show);

  const handleMouseDownPassword = (event) => {
    event.preventDefault();
  };

  return (
    <Box sx={{ display: 'flex', flexWrap: 'wrap' }}>
      <div>
        <TextField
          label="With normal TextField"
          id="outlined-start-adornment"
          sx={{ m: 1, width: '25ch' }}
          slotProps={{
            input: {
              startAdornment: <InputAdornment position="start">kg</InputAdornment>,
            },
          }}
        />
        <FormControl sx={{ m: 1, width: '25ch' }} variant="outlined">
          <OutlinedInput
            id="outlined-adornment-weight"
            endAdornment={<InputAdornment position="end">kg</InputAdornment>}
            aria-describedby="outlined-weight-helper-text"
          />
          <FormHelperText id="outlined-weight-helper-text">Weight</FormHelperText>
        </FormControl>
        <FormControl sx={{ m: 1, width: '25ch' }} variant="outlined">
          <InputLabel htmlFor="outlined-adornment-password">Password</InputLabel>
          <OutlinedInput
            id="outlined-adornment-password"
            type={showPassword ? 'text' : 'password'}
            endAdornment={
              <InputAdornment position="end">
                <IconButton
                  aria-label="toggle password visibility"
                  onClick={handleClickShowPassword}
                  onMouseDown={handleMouseDownPassword}
                  edge="end"
                >
                  {showPassword ? <VisibilityOff /> : <Visibility />}
                </IconButton>
              </InputAdornment>
            }
            label="Password"
          />
        </FormControl>
        <FormControl fullWidth sx={{ m: 1 }}>
          <InputLabel htmlFor="outlined-adornment-amount">Amount</InputLabel>
          <OutlinedInput
            id="outlined-adornment-amount"
            startAdornment={<InputAdornment position="start">$</InputAdornment>}
            label="Amount"
          />
        </FormControl>
      </div>
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    sx = list(display = 'flex', flexWrap = 'wrap'),
    div(
      TextField.shinyInput(
        inputId = "outlined-start-adornment",
        value = "",
        label = "With normal TextField",
        sx = list(m = 1, width = '25ch'),
        slotProps = list(
          input = list(
            startAdornment = InputAdornment(position = "start", "kg")
          )
        )
      ),
      FormControl(
        sx = list(m = 1, width = '25ch'),
        variant = "outlined",
        OutlinedInput.shinyInput(
          inputId = "outlined-adornment-weight",
          endAdornment = InputAdornment(position = "end", "kg"),
          "aria-describedby" = "outlined-weight-helper-text"
        ),
        FormHelperText(id = "outlined-weight-helper-text", "Weight")
      ),
      FormControl(
        sx = list(m = 1, width = '25ch'),
        variant = "outlined",
        InputLabel(htmlFor = "outlined-adornment-password", "Password"),
        OutlinedInput.shinyInput(
          inputId = "outlined-adornment-password-input",
          type = "password",
          endAdornment = InputAdornment(
            position = "end",
            IconButton(
              "aria-label" = "toggle password visibility",
              id = "toggle-password",
              edge = "end",
              shiny::icon("eye")
            )
          ),
          label = "Password"
        )
      ),
      FormControl(
        fullWidth = TRUE,
        sx = list(m = 1),
        InputLabel(htmlFor = "outlined-adornment-amount", "Amount"),
        OutlinedInput.shinyInput(
          inputId = "outlined-adornment-amount",
          startAdornment = InputAdornment(position = "start", "$"),
          label = "Amount"
        )
      )
    )
  )
)

server <- function(input, output, session) {
  showPassword <- reactiveVal(FALSE)
  
  observeEvent(input$`toggle-password`, {
    showPassword(!showPassword())
    updateOutlinedInput.shinyInput(
      session,
      inputId = "outlined-adornment-password-input",
      type = if(showPassword()) "text" else "password",
      endAdornment = InputAdornment(
            position = "end",
            IconButton(
              "aria-label" = "toggle password visibility",
              id = "toggle-password",
              edge = "end",
              shiny::icon("eye-slash")
            )
          )
    )
  })
}

shinyApp(ui, server)
```

## Sizes

Fancy smaller inputs? Use the size prop.

The filled variant input height can be further reduced by rendering the label outside of it.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import TextField from '@mui/material/TextField';

export default function TextFieldSizes() {
  return (
    <div>
      <TextField
        label="Size"
        id="outlined-size-small"
        defaultValue="Small"
        size="small"
      />
      <TextField
        label="Size"
        id="outlined-size-normal"
        defaultValue="Normal"
      />
    </div>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    Box(
      p = 3,
      TextField.shinyInput(
        inputId = "outlined-size-small",
        value = "Small",
        label = "Size",
        size = "small"
      ),
      TextField.shinyInput(
        inputId = "outlined-size-normal",
        value = "Normal",
        label = "Size"
      )
    )
  )
}

server <- function(input, output, session) { }

shinyApp(ui, server)

```

## Margin

The margin prop can be used to alter the vertical spacing of the text field.
Using none (default) doesn't apply margins to the FormControl whereas dense and normal do.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function LayoutTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& .MuiTextField-root': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <div>
        <TextField
          label={'margin="none"'}
          id="margin-none"
        />
        <TextField
          label={'margin="dense"'}
          id="margin-dense"
          margin="dense"
        />
        <TextField
          label={'margin="normal"'}
          id="margin-normal"
          margin="normal"
        />
      </div>
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    Box(
      component = "form",
      sx = list('& .MuiTextField-root' = list(m = 1, width = '25ch')),
      noValidate = TRUE,
      autoComplete = "off",
      div(
        TextField.shinyInput(
          inputId = "margin-none",
          value = "",
          label = 'margin="none"'
        ),
        TextField.shinyInput(
          inputId = "margin-dense",
          value = "",
          label = 'margin="dense"',
          margin = "dense"
        ),
        TextField.shinyInput(
          inputId = "margin-normal",
          value = "",
          label = 'margin="normal"',
          margin = "normal"
        )
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Full width

fullWidth can be used to make the input take up the full width of its container.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function FullWidthTextField() {
  return (
    <Box
      sx={{ width: 500, maxWidth: '100%' }}
    >
      <TextField fullWidth label="fullWidth" id="fullWidth" />
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    Box(
      sx = list(width = 500, maxWidth = '100%', p = 3),
      TextField.shinyInput(
        inputId = "fullWidth",
        value = "",
        fullWidth = TRUE,
        label = "fullWidth"
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Uncontrolled vs. Controlled

The component can be controlled or uncontrolled.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function StateTextFields() {
  const [name, setName] = React.useState('Cat in the Hat');

  return (
    <Box
      component="form"
      sx={{ '& > :not(style)': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <TextField
        id="outlined-controlled"
        label="Controlled"
        value={name}
        onChange={(event: React.ChangeEvent<HTMLInputElement>) => {
          setName(event.target.value);
        }}
      />
      <TextField
        id="outlined-uncontrolled"
        label="Uncontrolled"
        defaultValue="foo"
      />
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- shinyMaterialUIPage(
  Box(
    component = "form",
    sx = list('& > :not(style)' = list(m = 1, width = '25ch')),
    noValidate = TRUE,
    autoComplete = "off",
    TextField.shinyInput(
      inputId = "outlined-controlled",
      value = "Cat in the Hat",
      label = "Controlled"
    ),
    TextField.shinyInput(
      inputId = "outlined-uncontrolled",
      value = "foo",
      label = "Uncontrolled"
    ),
    br(),
    textOutput("controlled_value")
  )
)

server <- function(input, output, session) {
  output$controlled_value <- renderText({
    paste("Controlled value:", input$`outlined-controlled`)
  })
}

shinyApp(ui, server)
```

## Components

TextField is composed of smaller components (
FormControl,
Input,
FilledInput,
InputLabel,
OutlinedInput,
and FormHelperText
) that you can leverage directly to significantly customize your form inputs.

You might also have noticed that some native HTML input properties are missing from the TextField component.
This is on purpose.
The component takes care of the most used properties.
Then, it's up to the user to use the underlying component shown in the following demo. Still, you can use slotProps.htmlInput (and slotProps.input, slotProps.inputLabel properties) if you want to avoid some boilerplate.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Input from '@mui/material/Input';
import InputLabel from '@mui/material/InputLabel';
import InputAdornment from '@mui/material/InputAdornment';
import FormControl from '@mui/material/FormControl';
import AccountCircle from '@mui/icons-material/AccountCircle';

export default function InputWithIcon() {
  return (
    <Box sx={{ '& > :not(style)': { m: 1 } }}>
      <FormControl variant="standard">
        <InputLabel htmlFor="input-with-icon-adornment">
          With a start adornment
        </InputLabel>
        <Input
          id="input-with-icon-adornment"
          startAdornment={
            <InputAdornment position="start">
              <AccountCircle />
            </InputAdornment>
          }
        />
      </FormControl>
    </Box>
  );
}
```
</details>

```{r}
library(shiny)

InputWithIcon <- function() {
  CssBaseline(
    Box(
      sx = list('& > :not(style)' = list(m = 1)),
      FormControl(
        variant = "standard",
        InputLabel(
          htmlFor = "input-with-icon-adornment",
          "With a start adornment"
        ),
        Input(
          id = "input-with-icon-adornment",
          startAdornment = InputAdornment(
            position = "start",
            shiny::icon("user-circle")
          )
        )
      )
    )
  )
}
InputWithIcon()
```

## Color

The color prop changes the highlight color of the text field when focused.

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import TextField from '@mui/material/TextField';

export default function ColorTextFields() {
  return (
    <Box
      component="form"
      sx={{ '& > :not(style)': { m: 1, width: '25ch' } }}
      noValidate
      autoComplete="off"
    >
      <TextField
        label="Outlined secondary"
        color="secondary"
        focused
      />
      <TextField
        label="Filled success"
        variant="filled"
        color="success"
        focused
      />
      <TextField
        label="Standard warning"
        variant="standard"
        color="warning"
        focused
      />
    </Box>
  );
}
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    Box(
      component = "form",
      sx = list('& > :not(style)' = list(m = 1, width = '25ch')),
      noValidate = TRUE,
      autoComplete = "off",
      TextField.shinyInput(
        inputId = "outlined-secondary",
        value = "",
        label = "Outlined secondary",
        color = "secondary",
        focused = TRUE
      ),
      TextField.shinyInput(
        inputId = "filled-success",
        value = "",
        label = "Filled success",
        variant = "filled",
        color = "success",
        focused = TRUE
      ),
      TextField.shinyInput(
        inputId = "standard-warning",
        value = "",
        label = "Standard warning",
        variant = "standard",
        color = "warning",
        focused = TRUE
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Customization

Here are some examples of customizing the component.
You can learn more about this in the overrides documentation page.

### Using the styled API

<details>
<summary>JS code</summary>
```jsx
import * as React from 'react';
import { alpha, styled } from '@mui/material/styles';
import InputBase from '@mui/material/InputBase';
import Box from '@mui/material/Box';
import InputLabel from '@mui/material/InputLabel';
import TextField from '@mui/material/TextField';
import FormControl from '@mui/material/FormControl';

const BootstrapInput = styled(InputBase)(({ theme }) => ({
  'label + &': {
    marginTop: theme.spacing(3),
  },
  '& .MuiInputBase-input': {
    borderRadius: 4,
    position: 'relative',
    backgroundColor: '#F3F6F9',
    border: '1px solid',
    borderColor: '#E0E3E7',
    fontSize: 16,
    width: 'auto',
    padding: '10px 12px',
    transition: theme.transitions.create([
      'border-color',
      'background-color',
      'box-shadow',
    ]),
    // Use the system font instead of the default Roboto font.
    fontFamily: [
      '-apple-system',
      'BlinkMacSystemFont',
      '"Segoe UI"',
      'Roboto',
      '"Helvetica Neue"',
      'Arial',
      'sans-serif',
      '"Apple Color Emoji"',
      '"Segoe UI Emoji"',
      '"Segoe UI Symbol"',
    ].join(','),
    '&:focus': {
      boxShadow: `${alpha(theme.palette.primary.main, 0.25)} 0 0 0 0.2rem`,
      borderColor: theme.palette.primary.main,
    },
  },
}));

export default function CustomizedInputsStyled() {
  return (
    <Box
      component="form"
      noValidate
      sx={{ display: 'flex', flexWrap: 'wrap' }}
    >
      <FormControl variant="standard">
        <InputLabel shrink htmlFor="bootstrap-input">
          Bootstrap
        </InputLabel>
        <BootstrapInput defaultValue="react-bootstrap" id="bootstrap-input" />
      </FormControl>
    </Box>
  );
}
```
</details>

```{r}
library(shiny)

CustomizedInputsStyled <- function() {
  CssBaseline(
    ThemeProvider(
      theme = list(
        components = list(
          MuiInputBase = list(
            styleOverrides = list(
              root = list(
                "&.bootstrap-input" = list(
                  'label + &' = list(
                    marginTop = "24px"
                  ),
                  '& .MuiInputBase-input' = list(
                    borderRadius = 4,
                    position = 'relative',
                    backgroundColor = '#F3F6F9',
                    border = '1px solid',
                    borderColor = '#E0E3E7',
                    fontSize = 16,
                    width = 'auto',
                    padding = '10px 12px',
                    transition = 'border-color 0.15s ease-in-out, background-color 0.15s ease-in-out, box-shadow 0.15s ease-in-out',
                    fontFamily = paste(
                      "-apple-system",
                      "BlinkMacSystemFont",
                      '"Segoe UI"',
                      "Roboto",
                      '"Helvetica Neue"',
                      "Arial",
                      "sans-serif",
                      '"Apple Color Emoji"',
                      '"Segoe UI Emoji"',
                      '"Segoe UI Symbol"',
                      sep = ","
                    ),
                    '&:focus' = list(
                      boxShadow = 'rgba(25, 118, 210, 0.25) 0 0 0 0.2rem',
                      borderColor = '#1976d2'
                    )
                  )
                )
              )
            )
          )
        )
      ),
      Box(
        component = "form",
        noValidate = TRUE,
        sx = list(display = 'flex', flexWrap = 'wrap'),
        FormControl(
          variant = "standard",
          InputLabel(
            shrink = TRUE,
            htmlFor = "bootstrap-input",
            "Bootstrap"
          ),
          InputBase(
            className = "bootstrap-input",
            defaultValue = "react-bootstrap",
            id = "bootstrap-input"
          )
        )
      )
    )
  )
}
```

### Using the theme style overrides API

Use the styleOverrides key to change any style injected by Material UI into the DOM.
See the theme style overrides documentation for further details.

Customization does not stop at CSS.
You can use composition to build custom components and give your app a unique feel.
Below is an example using the InputBase component, inspired by Google Maps.

🎨 If you are looking for inspiration, you can check [MUI Treasury's customization examples](https://mui-treasury.com/?path=/docs/textField-introduction--docs).

## Limitations

### Shrink

The input label "shrink" state isn't always correct. The input label is supposed to shrink as soon as the input is displaying something. In some circumstances, we can't determine the "shrink" state (number input, datetime input, Stripe input). You might notice an overlap.

To workaround the issue, you can force the "shrink" state of the label.

<details>
<summary>JS code</summary>
```jsx
<TextField slotProps={{ inputLabel: { shrink: true } }} />
```
</details>

or

<details>
<summary>JS code</summary>
```jsx
<InputLabel shrink>Count</InputLabel>
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    Box(
      p = 3,
      TextField.shinyInput(
        inputId = "shrink-example",
        value = "",
        slotProps = list(
          inputLabel = list(shrink = TRUE)
        )
      ),
      br(),
      FormControl(
        InputLabel(shrink = TRUE, "Count"),
        Input()
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

### Floating label

The floating label is absolutely positioned. It won't impact the layout of the page. Make sure that the input is larger than the label to display correctly.

### type="number"

If you need a text field with number validation, you can use MUI Base's Number Input instead.
You can follow this GitHub issue to track the progress of introducing the Number Input component to Material UI.

### Helper text

The helper text prop affects the height of the text field. If two text fields are placed side by side, one with a helper text and one without, they will have different heights. For example:

<details>
<summary>JS code</summary>
```jsx
<TextField
  helperText="Please enter your name"
  id="demo-helper-text-misaligned"
  label="Name"
/>
<TextField
  id="demo-helper-text-misaligned-no-helper"
  label="Name"
/>
```
</details>

This can be fixed by passing a space character to the helperText prop:

<details>
<summary>JS code</summary>
```jsx
<TextField
  helperText="Please enter your name"
  id="demo-helper-text-aligned"
  label="Name"
/>
<TextField
  helperText=" "
  id="demo-helper-text-aligned-no-helper"
  label="Name"
/>
```
</details>

```{r eval=FALSE}
library(shiny)

ui <- function() {
  CssBaseline(
    div(
      h4("Misaligned:"),
      TextField.shinyInput(
        inputId = "demo-helper-text-misaligned",
        value = "",
        helperText = "Please enter your name",
        label = "Name"
      ),
      TextField.shinyInput(
        inputId = "demo-helper-text-misaligned-no-helper",
        value = "",
        label = "Name"
      ),
      h4("Aligned:"),
      TextField.shinyInput(
        inputId = "demo-helper-text-aligned",
        value = "",
        helperText = "Please enter your name",
        label = "Name"
      ),
      TextField.shinyInput(
        inputId = "demo-helper-text-aligned-no-helper",
        value = "",
        helperText = " ",
        label = "Name"
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Integration with 3rd party input libraries

You can use third-party libraries to format an input.
You have to provide a custom implementation of the `<input>` element with the inputComponent property.

The following demo uses the react-imask and react-number-format libraries. The same concept could be applied to, for example react-stripe-element.

The provided input component should expose a ref with a value that implements the following interface:

```jsx
interface InputElement {
  focus(): void;
  value?: string;
}
```

<details>
<summary>JS code</summary>
```jsx
const MyInputComponent = React.forwardRef((props, ref) => {
  const { component: Component, ...other } = props;

  // implement `InputElement` interface
  React.useImperativeHandle(ref, () => ({
    focus: () => {
      // logic to focus the rendered component from 3rd party belongs here
    },
    // hiding the value e.g. react-stripe-elements
  }));

  // `Component` will be your `SomeThirdPartyComponent` from below
  return <Component {...other} />;
});

// usage
<TextField
  slotProps={{
    input: {
      inputComponent: MyInputComponent,
      inputProps: {
        component: SomeThirdPartyComponent,
      },
    },
  }}
/>;
```
</details>

## Accessibility

In order for the text field to be accessible, the input should be linked to the label and the helper text. The underlying DOM nodes should have this structure:

```html
<div class="form-control">
  <label for="my-input">Email address</label>
  <input id="my-input" aria-describedby="my-helper-text" />
  <span id="my-helper-text">We'll never share your email.</span>
</div>
```

- If you are using the TextField component, you just have to provide a unique id unless you're using the TextField only client-side. Until the UI is hydrated TextField without an explicit id will not have associated labels.
- If you are composing the component:

<details>
<summary>JS code</summary>
```jsx
<FormControl>
  <InputLabel htmlFor="my-input">Email address</InputLabel>
  <Input id="my-input" aria-describedby="my-helper-text" />
  <FormHelperText id="my-helper-text">We'll never share your email.</FormHelperText>
</FormControl>
```
</details>

```{r eval=FALSE}
ui <- function() {
  CssBaseline(
    FormControl(
      InputLabel(htmlFor = "my-input", "Email address"),
      Input(
        id = "my-input",
        "aria-describedby" = "my-helper-text"
      ),
      FormHelperText(
        id = "my-helper-text",
        "We'll never share your email."
      )
    )
  )
}

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```
