---
title: "Timeline"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Timeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-timeline/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
```

# Timeline

The timeline displays a list of events in chronological order.

<div class="alert alert-info">
This component is not documented in the [Material Design guidelines](https://m2.material.io/), but it is available in Material UI.
The Timeline components come from [`@mui/lab`](https://mui.com/material-ui/about-the-lab/), which is published on the MUI beta channel.
</div>

The demos use this helper to build one `TimelineItem()` per event, with a
connector after every dot but the last:

```{r}
timelineItems <- function(labels, dot = function(i) TimelineDot(), opposite = NULL) {
  lapply(seq_along(labels), function(i) {
    TimelineItem(
      if (!is.null(opposite)) opposite(i),
      TimelineSeparator(dot(i), if (i < length(labels)) TimelineConnector()),
      TimelineContent(labels[i])
    )
  })
}
```

## Basic timeline

A basic timeline showing list of events.

```{r}
muiMaterialPage(CssBaseline(), Timeline(timelineItems(c("Eat", "Code", "Sleep"))))
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function BasicTimeline() {
  return (
    <Timeline>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Left-positioned timeline

The main content of the timeline can be positioned on the left side relative to the time axis.

```{r}
muiMaterialPage(CssBaseline(), Timeline(position = "left", timelineItems(c("Eat", "Code", "Sleep", "Repeat"))))
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function LeftPositionedTimeline() {
  return (
    <Timeline position="left">
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Repeat</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Alternating timeline

The timeline can display the events on alternating sides.

```{r}
muiMaterialPage(CssBaseline(), Timeline(position = "alternate", timelineItems(c("Eat", "Code", "Sleep", "Repeat"))))
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function AlternateTimeline() {
  return (
    <Timeline position="alternate">
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Repeat</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Reverse Alternating timeline

The timeline can display the events on alternating sides in reverse order.

```{r}
muiMaterialPage(CssBaseline(), Timeline(position = "alternate-reverse", timelineItems(c("Eat", "Code", "Sleep", "Repeat"))))
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function AlternateReverseTimeline() {
  return (
    <Timeline position="alternate-reverse">
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Repeat</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Color

The `TimelineDot` can appear in different colors from theme palette.

```{r}
muiMaterialPage(
  CssBaseline(),
  Timeline(
    position = "alternate",
    timelineItems(c("Secondary", "Success"), dot = function(i) TimelineDot(color = c("secondary", "success")[i]))
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function ColorsTimeline() {
  return (
    <Timeline position="alternate">
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot color="secondary" />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Secondary</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot color="success" />
        </TimelineSeparator>
        <TimelineContent>Success</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Outlined

```{r}
muiMaterialPage(
  CssBaseline(),
  Timeline(
    position = "alternate",
    timelineItems(
      c("Eat", "Code", "Sleep", "Repeat"),
      dot = function(i) TimelineDot(variant = "outlined", color = c("grey", "primary", "secondary", "grey")[i])
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function OutlinedTimeline() {
  return (
    <Timeline position="alternate">
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot variant="outlined" />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot variant="outlined" color="primary" />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot variant="outlined" color="secondary" />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot variant="outlined" />
        </TimelineSeparator>
        <TimelineContent>Repeat</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Opposite content

The timeline can display content on opposite sides.

```{r}
times <- c("09:30 am", "10:00 am", "12:00 am", "9:00 am")

muiMaterialPage(
  CssBaseline(),
  Timeline(
    position = "alternate",
    timelineItems(
      c("Eat", "Code", "Sleep", "Repeat"),
      opposite = function(i) TimelineOppositeContent(sx = list(color = "text.secondary"), times[i])
    )
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';
import TimelineOppositeContent from '@mui/lab/TimelineOppositeContent';

export default function OppositeContentTimeline() {
  return (
    <Timeline position="alternate">
      <TimelineItem>
        <TimelineOppositeContent
          sx={{
            color: 'text.secondary',
          }}
        >
          09:30 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent
          sx={{
            color: 'text.secondary',
          }}
        >
          10:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent
          sx={{
            color: 'text.secondary',
          }}
        >
          12:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Sleep</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent
          sx={{
            color: 'text.secondary',
          }}
        >
          9:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Repeat</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Customization

Here is an example of customizing the component.
You can learn more about this in the [overrides documentation page](https://mui.com/material-ui/customization/how-to-customize/).

```{r}
events <- list(
  list(time = "9:30 am", title = "Eat", text = "Because you need strength", icon = "fastfood", dot = list()),
  list(time = "10:00 am", title = "Code", text = "Because it's awesome!", icon = "laptop_mac", dot = list(color = "primary")),
  list(title = "Sleep", text = "Because you need rest", icon = "hotel", dot = list(color = "primary", variant = "outlined")),
  list(title = "Repeat", text = "Because this is the life you love!", icon = "repeat", dot = list(color = "secondary"))
)

muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Timeline(
    position = "alternate",
    lapply(seq_along(events), function(i) {
      e <- events[[i]]
      TimelineItem(
        if (!is.null(e$time)) {
          TimelineOppositeContent(
            sx = list(m = "auto 0"),
            align = if (i == 1) "right" else "left",
            variant = "body2",
            color = "text.secondary",
            e$time
          )
        },
        TimelineSeparator(
          TimelineConnector(sx = if (i == 4) list(bgcolor = "secondary.main")),
          do.call(TimelineDot, c(e$dot, list(Icon(e$icon)))),
          TimelineConnector(sx = if (i == 3) list(bgcolor = "secondary.main"))
        ),
        TimelineContent(
          sx = list(py = "12px", px = 2),
          Typography(variant = "h6", component = "span", e$title),
          Typography(e$text)
        )
      )
    })
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineOppositeContent from '@mui/lab/TimelineOppositeContent';
import TimelineDot from '@mui/lab/TimelineDot';
import FastfoodIcon from '@mui/icons-material/Fastfood';
import LaptopMacIcon from '@mui/icons-material/LaptopMac';
import HotelIcon from '@mui/icons-material/Hotel';
import RepeatIcon from '@mui/icons-material/Repeat';
import Typography from '@mui/material/Typography';

export default function CustomizedTimeline() {
  return (
    <Timeline position="alternate">
      <TimelineItem>
        <TimelineOppositeContent
          align="right"
          variant="body2"
          sx={{
            color: 'text.secondary',
            m: 'auto 0',
          }}
        >
          9:30 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineConnector />
          <TimelineDot>
            <FastfoodIcon />
          </TimelineDot>
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent sx={{ py: '12px', px: 2 }}>
          <Typography variant="h6" component="span">
            Eat
          </Typography>
          <Typography>Because you need strength</Typography>
        </TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent
          variant="body2"
          sx={{
            color: 'text.secondary',
            m: 'auto 0',
          }}
        >
          10:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineConnector />
          <TimelineDot color="primary">
            <LaptopMacIcon />
          </TimelineDot>
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent sx={{ py: '12px', px: 2 }}>
          <Typography variant="h6" component="span">
            Code
          </Typography>
          <Typography>Because it&apos;s awesome!</Typography>
        </TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineConnector />
          <TimelineDot color="primary" variant="outlined">
            <HotelIcon />
          </TimelineDot>
          <TimelineConnector sx={{ bgcolor: 'secondary.main' }} />
        </TimelineSeparator>
        <TimelineContent sx={{ py: '12px', px: 2 }}>
          <Typography variant="h6" component="span">
            Sleep
          </Typography>
          <Typography>Because you need rest</Typography>
        </TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineConnector sx={{ bgcolor: 'secondary.main' }} />
          <TimelineDot color="secondary">
            <RepeatIcon />
          </TimelineDot>
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent sx={{ py: '12px', px: 2 }}>
          <Typography variant="h6" component="span">
            Repeat
          </Typography>
          <Typography>Because this is the life you love!</Typography>
        </TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Alignment

There are different ways in which a Timeline can be placed within the container.

You can do it by overriding the styles.

A Timeline centers itself in the container by default.

The demos below show how to adjust the relative width of the left and right sides of a Timeline:

### Left-aligned

```{r}
twoTimes <- function(i) TimelineOppositeContent(color = "textSecondary", times[i])

muiMaterialPage(
  CssBaseline(),
  Timeline(
    sx = list("& .MuiTimelineOppositeContent-root" = list(flex = 0.2)),
    timelineItems(c("Eat", "Code"), opposite = twoTimes)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';
import TimelineOppositeContent, {
  timelineOppositeContentClasses,
} from '@mui/lab/TimelineOppositeContent';

export default function LeftAlignedTimeline() {
  return (
    <Timeline
      sx={{
        [`& .${timelineOppositeContentClasses.root}`]: {
          flex: 0.2,
        },
      }}
    >
      <TimelineItem>
        <TimelineOppositeContent color="textSecondary">
          09:30 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent color="textSecondary">
          10:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

### Right-aligned

```{r}
muiMaterialPage(
  CssBaseline(),
  Timeline(
    sx = list("& .MuiTimelineContent-root" = list(flex = 0.2)),
    timelineItems(c("Eat", "Code"), opposite = twoTimes)
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent, { timelineContentClasses } from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';
import TimelineOppositeContent from '@mui/lab/TimelineOppositeContent';

export default function RightAlignedTimeline() {
  return (
    <Timeline
      sx={{
        [`& .${timelineContentClasses.root}`]: {
          flex: 0.2,
        },
      }}
    >
      <TimelineItem>
        <TimelineOppositeContent color="textSecondary">
          09:30 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineOppositeContent color="textSecondary">
          10:00 am
        </TimelineOppositeContent>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

### Left-aligned with no opposite content

```{r}
muiMaterialPage(
  CssBaseline(),
  Timeline(
    sx = list("& .MuiTimelineItem-root:before" = list(flex = 0, padding = 0)),
    timelineItems(c("Eat", "Code"))
  )
)
```

<details><summary>JS code</summary>
```jsx
import Timeline from '@mui/lab/Timeline';
import TimelineItem, { timelineItemClasses } from '@mui/lab/TimelineItem';
import TimelineSeparator from '@mui/lab/TimelineSeparator';
import TimelineConnector from '@mui/lab/TimelineConnector';
import TimelineContent from '@mui/lab/TimelineContent';
import TimelineDot from '@mui/lab/TimelineDot';

export default function NoOppositeContent() {
  return (
    <Timeline
      sx={{
        [`& .${timelineItemClasses.root}:before`]: {
          flex: 0,
          padding: 0,
        },
      }}
    >
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
          <TimelineConnector />
        </TimelineSeparator>
        <TimelineContent>Eat</TimelineContent>
      </TimelineItem>
      <TimelineItem>
        <TimelineSeparator>
          <TimelineDot />
        </TimelineSeparator>
        <TimelineContent>Code</TimelineContent>
      </TimelineItem>
    </Timeline>
  );
}
```
</details>

## Timelines from data

In R, a timeline is often built from a data frame of events:

```{r}
milestones <- data.frame(
  date = c("2026-01-15", "2026-03-02", "2026-05-20"),
  event = c("Project kick-off", "First prototype", "Public release")
)

muiMaterialPage(
  CssBaseline(),
  Timeline(
    sx = list("& .MuiTimelineOppositeContent-root" = list(flex = 0.3)),
    timelineItems(
      milestones$event,
      dot = function(i) TimelineDot(color = if (i == nrow(milestones)) "primary" else "grey"),
      opposite = function(i) TimelineOppositeContent(color = "textSecondary", milestones$date[i])
    )
  )
)
```
