---
title: "Transitions"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Transitions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/transitions/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(reactRouter)
library(shiny)
```

# Transitions

Transitions help to make a UI expressive and easy to use.

Material UI provides transitions that can be used to introduce some basic [motion](https://m2.material.io/design/motion/) to your applications.

<div class="alert alert-secondary">
**How the live demos on this page work.** A transition animates when its `in`
prop changes. In React, `in` comes from a `useState()` hook toggled by the
"Show" switch. Here, the switch navigates between two
[reactRouter](https://felixluginbuhl.com/reactRouter/) routes (for example
`#/transition/collapse` and `#/`), which render the demo with `in = TRUE` or
`in = FALSE`. Changing route replaces the elements, so `appear = TRUE` animates
the entering; the exit is immediate. In a Shiny app, `renderReact()` updates
`in` on the same element, so both directions are animated: see
[Transitions in Shiny apps](#transitions-in-shiny-apps).
</div>

```{r}
# The triangle icon of the MUI demos
icon <- Paper(
  sx = list(m = 1, width = 100, height = 100),
  elevation = 4,
  tags$svg(
    width = "100", height = "100",
    Box(component = "polygon", points = "0,100 50,00, 100,100", sx = list(fill = "common.white", stroke = "divider", strokeWidth = 1))
  )
)

# A demo with a "Show" switch: `render(checked)` builds the content
transitionDemo <- function(name, render) {
  showSwitch <- function(checked) {
    FormControlLabel(
      control = Switch(
        checked = checked,
        onChange = JS(sprintf("() => { window.location.hash = '%s'; }", if (checked) "/" else paste0("/transition/", name)))
      ),
      label = "Show"
    )
  }
  RouterProvider(
    router = createHashRouter(
      Route(path = paste0("transition/", name), element = render(TRUE, showSwitch(TRUE))),
      Route(path = "*", element = render(FALSE, showSwitch(FALSE)))
    )
  )
}
```

## Collapse

Expand from the start edge of the child element.
Use the `orientation` prop if you need a horizontal collapse.
The `collapsedSize` prop can be used to set the minimum width/height when not expanded.

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("collapse", function(checked, toggle) {
    Box(
      sx = list(height = 300),
      toggle,
      Box(
        sx = list("& > :not(style)" = list(display = "flex", justifyContent = "space-around", height = 120, width = 250)),
        Box(
          Collapse(`in` = checked, appear = TRUE, icon),
          Collapse(`in` = checked, appear = TRUE, collapsedSize = 40, icon)
        ),
        Box(
          Box(sx = list(width = "50%"), Collapse(orientation = "horizontal", `in` = checked, appear = TRUE, icon)),
          Box(sx = list(width = "50%"), Collapse(orientation = "horizontal", `in` = checked, appear = TRUE, collapsedSize = 40, icon))
        )
      )
    )
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Collapse from '@mui/material/Collapse';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SimpleCollapse() {
  const [checked, setChecked] = React.useState(false);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box sx={{ height: 300 }}>
      <FormControlLabel
        control={<Switch checked={checked} onChange={handleChange} />}
        label="Show"
      />
      <Box
        sx={{
          '& > :not(style)': {
            display: 'flex',
            justifyContent: 'space-around',
            height: 120,
            width: 250,
          },
        }}
      >
        <div>
          <Collapse in={checked}>{icon}</Collapse>
          <Collapse in={checked} collapsedSize={40}>
            {icon}
          </Collapse>
        </div>
        <div>
          <Box sx={{ width: '50%' }}>
            <Collapse orientation="horizontal" in={checked}>
              {icon}
            </Collapse>
          </Box>
          <Box sx={{ width: '50%' }}>
            <Collapse orientation="horizontal" in={checked} collapsedSize={40}>
              {icon}
            </Collapse>
          </Box>
        </div>
      </Box>
    </Box>
  );
}
```
</details>

## Fade

Fade in from transparent to opaque.

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("fade", function(checked, toggle) {
    Box(sx = list(height = 180), toggle, Box(sx = list(display = "flex"), Fade(`in` = checked, appear = TRUE, icon)))
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Fade from '@mui/material/Fade';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SimpleFade() {
  const [checked, setChecked] = React.useState(false);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box sx={{ height: 180 }}>
      <FormControlLabel
        control={<Switch checked={checked} onChange={handleChange} />}
        label="Show"
      />
      <Box sx={{ display: 'flex' }}>
        <Fade in={checked}>{icon}</Fade>
      </Box>
    </Box>
  );
}
```
</details>

## Grow

Expands outwards from the center of the child element, while also fading in from transparent to opaque.

The second example demonstrates how to change the `transform-origin`, and conditionally applies
the `timeout` prop to change the entry speed.

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("grow", function(checked, toggle) {
    Box(
      sx = list(height = 180),
      toggle,
      Box(
        sx = list(display = "flex"),
        Grow(`in` = checked, appear = TRUE, icon),
        # Conditionally applies the timeout prop to change the entry speed.
        do.call(Grow, c(
          list(`in` = checked, appear = TRUE, style = list(transformOrigin = "0 0 0"), icon),
          if (checked) list(timeout = 1000)
        ))
      )
    )
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Grow from '@mui/material/Grow';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SimpleGrow() {
  const [checked, setChecked] = React.useState(false);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box sx={{ height: 180 }}>
      <FormControlLabel
        control={<Switch checked={checked} onChange={handleChange} />}
        label="Show"
      />
      <Box sx={{ display: 'flex' }}>
        <Grow in={checked}>{icon}</Grow>
        {/* Conditionally applies the timeout prop to change the entry speed. */}
        <Grow
          in={checked}
          style={{ transformOrigin: '0 0 0' }}
          {...(checked ? { timeout: 1000 } : {})}
        >
          {icon}
        </Grow>
      </Box>
    </Box>
  );
}
```
</details>

## Slide

Slide in from the edge of the screen.
The `direction` prop controls which edge of the screen the transition starts from.

The `mountOnEnter` prop prevents the child component from being mounted
until `in` is `TRUE`.
This prevents the relatively positioned component from scrolling into view
from its off-screen position.
Similarly, the `unmountOnExit` prop removes the component from the DOM after it has been transition off-screen.

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("slide", function(checked, toggle) {
    Box(
      sx = list(height = 180, width = 130, position = "relative", zIndex = 1),
      toggle,
      Slide(direction = "up", `in` = checked, appear = TRUE, mountOnEnter = TRUE, unmountOnExit = TRUE, icon)
    )
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Slide from '@mui/material/Slide';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SimpleSlide() {
  const [checked, setChecked] = React.useState(false);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box sx={{ height: 180, width: 130, position: 'relative', zIndex: 1 }}>
      <FormControlLabel
        control={<Switch checked={checked} onChange={handleChange} />}
        label="Show"
      />
      <Slide direction="up" in={checked} mountOnEnter unmountOnExit>
        {icon}
      </Slide>
    </Box>
  );
}
```
</details>

### Slide relative to a container

The Slide component also accepts `container` prop, which is a reference to a DOM node.
If this prop is set, the Slide component will slide from the edge of that DOM node.
In R, the node is looked up by `id` with a `JS()` function:

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("slide-container", function(checked, toggle) {
    Box(
      sx = list(width = 240, borderRadius = 2, border = "1px solid", borderColor = "divider", backgroundColor = "background.default"),
      Box(
        id = "slide-container",
        sx = list(p = 2, height = 200, overflow = "hidden"),
        toggle,
        Slide(`in` = checked, appear = TRUE, container = JS("() => document.getElementById('slide-container')"), icon)
      )
    )
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Slide from '@mui/material/Slide';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SlideFromContainer() {
  const [checked, setChecked] = React.useState(false);
  const containerRef = React.useRef(null);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box
      sx={{
        width: 240,
        borderRadius: 2,
        border: '1px solid',
        borderColor: 'divider',
        backgroundColor: 'background.default',
      }}
    >
      <Box sx={{ p: 2, height: 200, overflow: 'hidden' }} ref={containerRef}>
        <FormControlLabel
          control={<Switch checked={checked} onChange={handleChange} />}
          label="Show from target"
        />
        <Slide in={checked} container={containerRef.current}>
          {icon}
        </Slide>
      </Box>
    </Box>
  );
}
```
</details>

## Zoom

Expand outwards from the center of the child element.

This example also demonstrates how to delay the enter transition.

```{r}
muiMaterialPage(
  CssBaseline(),
  transitionDemo("zoom", function(checked, toggle) {
    Box(
      sx = list(height = 180),
      toggle,
      Box(
        sx = list(display = "flex"),
        Zoom(`in` = checked, appear = TRUE, icon),
        Zoom(`in` = checked, appear = TRUE, style = list(transitionDelay = if (checked) "500ms" else "0ms"), icon)
      )
    )
  })
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Box from '@mui/material/Box';
import Switch from '@mui/material/Switch';
import Paper from '@mui/material/Paper';
import Zoom from '@mui/material/Zoom';
import FormControlLabel from '@mui/material/FormControlLabel';

const icon = (
  <Paper sx={{ m: 1, width: 100, height: 100 }} elevation={4}>
    <svg width="100" height="100">
      <Box
        component="polygon"
        points="0,100 50,00, 100,100"
        sx={(theme) => ({
          fill: theme.palette.common.white,
          stroke: theme.palette.divider,
          strokeWidth: 1,
        })}
      />
    </svg>
  </Paper>
);

export default function SimpleZoom() {
  const [checked, setChecked] = React.useState(false);

  const handleChange = () => {
    setChecked((prev) => !prev);
  };

  return (
    <Box sx={{ height: 180 }}>
      <FormControlLabel
        control={<Switch checked={checked} onChange={handleChange} />}
        label="Show"
      />
      <Box sx={{ display: 'flex' }}>
        <Zoom in={checked}>{icon}</Zoom>
        <Zoom in={checked} style={{ transitionDelay: checked ? '500ms' : '0ms' }}>
          {icon}
        </Zoom>
      </Box>
    </Box>
  );
}
```
</details>

## Reduced motion

Transitions can opt in to reduced-motion support through the theme:

```r
ThemeProvider(theme = list(motion = list(reducedMotion = "system")), ...)
```

When enabled, Material UI transition components preserve lifecycle callbacks and mount/unmount
behavior. When reduced motion is active, entering content appears in its final state and exiting
content disappears without the normal animation. Components still run the same callbacks, so code
that waits for `onEntered` or `onExited` continues to work.

Use `disablePrefersReducedMotion` only when a specific transition should intentionally keep its
normal motion:

```r
Fade(`in` = TRUE, disablePrefersReducedMotion = TRUE, Box())
```

## Child requirement

- **Forward the style**: To better support server rendering, Material UI provides a `style` prop to the children of some transition components (Fade, Grow, Zoom, Slide).
  The `style` prop must be applied to the DOM for the animation to work as expected.
- **Forward the ref**: The transition components require the first child element to forward its ref to the DOM node.
- **Single element**: The transition components require only one child element.

All muiMaterial components and HTML tags satisfy the first two requirements.
Wrap several elements in a single `Box()` or `tags$div()`.

## TransitionGroup

In React, the [`TransitionGroup`](https://reactcommunity.org/react-transition-group/transition-group/)
component of react-transition-group animates components when they are mounted or
unmounted, for example items added to or removed from a list. It is not
exported by muiMaterial.

## Transition slots

Many Material UI components use these transitions internally.
Use `slots$transition` and `slotProps$transition` to customize the default transition.
A component is expected, so pass it as a JavaScript reference with `JS()`, for example
`slots = list(transition = JS("jsmodule['@mui/material'].Fade"))`.
It should respect the following conditions:

- Accepts an `in` prop. This corresponds to the open/close state.
- Call the `onEnter` callback prop when the enter transition starts.
- Call the `onExited` callback prop when the exit transition is completed.
  These two callbacks allow to unmount the children when in a closed state and fully transitioned.

You can also visit the dedicated sections of some of the components:

- [Modal](modal.html#transitions)
- [Dialog](dialog.html#transitions)
- [Popper](popper.html#transitions)
- [Snackbar](snackbar.html#transitions)
- [Tooltip](tooltip.html#transitions)

## Performance & SEO

The content of transition component is mounted by default even if `in = FALSE`.
This default behavior has server-side rendering and SEO in mind.
If you render expensive component trees inside your transition it might be a good idea to change this default behavior by enabling the
`unmountOnExit` prop:

```r
Fade(`in` = FALSE, unmountOnExit = TRUE, Box())
```

As with any performance optimization this is not a silver bullet.
Be sure to identify bottlenecks first and then try out these optimization strategies.

## Transitions in Shiny apps

With `renderReact()`, the server updates the `in` prop of the rendered
element. React keeps the element, so the transition runs in both directions:

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Box(
    sx = list(p = 2),
    FormControlLabel(control = Switch.shinyInput("show", value = FALSE), label = "Show"),
    reactOutput("collapse")
  )
)

server <- function(input, output, session) {
  output$collapse <- renderReact({
    Collapse(`in` = isTRUE(input$show), Alert(severity = "info", "Collapsed content"))
  })
}

shinyApp(ui, server)
```
