---
title: "Transfer List"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Transfer List}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<div class="alert alert-primary">This page is an adaptation of the related [MUI Material UI documentation page](https://mui.com/material-ui/react-transfer-list/).</div>

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

# Transfer List

A Transfer List (or "shuttle") enables the user to move one or more list items between lists.

A transfer list has a lot of state: which items are checked, and in which list
each item is. In a Shiny app, that state naturally lives on the server. The
demos below show the layout; the [complete Shiny app](#transfer-lists-in-shiny-apps)
is at the end of the page.

## Basic transfer list

For completeness, this example includes buttons for "move all", but not every transfer list needs these.

```{r}
customList <- function(items, checked = integer()) {
  Paper(
    sx = list(width = 200, height = 230, overflow = "auto"),
    List(
      dense = TRUE,
      component = "div",
      role = "list",
      lapply(items, function(value) {
        labelId <- sprintf("transfer-list-item-%s-label", value)
        ListItemButton(
          role = "listitem",
          ListItemIcon(
            Checkbox(
              defaultChecked = value %in% checked,
              tabIndex = -1,
              disableRipple = TRUE,
              slotProps = list(input = list(`aria-labelledby` = labelId))
            )
          ),
          ListItemText(id = labelId, primary = paste("List item", value + 1))
        )
      })
    )
  )
}
moveButton <- function(label, ariaLabel, disabled = FALSE) {
  Button(sx = list(my = 0.5), variant = "outlined", size = "small", disabled = disabled, `aria-label` = ariaLabel, label)
}

muiMaterialPage(
  CssBaseline(),
  Grid(
    container = TRUE,
    spacing = 2,
    sx = list(justifyContent = "center", alignItems = "center"),
    Grid(customList(0:3)),
    Stack(
      moveButton("≫", "move all right"),
      moveButton(">", "move selected right", disabled = TRUE),
      moveButton("<", "move selected left", disabled = TRUE),
      moveButton("≪", "move all left")
    ),
    Grid(customList(4:7))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Grid from '@mui/material/Grid';
import List from '@mui/material/List';
import ListItemButton from '@mui/material/ListItemButton';
import ListItemIcon from '@mui/material/ListItemIcon';
import ListItemText from '@mui/material/ListItemText';
import Checkbox from '@mui/material/Checkbox';
import Button from '@mui/material/Button';
import Paper from '@mui/material/Paper';
import Stack from '@mui/material/Stack';

function not(a, b) {
  return a.filter((value) => !b.includes(value));
}

function intersection(a, b) {
  return a.filter((value) => b.includes(value));
}

export default function TransferList() {
  const [checked, setChecked] = React.useState([]);
  const [left, setLeft] = React.useState([0, 1, 2, 3]);
  const [right, setRight] = React.useState([4, 5, 6, 7]);

  const leftChecked = intersection(checked, left);
  const rightChecked = intersection(checked, right);

  const handleToggle = (value) => () => {
    const currentIndex = checked.indexOf(value);
    const newChecked = [...checked];

    if (currentIndex === -1) {
      newChecked.push(value);
    } else {
      newChecked.splice(currentIndex, 1);
    }

    setChecked(newChecked);
  };

  const handleAllRight = () => {
    setRight(right.concat(left));
    setLeft([]);
  };

  const handleCheckedRight = () => {
    setRight(right.concat(leftChecked));
    setLeft(not(left, leftChecked));
    setChecked(not(checked, leftChecked));
  };

  const handleCheckedLeft = () => {
    setLeft(left.concat(rightChecked));
    setRight(not(right, rightChecked));
    setChecked(not(checked, rightChecked));
  };

  const handleAllLeft = () => {
    setLeft(left.concat(right));
    setRight([]);
  };

  const customList = (items) => (
    <Paper sx={{ width: 200, height: 230, overflow: 'auto' }}>
      <List dense component="div" role="list">
        {items.map((value) => {
          const labelId = `transfer-list-item-${value}-label`;

          return (
            <ListItemButton
              key={value}
              role="listitem"
              onClick={handleToggle(value)}
            >
              <ListItemIcon>
                <Checkbox
                  checked={checked.includes(value)}
                  tabIndex={-1}
                  disableRipple
                  slotProps={{
                    input: { 'aria-labelledby': labelId },
                  }}
                />
              </ListItemIcon>
              <ListItemText id={labelId} primary={`List item ${value + 1}`} />
            </ListItemButton>
          );
        })}
      </List>
    </Paper>
  );

  return (
    <Grid
      container
      spacing={2}
      sx={{ justifyContent: 'center', alignItems: 'center' }}
    >
      <Grid>{customList(left)}</Grid>
      <Stack>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleAllRight}
          disabled={left.length === 0}
          aria-label="move all right"
        >
          ≫
        </Button>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleCheckedRight}
          disabled={leftChecked.length === 0}
          aria-label="move selected right"
        >
          &gt;
        </Button>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleCheckedLeft}
          disabled={rightChecked.length === 0}
          aria-label="move selected left"
        >
          &lt;
        </Button>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleAllLeft}
          disabled={right.length === 0}
          aria-label="move all left"
        >
          ≪
        </Button>
      </Stack>
      <Grid>{customList(right)}</Grid>
    </Grid>
  );
}
```
</details>

## Enhanced transfer list

This example exchanges the "move all" buttons for a "select all / select none" checkbox and adds a counter.

```{r}
enhancedList <- function(title, items, checked = integer()) {
  Card(
    CardHeader(
      sx = list(px = 2, py = 1),
      avatar = Checkbox(
        checked = length(checked) == length(items) && length(items) > 0,
        indeterminate = length(checked) > 0 && length(checked) < length(items),
        disabled = length(items) == 0,
        onChange = JS("() => {}"),
        slotProps = list(input = list(`aria-label` = "all items selected"))
      ),
      title = title,
      subheader = sprintf("%s/%s selected", length(checked), length(items))
    ),
    Divider(),
    List(
      sx = list(width = 200, height = 230, bgcolor = "background.paper", overflow = "auto"),
      dense = TRUE,
      component = "div",
      role = "list",
      lapply(items, function(value) {
        ListItemButton(
          role = "listitem",
          ListItemIcon(Checkbox(checked = value %in% checked, onChange = JS("() => {}"), tabIndex = -1, disableRipple = TRUE)),
          ListItemText(primary = paste("List item", value + 1))
        )
      })
    )
  )
}

muiMaterialPage(
  CssBaseline(),
  Grid(
    container = TRUE,
    spacing = 2,
    sx = list(justifyContent = "center", alignItems = "center"),
    Grid(enhancedList("Choices", 0:3, checked = c(0, 2))),
    Stack(
      moveButton(">", "move selected right"),
      moveButton("<", "move selected left", disabled = TRUE)
    ),
    Grid(enhancedList("Chosen", 4:7))
  )
)
```

<details><summary>JS code</summary>
```jsx
import * as React from 'react';
import Grid from '@mui/material/Grid';
import List from '@mui/material/List';
import Card from '@mui/material/Card';
import CardHeader from '@mui/material/CardHeader';
import ListItemButton from '@mui/material/ListItemButton';
import ListItemText from '@mui/material/ListItemText';
import ListItemIcon from '@mui/material/ListItemIcon';
import Checkbox from '@mui/material/Checkbox';
import Button from '@mui/material/Button';
import Divider from '@mui/material/Divider';
import Stack from '@mui/material/Stack';

function not(a, b) {
  return a.filter((value) => !b.includes(value));
}

function intersection(a, b) {
  return a.filter((value) => b.includes(value));
}

function union(a, b) {
  return [...a, ...not(b, a)];
}

export default function SelectAllTransferList() {
  const [checked, setChecked] = React.useState([]);
  const [left, setLeft] = React.useState([0, 1, 2, 3]);
  const [right, setRight] = React.useState([4, 5, 6, 7]);

  const leftChecked = intersection(checked, left);
  const rightChecked = intersection(checked, right);

  const handleToggle = (value) => () => {
    const currentIndex = checked.indexOf(value);
    const newChecked = [...checked];

    if (currentIndex === -1) {
      newChecked.push(value);
    } else {
      newChecked.splice(currentIndex, 1);
    }

    setChecked(newChecked);
  };

  const numberOfChecked = (items) => intersection(checked, items).length;

  const handleToggleAll = (items) => () => {
    if (numberOfChecked(items) === items.length) {
      setChecked(not(checked, items));
    } else {
      setChecked(union(checked, items));
    }
  };

  const handleCheckedRight = () => {
    setRight(right.concat(leftChecked));
    setLeft(not(left, leftChecked));
    setChecked(not(checked, leftChecked));
  };

  const handleCheckedLeft = () => {
    setLeft(left.concat(rightChecked));
    setRight(not(right, rightChecked));
    setChecked(not(checked, rightChecked));
  };

  const customList = (title, items) => (
    <Card>
      <CardHeader
        sx={{ px: 2, py: 1 }}
        avatar={
          <Checkbox
            onClick={handleToggleAll(items)}
            checked={numberOfChecked(items) === items.length && items.length !== 0}
            indeterminate={
              numberOfChecked(items) !== items.length && numberOfChecked(items) !== 0
            }
            disabled={items.length === 0}
            slotProps={{
              input: { 'aria-label': 'all items selected' },
            }}
          />
        }
        title={title}
        subheader={`${numberOfChecked(items)}/${items.length} selected`}
      />
      <Divider />
      <List
        sx={{
          width: 200,
          height: 230,
          bgcolor: 'background.paper',
          overflow: 'auto',
        }}
        dense
        component="div"
        role="list"
      >
        {items.map((value) => {
          const labelId = `transfer-list-all-item-${value}-label`;

          return (
            <ListItemButton
              key={value}
              role="listitem"
              onClick={handleToggle(value)}
            >
              <ListItemIcon>
                <Checkbox
                  checked={checked.includes(value)}
                  tabIndex={-1}
                  disableRipple
                  slotProps={{
                    input: { 'aria-labelledby': labelId },
                  }}
                />
              </ListItemIcon>
              <ListItemText id={labelId} primary={`List item ${value + 1}`} />
            </ListItemButton>
          );
        })}
      </List>
    </Card>
  );

  return (
    <Grid
      container
      spacing={2}
      sx={{ justifyContent: 'center', alignItems: 'center' }}
    >
      <Grid>{customList('Choices', left)}</Grid>
      <Stack>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleCheckedRight}
          disabled={leftChecked.length === 0}
          aria-label="move selected right"
        >
          &gt;
        </Button>
        <Button
          sx={{ my: 0.5 }}
          variant="outlined"
          size="small"
          onClick={handleCheckedLeft}
          disabled={rightChecked.length === 0}
          aria-label="move selected left"
        >
          &lt;
        </Button>
      </Stack>
      <Grid>{customList('Chosen', right)}</Grid>
    </Grid>
  );
}
```
</details>

## Limitations

The component comes with a couple of limitations:

- It only works on desktop.
  If you have a limited amount of options to select, prefer the [Autocomplete](autocomplete.html) component.
  If mobile support is important for you, have a look at [#27579](https://github.com/mui/material-ui/issues/27579).
- There are no high-level components exported from npm. The demos are based on composition.
  If this is important for you, have a look at [#27579](https://github.com/mui/material-ui/issues/27579).

## Transfer lists in Shiny apps

The server keeps the items of each list and the checked items in
`reactiveVal()`s. The lists are rendered with `renderReact()`, so they update
when the state changes:

- A click on a row sends a single custom event with the clicked item as payload
  (`Shiny.setInputValue(..., {priority: 'event'})`). This is lighter than one
  `.shinyInput()` per row, as the rows are re-generated on every move.
- The checkboxes are plain `Checkbox()` controlled by `checked`. The no-op
  `onChange` marks them as controlled for React; the row's click does the
  toggling.
- The move buttons are `Button.shinyInput()`s. Their `disabled` state is
  updated from the server with `updateButton.shinyInput()`.

```{r eval=FALSE}
library(shiny)
library(muiMaterial)

transferRow <- function(value, checked, event) {
  ListItemButton(
    role = "listitem",
    onClick = JS(sprintf("() => Shiny.setInputValue('%s', %s, {priority: 'event'})", event, value)),
    ListItemIcon(Checkbox(checked = value %in% checked, onChange = JS("() => {}"), tabIndex = -1, disableRipple = TRUE)),
    ListItemText(primary = paste("List item", value + 1))
  )
}
listPaper <- function(outputId) {
  Paper(sx = list(width = 200, height = 230, overflow = "auto"), List(dense = TRUE, component = "div", role = "list", reactOutput(outputId)))
}

ui <- muiMaterialPage(
  CssBaseline(),
  Grid(
    container = TRUE,
    spacing = 2,
    sx = list(justifyContent = "center", alignItems = "center", p = 2),
    Grid(listPaper("left")),
    Stack(
      Button.shinyInput("all_right", "≫", sx = list(my = 0.5), variant = "outlined", size = "small"),
      Button.shinyInput("checked_right", ">", sx = list(my = 0.5), variant = "outlined", size = "small"),
      Button.shinyInput("checked_left", "<", sx = list(my = 0.5), variant = "outlined", size = "small"),
      Button.shinyInput("all_left", "≪", sx = list(my = 0.5), variant = "outlined", size = "small")
    ),
    Grid(listPaper("right"))
  )
)

server <- function(input, output, session) {
  left <- reactiveVal(0:3)
  right <- reactiveVal(4:7)
  checked <- reactiveVal(integer())

  toggle <- function(value) {
    checked(if (value %in% checked()) setdiff(checked(), value) else c(checked(), value))
  }
  observeEvent(input$left_toggle, toggle(input$left_toggle))
  observeEvent(input$right_toggle, toggle(input$right_toggle))

  observeEvent(input$all_right, {
    right(c(right(), left()))
    left(integer())
  })
  observeEvent(input$all_left, {
    left(c(left(), right()))
    right(integer())
  })
  observeEvent(input$checked_right, {
    moved <- intersect(checked(), left())
    right(c(right(), moved))
    left(setdiff(left(), moved))
    checked(setdiff(checked(), moved))
  })
  observeEvent(input$checked_left, {
    moved <- intersect(checked(), right())
    left(c(left(), moved))
    right(setdiff(right(), moved))
    checked(setdiff(checked(), moved))
  })

  # Disable the buttons that would do nothing
  observe({
    updateButton.shinyInput(inputId = "all_right", disabled = length(left()) == 0)
    updateButton.shinyInput(inputId = "all_left", disabled = length(right()) == 0)
    updateButton.shinyInput(inputId = "checked_right", disabled = length(intersect(checked(), left())) == 0)
    updateButton.shinyInput(inputId = "checked_left", disabled = length(intersect(checked(), right())) == 0)
  })

  output$left <- renderReact(lapply(left(), transferRow, checked = checked(), event = "left_toggle"))
  output$right <- renderReact(lapply(right(), transferRow, checked = checked(), event = "right_toggle"))
}

shinyApp(ui, server)
```

Run another version with `muiMaterialExample("TransferList")`.
