---
title: "Overlays with .triggerId"
format:
  html:
    minimal: true
vignette: >
  %\VignetteIndexEntry{Overlays with .triggerId}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE}
library(muiMaterial)
library(shiny)
```

Overlays (dialogs, drawers, menus, popovers) have an open state. In React, it
lives in a `useState()` hook of the parent component, which R code cannot
write. muiMaterial offers several ways to handle it:

| Approach | Open state lives in | Server code | Works in Quarto / static HTML |
|---|---|---|---|
| **`.triggerId()` wrappers** (this page) | the browser | none | yes |
| [reactRouter](https://felixluginbuhl.com/reactRouter/) routes (see [Dialog](dialog.html)) | the URL | none | yes |
| `.shinyInput()` wrappers + `update*.shinyInput()` | the Shiny server | yes | no |

The `.triggerId()` wrappers cover the most common case: **open an overlay when
an element is clicked**. They are available for six components:

| Function | Opens | Closes |
|---|---|---|
| `Dialog.triggerId()` | a dialog | backdrop click, Escape |
| `Modal.triggerId()` | a modal | backdrop click, Escape |
| `Drawer.triggerId()` | a temporary drawer | backdrop click, Escape, a click on a link inside |
| `SwipeableDrawer.triggerId()` | a swipeable drawer | same as `Drawer.triggerId()`, and swiping |
| `Menu.triggerId()` | a menu, next to the trigger | click outside, Escape, a click on an item |
| `Popover.triggerId()` | a popover, next to the trigger | click outside, Escape |

## How it works

Give any element an `id`, and pass the same id as the `triggerId` of the
overlay. A click on the element opens the overlay:

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(id = "open-dialog", variant = "outlined", "Open dialog"),
  Dialog.triggerId(
    triggerId = "open-dialog",
    DialogTitle("Hello"),
    DialogContent(DialogContentText("Open/close is managed entirely in the browser."))
  )
)
```

- The trigger can be **any element with an `id`**: a muiMaterial component, an
  HTML tag, or a Shiny widget such as `shiny::actionButton()`.
- The overlay can be anywhere on the page. It does not need to be next to its
  trigger.
- The binding survives re-renders: a trigger rendered later with
  `renderUI()`/`uiOutput()` (or replaced by it) keeps working.
- Several triggers cannot share one id, but several overlays can each have
  their own trigger.

## The six wrappers

```{r}
muiMaterialPage(
  useMaterialIconsFilled = TRUE,
  CssBaseline(),
  Stack(
    direction = "row",
    spacing = 1,
    sx = list(flexWrap = "wrap"),
    Button(id = "t-dialog", variant = "outlined", "Dialog"),
    Button(id = "t-modal", variant = "outlined", "Modal"),
    Button(id = "t-drawer", variant = "outlined", "Drawer"),
    Button(id = "t-swipeable", variant = "outlined", "Swipeable drawer"),
    Button(id = "t-menu", variant = "outlined", "Menu"),
    Button(id = "t-popover", variant = "outlined", "Popover")
  ),
  Dialog.triggerId("t-dialog", DialogTitle("Dialog"), DialogContent("Press Escape to close.")),
  Modal.triggerId(
    "t-modal",
    Box(
      sx = list(position = "absolute", top = "50%", left = "50%", transform = "translate(-50%, -50%)",
                width = 300, bgcolor = "background.paper", boxShadow = 24, p = 4),
      Typography("Modal content")
    )
  ),
  Drawer.triggerId("t-drawer", anchor = "left", width = 250, Box(sx = list(p = 2), Typography("Drawer content"))),
  SwipeableDrawer.triggerId("t-swipeable", anchor = "bottom", width = "auto", Box(sx = list(p = 2), Typography("Swipe down to close"))),
  Menu.triggerId("t-menu", MenuItem("Profile"), MenuItem("My account"), MenuItem("Logout")),
  Popover.triggerId(
    "t-popover",
    anchorOrigin = list(vertical = "bottom", horizontal = "left"),
    Typography(sx = list(p = 2), "Popover content")
  )
)
```

## Options

All the other arguments are passed to the MUI component as props, so the
options of the MUI documentation apply: `anchor`, `maxWidth`, `anchorOrigin`,
`slotProps`, `sx`, ... A few arguments are specific to the wrappers:

- `Drawer.triggerId()` and `SwipeableDrawer.triggerId()`:
  - `width` sizes the drawer paper (280 by default). Use `width = "auto"` for
    top and bottom drawers.
  - `closeOnLinkClick = FALSE` keeps the drawer open when a link (`<a>`) inside
    it is clicked. By default, the drawer closes, which suits navigation drawers.
  - `sx` styles the drawer **root**, like on any other component. Style the paper
    with `slotProps = list(paper = list(sx = ...))`.
- `Menu.triggerId()`: `closeOnItemClick = FALSE` keeps the menu open after a
  click, for example for a menu of checkboxes.

```{r}
muiMaterialPage(
  CssBaseline(),
  Button(id = "filters-menu", variant = "outlined", "Filters"),
  Menu.triggerId(
    triggerId = "filters-menu",
    closeOnItemClick = FALSE,
    lapply(c("Open", "In progress", "Closed"), function(status) {
      MenuItem(FormControlLabel(control = Checkbox(defaultChecked = TRUE), label = status))
    })
  )
)
```

## Callbacks

The open state belongs to the wrapper: passing `open` (or `anchorEl`) has no
effect. Callbacks you pass are **composed** with the wrapper's own handlers:
the wrapper updates its state first, then calls yours. For example, send the
closing of a dialog to the server with `triggerEvent()`:

```{r, eval = FALSE}
Dialog.triggerId(
  triggerId = "open-dialog",
  onClose = triggerEvent("dialog_closed"),
  DialogTitle("Hello")
)
```

The same applies to `onClick` and `onClose` of `Menu.triggerId()`, and to
`onOpen`/`onClose` of `SwipeableDrawer.triggerId()`.

## Limitations

- A `.triggerId()` overlay closes on its built-in events only (see the tables
  above). A button inside a dialog does not close it. For dialogs with
  "Cancel"/"Confirm" buttons, keep the open state in the URL with
  [reactRouter](dialog.html) or on the server with `Dialog.shinyInput()`.
- The server cannot open or close a `.triggerId()` overlay. Use the
  `.shinyInput()` variant when the server must decide.

## In a Shiny app

The trigger can be a Shiny widget, and it can be rendered by the server:

```{r, eval = FALSE}
library(shiny)
library(muiMaterial)

ui <- muiMaterialPage(
  CssBaseline(),
  Box(sx = list(p = 2), uiOutput("toolbar")),
  Drawer.triggerId(
    triggerId = "open-settings",
    anchor = "right",
    Box(sx = list(p = 2), Typography(variant = "h6", "Settings"), Switch.shinyInput("compact", value = FALSE))
  )
)

server <- function(input, output, session) {
  # The trigger is rendered (and re-rendered) by the server
  output$toolbar <- renderUI({
    IconButton(id = "open-settings", `aria-label` = "settings", shiny::icon("gear"))
  })
}

shinyApp(ui, server)
```

Run complete examples with `muiMaterialExample("DrawerTriggerId")` and
`muiMaterialExample("Menu")`.
